---
title: "2_curate_taxonomy - Grace Lewin"
output: html_document
date: "2023-08-09"
---


curate taxonomy for 12s and 16s


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

This section reads in the dataset:
```{r}
#gaviota_12s <- read_csv(here("metabarcoding_scripts_GL","gaviota_metabarcoding_results.csv"))

gaviota_12s_16s <- read_csv(here("metabarcoding_scripts_GL","gaviota_metabarcoding_results.csv"))

```

This section takes out human contamination and jaguar match:
```{r}
gaviota_12s_16s <- gaviota_12s_16s[!grepl("Homo sapiens", gaviota_12s_16s$Including_domestic.Local_species),]
gaviota_12s_16s <- gaviota_12s_16s[!grepl("Panthera onca", gaviota_12s_16s$Including_domestic.Local_species),]
```


This section splits the data into 1% and 0.5% filters

```{r}
#filter percentage_filter by 0.5 and select relevant columns to look at species determined though metabarcoding and visual inspection notes.
gaviota_12s_16s_0.5 <- gaviota_12s_16s %>% 
  filter(Percentage_filter == 0.5) 
  # select(c("Sample","Best_match_references", "Local_taxa", "Local_species", , "Visual_Inspection_categories", "Percentage_filter", "common_name"))
  
#filter percentage_filter by 1 and select relevant columns to look at species determined though metabarcoding and visual inspection notes.
gaviota_12s_16s_1 <- gaviota_12s_16s %>% 
  filter(Percentage_filter == 1)
  # select(c("Sample", "Best_match_references", "Local_taxa", "Local_species", "Visual_Inspection", "Visual_Inspection_categories", "Percentage_filter", "common_name"))
```



This section of the pipeline starts with the cleaned and combined "Lewin_Gaviota_mammal_scat_data_Results_Dec2024_v2" and "UCSB_Rerun_Samples_12s_March2025" file (either 1% or 0.5% filtering) and adds columns with taxonomic info across varying levels of specificity.

## choose 1% or 0.5% filtered data
```{r}
gaviota <- gaviota_12s_16s_0.5
```


## Curate the "Including_domestic.Local_species" column to represent the single most accurate species assignment
```{r}
#unique(gaviota$Including_domestic.Local_species)

gaviota$curated_species = NA

gaviota <- gaviota %>%
  mutate(curated_species = case_when(
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus" ~ "Canis latrans", # coyote
    
    Including_domestic.Local_species %in% "Phalacrocorax penicillatus, Phalacrocorax brasilianus, Phalacrocorax pelagicus" 
      ~ "Urile penicillatus", # Brandt's cormorant
    
    Including_domestic.Local_species %in% "Phalacrocorax pelagicus, Phalacrocorax penicillatus" 
      ~ "Urile penicillatus", # Brandt's cormorant
    
    Including_domestic.Local_species %in% "Microtus californicus" 
      ~ "Microtus californicus", # Californian vole
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei" ~ "Peromyscus", #Peromyscus - deer mouse
    
    Including_domestic.Local_species %in% "Sylvilagus bachmani" ~ "Sylvilagus bachmani", # Brush rabbit
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Arborimus albipes, Arborimus longicaudus, Arborimus pomo, Myodes californicus, Lemmiscus curtatus, Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii, Ondatra zibethicus, Phenacomys intermedius, Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis, Onychomys leucogaster, Onychomys torridus, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei, Reithrodontomys raviventris, Reithrodontomys megalotis, Sigmodon arizonae, Sigmodon hispidus" ~ "Peromyscus sp", # 	Cricetidae (mouse or vole) --  choosing peromyscus sp
      
    Including_domestic.Local_species %in% "Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis" ~ "Neotoma", # Neotoma - woodrat
      
    Including_domestic.Local_species %in% "Bos taurus, Bos indicus, Bos javanicus, Bos frontalis" ~ "Bos taurus", # Bovine
      
    Including_domestic.Local_species %in% "Gallus gallus" ~ "Gallus gallus", #chicken
      
    Including_domestic.Local_species %in% "Urocyon cinereoargenteus, Urocyon littoralis" ~ "Urocyon cinereoargenteus", # grey fox
      
    Including_domestic.Local_species %in% "Glaucomys oregonensis, Glaucomys sabrinus, Callospermophilus lateralis" ~ "Sciuridae", # flying squirrel/golden mantle ground squirrel
      
    Including_domestic.Local_species %in% "Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii" ~ "Microtus californicus", # vole
    Including_domestic.Local_species %in% "Thomomys bottae" ~"Thomomys bottae", # botta's pocket gopher
    
    Including_domestic.Local_species %in% "Lynx rufus, Puma concolor" ~ "Lynx rufus", # bobcat
      
    Including_domestic.Local_species %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Odocoileus hemionus", # mule deer
      
    Including_domestic.Local_species %in% "Sus scrofa" ~ "Sus scrofa", #wild boar
      
    Including_domestic.Local_species %in% "Chloris sinica, Spinus pinus, Plectrophenax nivalis, Zonotrichia leucophrys, Zonotrichia atricapilla, Zonotrichia albicollis, Sturnella neglecta, Sturnella magna, Serinus canaria" ~ "Zonotrichia sp", # Passeriformes
      
    Including_domestic.Local_species %in% "Peromyscus eremicus" ~ "Peromyscus eremicus", # cactus mouse
      
    Including_domestic.Local_species %in% "Zalophus californianus" ~ "Zalophus californianus", # sea lion
    
    Including_domestic.Local_species %in% "Canis lupus" ~ "Canis latrans", # coyote
      
    Including_domestic.Local_species %in% "Peromyscus crinitus, Peromyscus eremicus" ~ "Peromyscus", # Peromyscus
      
    Including_domestic.Local_species %in% "Callipepla californica, Callipepla gambelii" ~ "Callipepla californica", # CA quail 
      
    Including_domestic.Local_species %in% "Alosa sapidissima, Dorosoma petenense" ~ "Alosa sapidissima", # American shad
      
    Including_domestic.Local_species %in% "Leuconotopicus albolarvatus, Jynx torquilla, Melanerpes lewis, Melanerpes erythrocephalus, Melanerpes formicivorus, Melanerpes uropygialis, Sphyrapicus thyroideus, Sphyrapicus varius, Sphyrapicus nuchalis, Sphyrapicus ruber, Picoides arcticus, Dryobates pubescens, Dryobates nuttallii, Dryobates scalaris, Picoides villosus, Colaptes auratus, Colaptes chrysoides, Dryocopus pileatus" ~ "Melanerpes formicivorus", # acorn woodpecker
      
    Including_domestic.Local_species %in% "Plestiodon gilberti" ~ "Plestiodon gilberti", # Gilbert's skink
      
    Including_domestic.Local_species %in% "Taxidea taxus" ~ "Taxidea taxus", # American badger
      
    Including_domestic.Local_species %in% "Microtus townsendii" ~ "Microtus townsendii", # Microtus - townsend's vole
      
    Including_domestic.Local_species %in% "Neotoma macrotis" ~ "Neotoma macrotis", # big eared woodrat
    
    Including_domestic.Local_species %in% "Aphelocoma insularis, Aphelocoma californica, Aphelocoma woodhouseii" ~ "Aphelocoma californica", # CA scrub jay
      
    Including_domestic.Local_species %in% "Puma concolor" ~ "Puma concolor", # puma
      
    Including_domestic.Local_species %in% "Meleagris gallopavo, Gallus gallus" ~ "Meleagris gallopavo", # wild turkey
      
    Including_domestic.Local_species %in% "Mephitis mephitis" ~ "Mephitis mephitis", # striped skunk
      
    Including_domestic.Local_species %in% "Vulpes vulpes" ~ "Vulpes vulpes", # red fox
      
    Including_domestic.Local_species %in% "Chaetodipus californicus" ~ "Chaetodipus californicus", # California pocket mouse
      
    Including_domestic.Local_species %in% "Mus musculus" ~ "Mus musculus", # house mouse
    
    Including_domestic.Local_species %in% "Rattus rattus" ~ "Rattus rattus", # black rat
      
    Including_domestic.Local_species %in% "Peromyscus maniculatus, Peromyscus keeni" ~ "Peromyscus sonoriensis", # deer mice
      
    Including_domestic.Local_species %in% "Sylvilagus floridanus, Sylvilagus nuttallii" ~ "Sylvilagus bachmani", # brush rabbit
      
    Including_domestic.Local_species %in% "Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii" ~ "Microtus sp", # vole
      
    Including_domestic.Local_species %in% "Lemmiscus curtatus, Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii, Ondatra zibethicus, Onychomys leucogaster, Synaptomys borealis, Peromyscus maniculatus, Peromyscus keeni" ~ "Peromyscus sonoriensis", # cricetidae - some kind of rodent -- (choosing microtus sp)
   
     Including_domestic.Local_species %in% "Neotoma lepida" ~ "Neotoma lepida", # desert woodrat
    
     Including_domestic.Local_species %in% "Canis latrans, Canis lupus, Urocyon cinereoargenteus, Urocyon littoralis, Vulpes macrotis, Vulpes vulpes" ~ "Canis latrans", # coyote
    
    Including_domestic.Local_species %in% "Turdus obscurus, Turdus migratorius" ~ "Turdus migratorius", #American robin
    
    Including_domestic.Local_species %in% "Spizella passerina, Passerella iliaca, Melospiza melodia" ~ "Melospiza melodia", #Song sparrow, could  be chipping or fox sparrow too though.
    
    Including_domestic.Local_species %in% "Ascaphus truei, Incilius alvarius, Anaxyrus boreas, Anaxyrus californicus, Anaxyrus canorus, Anaxyrus cognatus, Anaxyrus exsul, Anaxyrus microscaphus, Anaxyrus punctatus, Anaxyrus woodhousii, Pseudacris cadaverina, Pseudacris hypochondriaca, Pseudacris regilla, Pseudacris sierra, Scaphiopus couchii, Spea hammondii, Spea intermontana, Rana aurora, Lithobates berlandieri, Rana boylii, Rana cascadae, Lithobates catesbeianus, Rana draytonii, Rana luteiventris, Rana muscosa, Lithobates pipiens, Rana pretiosa, Rana sierrae, Lithobates sphenocephalus, Lithobates yavapaiensis, Xenopus laevis" ~ "Anura", #frog or toad
    
    Including_domestic.Local_species %in% "Phoca fasciata, Phoca vitulina" ~ "Phoca vitulina", #harbor seal
    
    Including_domestic.Local_species %in% "Nucifraga columbiana, Corvus brachyrhynchos, Corvus corax" ~ "Corvidae", #corvidae -- Clark's nutcracker, American crow, Common raven
    
    Including_domestic.Local_species %in% "Pipilo maculatus" ~ "Pipilo maculatus", #spotted towhee
    
    Including_domestic.Local_species %in% "Onychomys torridus" ~ "Onychomys torridus", #southern grasshopper mouse
    
    Including_domestic.Local_species %in% "Arctocephalus townsendi" ~ "Arctocephalus townsendi", #guadelupe fur seal
    
    Including_domestic.Local_species %in% "Callospermophilus lateralis" ~ "Callospermophilus lateralis", #golden-mantled ground squirrel
    
    Including_domestic.Local_species %in% "Marmota flaviventris" ~ "Marmota flaviventris", #yellow-bellied marmot
    
    Including_domestic.Local_species %in% "Microtus montanus" ~ "Microtus montanus" #montane vole
  ))

# for the 16s run
gaviota <- gaviota %>%
  mutate(curated_species = case_when(
    Barcode_region == "16" & Best_match_references %in% "Bos taurus" ~ "Bos", # 100
    Barcode_region == "16" & Best_match_references %in% "Lynx rufus" ~ "Lynx rufus", # 100
    #Best_match_references %in% "Heliconoides inflatus" ~ "Heliconoides inflatus", # sea butterfly snail
    #Best_match_references %in% "Odocoileus virginianus" ~ "white tailed deer", # 100
    #Best_match_references %in% "Crotalus viridis" ~ "prairie rattlesnake", # 100
    Barcode_region == "16" & Best_match_references %in% "Puma concolor" ~ "Puma concolor", # puma
    Barcode_region == "16" & Best_match_references %in% "Rattus rattus" ~ "Rattus rattus", # rat
    #Best_match_references %in% "Gallus sonneratii" ~ "chicken", # 100
    Barcode_region == "16" & Best_match_references %in% "Callipepla californica" ~ "Callipepla californica", # California quail
    #Best_match_references %in% "Scomber japonicus" ~ "chub mackerel", # 100
    Barcode_region == "16" & Best_match_references %in% "Lamna ditropis" ~ "Lamna ditropis", #salmon shark
    Barcode_region == "16" & Best_match_references %in% "Magnisudis atlantica" ~ "Magnisudis atlantica", # duckbill barracuda
    Barcode_region == "16" & Best_match_references %in% "Ophiodon elongatus" ~ "Ophiodon elongatus", # lingcod
    Barcode_region == "16" & Best_match_references %in% "Thysanoessa inermis" ~ "Thysanoessa inermis", # Krill
    #Best_match_references %in% "Sebastes serranoides" ~ "Sebastes sp", # olive rockfish
    Barcode_region == "16" & Best_match_references %in% "Clupea pallasii" ~ "Clupea pallasii", # pacific herring
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa" ~ "Sus scrofa", #wild boar
    #Best_match_references %in% "Canis rufus" ~ "canid", #
    Barcode_region == "16" & Best_match_references %in% "Pilaira anomala" ~ "Pilaira anomala", #mold
    Barcode_region == "16" & Best_match_references %in% "Reithrodontomys megalotis" ~ "Reithrodontomys megalotis", #western harvest mouse
    Barcode_region == "16" & Best_match_references %in% "Microtus californicus" ~ "Microtus californicus", #California vole
    Barcode_region == "16" & Best_match_references %in% "Chaetodipus californicus" ~ "Chaetodipus californicus", #California pocket mouse
    Barcode_region == "16" & Best_match_references %in% "Zalophus californianus" ~ "Zalophus californianus", #California sea lion
    Barcode_region == "16" & Best_match_references %in% "Canis latrans" ~ "Canis latrans", #coyote
    Barcode_region == "16" & Best_match_references %in% "Neotoma lepida" ~ "Neotoma lepida", #desert woodrat
    #Best_match_references %in% "Meleagris gallopavo" ~ "turkey", #
    Barcode_region == "16" & Best_match_references %in% "Otospermophilus beecheyi" ~ "Otospermophilus beecheyi", #California ground squirrel
    Barcode_region == "16" & Best_match_references %in% "Taxidea taxus" ~ "Taxidea taxus", #american badger
    Barcode_region == "16" & Best_match_references %in% "Neotoma macrotis" ~ "Neotoma macrotis", #big-eared woodrat
    #Best_match_references %in% "Ravinia vagabunda" ~ "flesh fly", #
    Barcode_region == "16" & Best_match_references %in% "Canis lupus" ~ "Canis latrans", # canid
    Barcode_region == "16" & Best_match_references %in% "Forficula auricularia" ~ "Forficula auricularia", #European earwig
    Barcode_region == "16" & Best_match_references %in% "Mephitis mephitis" ~ "Mephitis mephitis", #striped skunk
    #Best_match_references %in% "Paralucilia pseudolyrcea" ~ "blow fly", #
    Barcode_region == "16" & Best_match_references %in% "Urocyon cinereoargenteus" ~ "Urocyon cinereoargenteus", #grey fox
    Barcode_region == "16" & Best_match_references %in% "Pandalus jordani" ~ "Pandalus jordani", #ocean pink shrimp
    #Best_match_references %in% "Merluccius productus" ~ "north pacific hake", #
    Barcode_region == "16" & Best_match_references %in% "Peromyscus maniculatus" ~ "Peromyscus sonoriensis", #deer mouse
    Barcode_region == "16" & Best_match_references %in% "Pituophis catenifer" ~ "Pituophis catenifer", #gopher snake
    Barcode_region == "16" & Best_match_references %in% "Sylvilagus bachmani" ~ "Sylvilagus bachmani", #brush rabbit
    Barcode_region == "16" & Best_match_references %in% "Plestiodon skiltonianus" ~ "Plestiodon skiltonianus", #western skink
    Barcode_region == "16" & Best_match_references %in% "Anchoa ischana" ~ "Anchoa ischana", #sharpnose anchovy
    Barcode_region == "16" & Best_match_references %in% "Gonatopsis borealis" ~ "Gonatopsis borealis", #Boreopacific armhook squid
    Barcode_region == "16" & Best_match_references %in% "Leucocarbo chalconotus" ~ "Urile pelagicus", # pelagic cormorant species, checked with BLAST. 3rd result, 92.99%match
    Barcode_region == "16" & Best_match_references %in% "Pseudopaludicola boliviana" ~ "Pseudopaludicola", # frog, Leptodactylidae family, found in regions in south america???
    Barcode_region == "16" & Best_match_references %in% "Microtus chrotorrhinus" ~ "Microtus sp", # rock vole, found in eastern US
    Barcode_region == "16" & Best_match_references %in% "Picoides tridactylus" ~ "Dryobates pubescens", # eurasian three-toed woodpecker, not found in US --- changed to Dryobates pubescens, downy woodpecker
    Barcode_region == "16" & Best_match_references %in% "Thomomys bottae" ~ "Thomomys bottae", # Botta's pocket gopher
    Barcode_region == "16" & Best_match_references %in% "Neodon bershulaensis" ~ "Microtus sp", # Bershula mountain vole, found in China -- BLAST shows microtus species second match
    #Best_match_references %in% "Chloropicon sieburthii" ~ "phytoplankton", # 90
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus" ~ "Creophilus maxillosus", # hairy rove beetle -- only shows up once
    Barcode_region == "16" & Best_match_references %in% "Rhipicephalus sanguineus" ~ "Rhipicephalus sanguineus", # Brown dog tick
    Barcode_region == "16" & Best_match_references %in% "Phakopsora meibomiae" ~ "Phakopsora meibomiae", # fungi
    Barcode_region == "16" & Best_match_references %in% "Tettigonia orientalis" ~ "Stenopelmatus sp", # bush cricket, but this species is found in Japan ---BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Isogenoides varians" ~ "Isogenoides varians", # stonefly
    Barcode_region == "16" & Best_match_references %in% "Setvena wahkeena" ~ "Stenopelmatus sp", # stonefly (found in socal?) -- BLAST - cricket match
    Barcode_region == "16" & Best_match_references %in% "Xiphidiopsis gurneyi" ~ "Stenopelmatus sp", #bush cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Abronia graminea" ~ "Elgaria multicarinata", # alligator lizard (found in socal?) -- BLAST - southern aligator lizard
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini" ~ "Stenopelmatus sp", # camel cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Chloropicon laureae" ~ "Chloropicon sp", # green algae
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae" ~ "Stenopelmatus sp", #Scarabaeidae (scarab) -- changed to species that came up with this species in results -- BLAST- cricket
    Barcode_region == "16" & Best_match_references %in% "Stenopelmatus fuscus" ~ "Stenopelmatus fuscus", # jerusalem cricket
    Barcode_region == "16" & Best_match_references %in% "Gymnaetoides acutilobatus" ~ "Stenopelmatus sp", # endemic to china??? cave cricket --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Microdipodops pallidus" ~ "Microdipodops pallidus", # pale kangaroo mouse
    Barcode_region == "16" & Best_match_references %in% "Onthophagus crinitus" ~ "Onthophagus crinitus", # dung beetle
    Barcode_region == "16" & Best_match_references %in% "Pheidole deima" ~ "Veromessor andrei", # ant (found in socal?) --- BLAST - Veromessor andrei - harvestor ant
    Barcode_region == "16" & Best_match_references %in% "Hazenia capsulata" ~ "Hazenia capsulata", # green algae
    Barcode_region == "16" & Best_match_references %in% "Anadara broughtonii" ~ "Anadara broughtonii", # Inflated ark clam
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea" ~ "Cavenderia subdiscoidea", # slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis" ~ "Cloropicon sp", # 70

    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos javanicus, Bos indicus, Bos grunniens, Bos grunniens x Bos taurus" ~ "Bos",
    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos primigenius, Bos indicus" ~ "Bos", #cow
    Barcode_region == "16" & Best_match_references %in% "Canis rufus, Canis lupus, Canis latrans" ~ "Canis latrans", # coyote
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea, Cavenderia pseudoaureostipes" ~ "Cavenderia sp", #slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Chloropicon sp", #green algae
    Barcode_region == "16" & Best_match_references %in% "Chloropicon sieburthii, Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Chloropicon sp", #green algae
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus, Anthrenus scrophulariae" ~ "Anthrenus scrophulariae", # Common carpet beetle/Buffalo carpet beetle
    Barcode_region == "16" & Best_match_references %in% "Crotalus viridis, Crotalus oreganus" ~ "Crotalus oreganus", #western rattlesnake
    Barcode_region == "16" & Best_match_references %in% "Gallus sonneratii, Gallus gallus" ~ "Gallus sp", # wild turkey probably, or chicken
    Barcode_region == "16" & Best_match_references %in% "Heliconoides inflatus, Cathaica polystigma, Nesiohelix moreletiana, Karaftohelix maackii, Atlanta turriculata, Aceria guerreronis, Hahnia ngai, Grylloprimevala jilina, Aconurella prolixa, Processina dashahensis, Tapinoma indicum, Homo sapiens" ~ "Heliconoides inflatus", # Heliconoides inflatus -- Sea butterfly (range is not Socal, but Mexico, and it's the closest species in the list) 
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae, Paratriodonta lentula" ~ "Paratriodonta lentula", #Scarabaeidae (scarab)
    Barcode_region == "16" & Best_match_references %in% "Meleagris gallopavo, Labidesthes sicculus" ~ "Meleagris gallopavo", #wild turkey
    Barcode_region == "16" & Best_match_references %in% "Merluccius productus, Merluccius angustimanus" ~ "Merluccius productus", #North Pacific Hake
    Barcode_region == "16" & Best_match_references %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Odocoileus hemionus", #mule deer
    Barcode_region == "16" & Best_match_references %in% "Paralucilia pseudolyrcea, Paralucilia paraensis, Hemilucilia semidiaphana, Calliphora vomitoria, Calliphora vicina, Calliphora sinensis, Calliphora nigribarbis, Calliphora calliphoroides" ~ "Diptera", # fly
    Barcode_region == "16" & Best_match_references %in% "Ravinia vagabunda, Ravinia stimulans, Ravinia derelicta" ~ "Ravinia", #flesh fly
    Barcode_region == "16" & Best_match_references %in% "Scomber japonicus, Scomber colias, Scomber australasicus"~ "Scomber japonicus", #pacific mackerel
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa, Thryonomys swinderianus" ~ "Sus scrofa", #wild pig
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini, Onthophagus obscurior" ~ "Onthophagus obscurior", #dung beetle
    Barcode_region == "16" & Best_match_references %in% "Sebastes serranoides, Sebastes melanops, Sebastes flavidus" ~ "Sebastes sp",  #rockfish
    Barcode_region == "16" & Best_match_references %in% "" ~ "no match",
    TRUE ~ curated_species
    ))


 

checkna <- gaviota[is.na(gaviota$curated_species), ] 
#Best_match_references ==rm(checkna)

```

## Curate common names

```{r}
gaviota$common_name = NA

gaviota <- gaviota %>%
  mutate(common_name = case_when(
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus" ~ "Coyote", # coyote
    
    Including_domestic.Local_species %in% "Phalacrocorax penicillatus, Phalacrocorax brasilianus, Phalacrocorax pelagicus" 
      ~ "Brandt's cormorant", # Brant's cormorant
    
    Including_domestic.Local_species %in% "Phalacrocorax pelagicus, Phalacrocorax penicillatus" 
      ~ "Brandt's cormorant", # Brant's cormorant
    
    Including_domestic.Local_species %in% "Microtus californicus" 
      ~ "California vole", # Californian vole
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei" ~ "Deer mouse", #Peromyscus - deer mouse
    
    Including_domestic.Local_species %in% "Sylvilagus bachmani" ~ "Brush rabbit", # Brush rabbit
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Arborimus albipes, Arborimus longicaudus, Arborimus pomo, Myodes californicus, Lemmiscus curtatus, Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii, Ondatra zibethicus, Phenacomys intermedius, Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis, Onychomys leucogaster, Onychomys torridus, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei, Reithrodontomys raviventris, Reithrodontomys megalotis, Sigmodon arizonae, Sigmodon hispidus" ~ "Deer mouse", # 	Cricetidae (mouse or vole) --  choosing peromyscus (deer mouse)
      
    Including_domestic.Local_species %in% "Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis" ~ "Woodrat", # Neotoma - woodrat
      
    Including_domestic.Local_species %in% "Bos taurus, Bos indicus, Bos javanicus, Bos frontalis" ~ "Cattle", # Bovine
      
    Including_domestic.Local_species %in% "Gallus gallus" ~ "Chicken", #chicken
      
    Including_domestic.Local_species %in% "Urocyon cinereoargenteus, Urocyon littoralis" ~ "Grey fox", # grey fox
      
    Including_domestic.Local_species %in% "Glaucomys oregonensis, Glaucomys sabrinus, Callospermophilus lateralis" ~ "Sciuridae", # flying squirrel/golden mantle ground squirrel
      
    Including_domestic.Local_species %in% "Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii" ~ "California vole", # vole
    
    Including_domestic.Local_species %in% "Thomomys bottae" ~"Botta's pocket gopher", # botta's pocket gopher
    
    Including_domestic.Local_species %in% "Lynx rufus, Puma concolor" ~ "Bobcat", # bobcat
      
    Including_domestic.Local_species %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Mule deer", # mule deer
      
    Including_domestic.Local_species %in% "Sus scrofa" ~ "Wild boar", #wild boar
      
    Including_domestic.Local_species %in% "Chloris sinica, Spinus pinus, Plectrophenax nivalis, Zonotrichia leucophrys, Zonotrichia atricapilla, Zonotrichia albicollis, Sturnella neglecta, Sturnella magna, Serinus canaria" ~ "Sparrow", # Sparrow
      
    Including_domestic.Local_species %in% "Peromyscus eremicus" ~ "Cactus mouse", # cactus mouse
      
    Including_domestic.Local_species %in% "Zalophus californianus" ~ "California sea lion", # sea lion
    
    Including_domestic.Local_species %in% "Canis lupus" ~ "Coyote", # coyote
      
    Including_domestic.Local_species %in% "Peromyscus crinitus, Peromyscus eremicus" ~ "Deer mouse", # Peromyscus
      
    Including_domestic.Local_species %in% "Callipepla californica, Callipepla gambelii" ~ "California quail", # CA quail 
      
    Including_domestic.Local_species %in% "Alosa sapidissima, Dorosoma petenense" ~ "American shad", # American shad
      
    Including_domestic.Local_species %in% "Leuconotopicus albolarvatus, Jynx torquilla, Melanerpes lewis, Melanerpes erythrocephalus, Melanerpes formicivorus, Melanerpes uropygialis, Sphyrapicus thyroideus, Sphyrapicus varius, Sphyrapicus nuchalis, Sphyrapicus ruber, Picoides arcticus, Dryobates pubescens, Dryobates nuttallii, Dryobates scalaris, Picoides villosus, Colaptes auratus, Colaptes chrysoides, Dryocopus pileatus" ~ "Acorn woodpecker", # woodpecker
      
    Including_domestic.Local_species %in% "Plestiodon gilberti" ~ "Gilbert's skink", # Gilbert's skink
      
    Including_domestic.Local_species %in% "Taxidea taxus" ~ "American badger", # American badger
      
    Including_domestic.Local_species %in% "Microtus townsendii" ~ "Townsend's vole", # Microtus - townsend's vole
      
    Including_domestic.Local_species %in% "Neotoma macrotis" ~ "Big-eared woodrat", # big eared woodrat
    
    Including_domestic.Local_species %in% "Aphelocoma insularis, Aphelocoma californica, Aphelocoma woodhouseii" ~ "California scrub jay", # CA scrub jay
      
    Including_domestic.Local_species %in% "Puma concolor" ~ "Puma", # puma
      
    Including_domestic.Local_species %in% "Meleagris gallopavo, Gallus gallus" ~ "Wild turkey", # wild turkey
      
    Including_domestic.Local_species %in% "Mephitis mephitis" ~ "Striped skunk", # striped skunk
      
    Including_domestic.Local_species %in% "Vulpes vulpes" ~ "Red fox", # red fox
      
    Including_domestic.Local_species %in% "Chaetodipus californicus" ~ "California pocket mouse", # California pocket mouse
      
    Including_domestic.Local_species %in% "Mus musculus" ~ "House mouse", # house mouse
    
    Including_domestic.Local_species %in% "Rattus rattus" ~ "Black rat", # black rat
      
    Including_domestic.Local_species %in% "Peromyscus maniculatus, Peromyscus keeni" ~ "Western deer mouse", # deer mice
      
    Including_domestic.Local_species %in% "Sylvilagus floridanus, Sylvilagus nuttallii" ~ "Brush rabbit", # brush rabbit
      
    Including_domestic.Local_species %in% "Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii" ~ "Vole", # vole
      
    Including_domestic.Local_species %in% "Lemmiscus curtatus, Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii, Ondatra zibethicus, Onychomys leucogaster, Synaptomys borealis, Peromyscus maniculatus, Peromyscus keeni" ~ "Western deer mouse", # cricetidae - some kind of rodent -- choosing microtus (vole)
    
    Including_domestic.Local_species %in% "Neotoma lepida" ~ "Desert woodrat", # desert woodrat
    
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus, Urocyon cinereoargenteus, Urocyon littoralis, Vulpes macrotis, Vulpes vulpes" ~ "Coyote", # coyote
    
        Including_domestic.Local_species %in% "Turdus obscurus, Turdus migratorius" ~ "American robin", #American robin
    
    Including_domestic.Local_species %in% "Spizella passerina, Passerella iliaca, Melospiza melodia" ~ "Song sparrow", #Song sparrow, could  be chipping or fox sparrow too though.
    
    Including_domestic.Local_species %in% "Ascaphus truei, Incilius alvarius, Anaxyrus boreas, Anaxyrus californicus, Anaxyrus canorus, Anaxyrus cognatus, Anaxyrus exsul, Anaxyrus microscaphus, Anaxyrus punctatus, Anaxyrus woodhousii, Pseudacris cadaverina, Pseudacris hypochondriaca, Pseudacris regilla, Pseudacris sierra, Scaphiopus couchii, Spea hammondii, Spea intermontana, Rana aurora, Lithobates berlandieri, Rana boylii, Rana cascadae, Lithobates catesbeianus, Rana draytonii, Rana luteiventris, Rana muscosa, Lithobates pipiens, Rana pretiosa, Rana sierrae, Lithobates sphenocephalus, Lithobates yavapaiensis, Xenopus laevis" ~ "Anura", #frog or toad
    
    Including_domestic.Local_species %in% "Phoca fasciata, Phoca vitulina" ~ "Harbor seal", #harbor seal
    
    Including_domestic.Local_species %in% "Nucifraga columbiana, Corvus brachyrhynchos, Corvus corax" ~ "Corvidae", #corvidae -- Clark's nutcracker, American crow, Common raven
    
        Including_domestic.Local_species %in% "Pipilo maculatus" ~ "Spotted towhee", #spotted towhee
    
    Including_domestic.Local_species %in% "Onychomys torridus" ~ "Southern grasshopper mouse", #southern grasshopper mouse
    
    Including_domestic.Local_species %in% "Arctocephalus townsendi" ~ "Guadelupe fur seal", #guadelupe fur seal
    
    Including_domestic.Local_species %in% "Callospermophilus lateralis" ~ "Golden-mantled ground squirrel", #golden-mantled ground squirrel
    
    Including_domestic.Local_species %in% "Marmota flaviventris" ~ "Yellow-bellied marmot", #yellow-bellied marmot
    
    Including_domestic.Local_species %in% "Microtus montanus" ~ "Montane vole" #montane vole
  ))

# for the 16s run - common names
gaviota <- gaviota %>%
  mutate(common_name = case_when(
    Barcode_region == "16" & Best_match_references %in% "Bos taurus" ~ "Cattle", # 100
    Barcode_region == "16" & Best_match_references %in% "Lynx rufus" ~ "Bobcat", # 100
    #Best_match_references %in% "Heliconoides inflatus" ~ "Heliconoides inflatus", # sea butterfly snail
    #Best_match_references %in% "Odocoileus virginianus" ~ "white tailed deer", # 100
    #Best_match_references %in% "Crotalus viridis" ~ "prairie rattlesnake", # 100
    Barcode_region == "16" & Best_match_references %in% "Puma concolor" ~ "Puma", # puma
    Barcode_region == "16" & Best_match_references %in% "Rattus rattus" ~ "Black rat", # rat
    #Best_match_references %in% "Gallus sonneratii" ~ "chicken", # 100
    Barcode_region == "16" & Best_match_references %in% "Callipepla californica" ~ "California quail", # California quail
    #Best_match_references %in% "Scomber japonicus" ~ "chub mackerel", # 100
    Barcode_region == "16" & Best_match_references %in% "Lamna ditropis" ~ "Salmon shark", #salmon shark
    Barcode_region == "16" & Best_match_references %in% "Magnisudis atlantica" ~ "Duckbill baracuda", # duckbill barracuda
    Barcode_region == "16" & Best_match_references %in% "Ophiodon elongatus" ~ "Lingcod", # lingcod
    Barcode_region == "16" & Best_match_references %in% "Thysanoessa inermis" ~ "Krill", # Krill
    #Best_match_references %in% "Sebastes serranoides" ~ "Sebastes sp", # olive rockfish
    Barcode_region == "16" & Best_match_references %in% "Clupea pallasii" ~ "Pacific herring", # pacific herring
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa" ~ "Wild boar", #wild boar
    #Best_match_references %in% "Canis rufus" ~ "canid", #
    Barcode_region == "16" & Best_match_references %in% "Pilaira anomala" ~ "Mold", #mold
    Barcode_region == "16" & Best_match_references %in% "Reithrodontomys megalotis" ~ "Western harvest mouse", #western harvest mouse
    Barcode_region == "16" & Best_match_references %in% "Microtus californicus" ~ "California vole", #California vole
    Barcode_region == "16" & Best_match_references %in% "Chaetodipus californicus" ~ "California pocket mouse", #California pocket mouse
    Barcode_region == "16" & Best_match_references %in% "Zalophus californianus" ~ "California sea lion", #California sea lion
    Barcode_region == "16" & Best_match_references %in% "Canis latrans" ~ "Coyote", #coyote
    Barcode_region == "16" & Best_match_references %in% "Neotoma lepida" ~ "Desert woodrat", #desert woodrat
    #Best_match_references %in% "Meleagris gallopavo" ~ "turkey", #
    Barcode_region == "16" & Best_match_references %in% "Otospermophilus beecheyi" ~ "California ground squirrel", #California ground squirrel
    Barcode_region == "16" & Best_match_references %in% "Taxidea taxus" ~ "American badger", #american badger
    Barcode_region == "16" & Best_match_references %in% "Neotoma macrotis" ~ "Big-eared woodrat", #big-eared woodrat
    #Best_match_references %in% "Ravinia vagabunda" ~ "flesh fly", #
    Barcode_region == "16" & Best_match_references %in% "Canis lupus" ~ "Coyote", # canid
    Barcode_region == "16" & Best_match_references %in% "Forficula auricularia" ~ "European earwig", #European earwig
    Barcode_region == "16" & Best_match_references %in% "Mephitis mephitis" ~ "Striped skunk", #striped skunk
    #Best_match_references %in% "Paralucilia pseudolyrcea" ~ "blow fly", #
    Barcode_region == "16" & Best_match_references %in% "Urocyon cinereoargenteus" ~ "Grey fox", #grey fox
    Barcode_region == "16" & Best_match_references %in% "Pandalus jordani" ~ "Ocean pink shrimp", #ocean pink shrimp
    #Best_match_references %in% "Merluccius productus" ~ "north pacific hake", #
    Barcode_region == "16" & Best_match_references %in% "Peromyscus maniculatus" ~ "Western deer mouse", #deer mouse
    Barcode_region == "16" & Best_match_references %in% "Pituophis catenifer" ~ "Gopher snake", #gopher snake
    Barcode_region == "16" & Best_match_references %in% "Sylvilagus bachmani" ~ "Brush rabbit", #brush rabbit
    Barcode_region == "16" & Best_match_references %in% "Plestiodon skiltonianus" ~ "Western skink", #western skink
    Barcode_region == "16" & Best_match_references %in% "Anchoa ischana" ~ "Sharpnose anchovy", #sharpnose anchovy
    Barcode_region == "16" & Best_match_references %in% "Gonatopsis borealis" ~ "Boreopacific armhook squid", #Boreopacific armhook squid
    Barcode_region == "16" & Best_match_references %in% "Leucocarbo chalconotus" ~ "Pelagic cormorant", # cormorant species, checked with BLAST. shag (stewart island shag, but those are just in NZ)
    Barcode_region == "16" & Best_match_references %in% "Pseudopaludicola boliviana" ~ "Anura", # frog, Leptodactylidae family, found in regions in south america???
    Barcode_region == "16" & Best_match_references %in% "Microtus chrotorrhinus" ~ "Vole", # rock vole, found in eastern US
    Barcode_region == "16" & Best_match_references %in% "Picoides tridactylus" ~ "Downy woodpecker", # eurasian three-toed woodpecker, not found in US --- changed to Dryobates pubescens, downy woodpecker
    Barcode_region == "16" & Best_match_references %in% "Thomomys bottae" ~ "Botta's pocket gopher", # Botta's pocket gopher
    Barcode_region == "16" & Best_match_references %in% "Neodon bershulaensis" ~ "Vole", # Bershula mountain vole, found in China -- BLAST shows microtus species second match
    #Best_match_references %in% "Chloropicon sieburthii" ~ "phytoplankton", # 90
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus" ~ "Hairy rove beetle", # hairy rove beetle -- only shows up once
    Barcode_region == "16" & Best_match_references %in% "Rhipicephalus sanguineus" ~ "Brown dog tick", # Brown dog tick
    Barcode_region == "16" & Best_match_references %in% "Phakopsora meibomiae" ~ "fungi", # fungi
    Barcode_region == "16" & Best_match_references %in% "Tettigonia orientalis" ~ "Cricket", # bush cricket, but this species is found in Japan --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Isogenoides varians" ~ "Stonefly", # stonefly
    Barcode_region == "16" & Best_match_references %in% "Setvena wahkeena" ~ "Cricket", # stonefly (found in socal?) -- BLAST - cricket match
    Barcode_region == "16" & Best_match_references %in% "Xiphidiopsis gurneyi" ~ "Cricket", #bush cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Abronia graminea" ~ "Southern alligator lizard", # alligator lizard (found in socal?) -- BLAST - southern aligator lizard
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini" ~ "Cricket", # camel cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Chloropicon laureae" ~ "Green algae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae" ~ "Cricket", #Scarabaeidae (scarab) -- changed to species that came up with this species in results -- BLAST- cricket
    Barcode_region == "16" & Best_match_references %in% "Stenopelmatus fuscus" ~ "Jerusalem cricket", # 80
    Barcode_region == "16" & Best_match_references %in% "Gymnaetoides acutilobatus" ~ "Cricket", # endemic to china??? --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Microdipodops pallidus" ~ "Pale kangaroo mouse", # pale kangaroo mouse
    Barcode_region == "16" & Best_match_references %in% "Onthophagus crinitus" ~ "Dung beetle", # dung beetle
    Barcode_region == "16" & Best_match_references %in% "Pheidole deima" ~ "Harvester ant", # ant (found in socal?) --- BLAST - Veromessor andrei - harvester ant
    Barcode_region == "16" & Best_match_references %in% "Hazenia capsulata" ~ "Green algae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Anadara broughtonii" ~ "Inflated ark clam", # Inflated ark clam
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea" ~ "mold", # slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis" ~ "green algae", # 70

    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos javanicus, Bos indicus, Bos grunniens, Bos grunniens x Bos taurus" ~ "Cattle",
    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos primigenius, Bos indicus" ~ "Cattle", #cow
    Barcode_region == "16" & Best_match_references %in% "Canis rufus, Canis lupus, Canis latrans" ~ "Coyote", # coyote
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea, Cavenderia pseudoaureostipes" ~ "mold", #slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "green algae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Chloropicon sieburthii, Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "green algae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus, Anthrenus scrophulariae" ~ "Common carpet beetle", # Common carpet beetle/Buffalo carpet beetle
    Barcode_region == "16" & Best_match_references %in% "Crotalus viridis, Crotalus oreganus" ~ "Western rattlesnake", #western rattlesnake
    Barcode_region == "16" & Best_match_references %in% "Gallus sonneratii, Gallus gallus" ~ "Chicken", # wild turkey probably
    Barcode_region == "16" & Best_match_references %in% "Heliconoides inflatus, Cathaica polystigma, Nesiohelix moreletiana, Karaftohelix maackii, Atlanta turriculata, Aceria guerreronis, Hahnia ngai, Grylloprimevala jilina, Aconurella prolixa, Processina dashahensis, Tapinoma indicum, Homo sapiens" ~ "Sea butterfly snail", # Heliconoides inflatus -- Sea butterfly (range is not Socal, but Mexico, and it's the closest species in the list) 
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae, Paratriodonta lentula" ~ "Scarab beetle", #Scarabaeidae (scarab)
    Barcode_region == "16" & Best_match_references %in% "Meleagris gallopavo, Labidesthes sicculus" ~ "Wild turkey", #wild turkey
    Barcode_region == "16" & Best_match_references %in% "Merluccius productus, Merluccius angustimanus" ~ "North Pacific Hake", #North Pacific Hake
    Barcode_region == "16" & Best_match_references %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Mule deer", #mule deer
    Barcode_region == "16" & Best_match_references %in% "Paralucilia pseudolyrcea, Paralucilia paraensis, Hemilucilia semidiaphana, Calliphora vomitoria, Calliphora vicina, Calliphora sinensis, Calliphora nigribarbis, Calliphora calliphoroides" ~ "Fly", # fly
    Barcode_region == "16" & Best_match_references %in% "Ravinia vagabunda, Ravinia stimulans, Ravinia derelicta" ~ "Flesh fly", #flesh fly
    Barcode_region == "16" & Best_match_references %in% "Scomber japonicus, Scomber colias, Scomber australasicus"~ "Pacific makerel", #pacific mackerel
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa, Thryonomys swinderianus" ~ "Wild boar", #wild pig
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini, Onthophagus obscurior" ~ "dung beetle", #dung beetle
    Barcode_region == "16" & Best_match_references %in% "Sebastes serranoides, Sebastes melanops, Sebastes flavidus" ~ "Rockfish",  #rockfish
    Barcode_region == "16" & Best_match_references %in% "" ~ "no match",
    TRUE ~ common_name
    ))


checkna <- gaviota[is.na(gaviota$common_name), ] 
#rm(checkna)
```

## Curate broad tax

```{r}
gaviota$broadtax = NA

gaviota <- gaviota %>%
  mutate(broadtax = case_when(
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus" ~ "Coyote", # coyote
    
    Including_domestic.Local_species %in% "Phalacrocorax penicillatus, Phalacrocorax brasilianus, Phalacrocorax pelagicus" 
      ~ "Brandt's cormorant", # Brant's cormorant
    
    Including_domestic.Local_species %in% "Phalacrocorax pelagicus, Phalacrocorax penicillatus" 
      ~ "Brandt's cormorant", # Brant's cormorant
    
    Including_domestic.Local_species %in% "Microtus californicus" 
      ~ "Small rodent", # Californian vole
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei" ~ "Small rodent", #Peromyscus - deer mouse
    
    Including_domestic.Local_species %in% "Sylvilagus bachmani" ~ "Brush rabbit", # Brush rabbit
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Arborimus albipes, Arborimus longicaudus, Arborimus pomo, Myodes californicus, Lemmiscus curtatus, Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii, Ondatra zibethicus, Phenacomys intermedius, Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis, Onychomys leucogaster, Onychomys torridus, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei, Reithrodontomys raviventris, Reithrodontomys megalotis, Sigmodon arizonae, Sigmodon hispidus" ~ "Small rodent", # 	Cricetidae (mouse or vole)
      
    Including_domestic.Local_species %in% "Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis" ~ "Small rodent", # Neotoma - woodrat
      
    Including_domestic.Local_species %in% "Bos taurus, Bos indicus, Bos javanicus, Bos frontalis" ~ "Cattle", # Bovine
      
    Including_domestic.Local_species %in% "Gallus gallus" ~ "Terrestrial bird", #chicken
      
    Including_domestic.Local_species %in% "Urocyon cinereoargenteus, Urocyon littoralis" ~ "Grey fox", # grey fox
      
    Including_domestic.Local_species %in% "Glaucomys oregonensis, Glaucomys sabrinus, Callospermophilus lateralis" ~ "Small rodent", # flying squirrel/golden mantle ground squirrel
      
    Including_domestic.Local_species %in% "Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii" ~ "Small rodent", # vole
    
    Including_domestic.Local_species %in% "Thomomys bottae" ~"Small rodent", # botta's pocket gopher
    
    Including_domestic.Local_species %in% "Lynx rufus, Puma concolor" ~ "Bobcat", # bobcat
      
    Including_domestic.Local_species %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Mule deer", # mule deer
      
    Including_domestic.Local_species %in% "Sus scrofa" ~ "Wild boar", #wild boar
      
    Including_domestic.Local_species %in% "Chloris sinica, Spinus pinus, Plectrophenax nivalis, Zonotrichia leucophrys, Zonotrichia atricapilla, Zonotrichia albicollis, Sturnella neglecta, Sturnella magna, Serinus canaria" ~ "Terrestrial bird", # Passeriformes
      
    Including_domestic.Local_species %in% "Peromyscus eremicus" ~ "Small rodent", # cactus mouse
      
    Including_domestic.Local_species %in% "Zalophus californianus" ~ "California sea lion", # sea lion
    
    Including_domestic.Local_species %in% "Canis lupus" ~ "Coyote", # coyote
      
    Including_domestic.Local_species %in% "Peromyscus crinitus, Peromyscus eremicus" ~ "Small rodent", # Peromyscus
      
    Including_domestic.Local_species %in% "Callipepla californica, Callipepla gambelii" ~ "Terrestrial bird", # CA quail 
      
    Including_domestic.Local_species %in% "Alosa sapidissima, Dorosoma petenense" ~ "American shad", # American shad
      
    Including_domestic.Local_species %in% "Leuconotopicus albolarvatus, Jynx torquilla, Melanerpes lewis, Melanerpes erythrocephalus, Melanerpes formicivorus, Melanerpes uropygialis, Sphyrapicus thyroideus, Sphyrapicus varius, Sphyrapicus nuchalis, Sphyrapicus ruber, Picoides arcticus, Dryobates pubescens, Dryobates nuttallii, Dryobates scalaris, Picoides villosus, Colaptes auratus, Colaptes chrysoides, Dryocopus pileatus" ~ "Terrestrial bird", # woodpecker
      
    Including_domestic.Local_species %in% "Plestiodon gilberti" ~ "Gilbert's skink", # Gilbert's skink
      
    Including_domestic.Local_species %in% "Taxidea taxus" ~ "American badger", # American badger
      
    Including_domestic.Local_species %in% "Microtus townsendii" ~ "Small rodent", # Microtus - townsend's vole
      
    Including_domestic.Local_species %in% "Neotoma macrotis" ~ "Small rodent", # big eared woodrat
    
    Including_domestic.Local_species %in% "Aphelocoma insularis, Aphelocoma californica, Aphelocoma woodhouseii" ~ "Terrestrial bird", # CA scrub jay
      
    Including_domestic.Local_species %in% "Puma concolor" ~ "Puma", # puma
      
    Including_domestic.Local_species %in% "Meleagris gallopavo, Gallus gallus" ~ "Terrestrial bird", # wild turkey
      
    Including_domestic.Local_species %in% "Mephitis mephitis" ~ "Striped skunk", # striped skunk
      
    Including_domestic.Local_species %in% "Vulpes vulpes" ~ "Red fox", # red fox
      
    Including_domestic.Local_species %in% "Chaetodipus californicus" ~ "Small rodent", # California pocket mouse
      
    Including_domestic.Local_species %in% "Mus musculus" ~ "Small rodent", # house mouse
    
    Including_domestic.Local_species %in% "Rattus rattus" ~ "Small rodent", # black rat
      
    Including_domestic.Local_species %in% "Peromyscus maniculatus, Peromyscus keeni" ~ "Small rodent", # deer mice
      
    Including_domestic.Local_species %in% "Sylvilagus floridanus, Sylvilagus nuttallii" ~ "Brush rabbit", # brush rabbit
      
    Including_domestic.Local_species %in% "Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii" ~ "Small rodent", # vole
      
    Including_domestic.Local_species %in% "Lemmiscus curtatus, Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii, Ondatra zibethicus, Onychomys leucogaster, Synaptomys borealis, Peromyscus maniculatus, Peromyscus keeni" ~ "Small rodent", # cricetidae - some kind of rodent
    
    Including_domestic.Local_species %in% "Neotoma lepida" ~ "Small rodent", # desert woodrat
    
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus, Urocyon cinereoargenteus, Urocyon littoralis, Vulpes macrotis, Vulpes vulpes" ~ "Coyote", # coyote
        
        Including_domestic.Local_species %in% "Turdus obscurus, Turdus migratorius" ~ "Terrestrial bird", #American robin
    
    Including_domestic.Local_species %in% "Spizella passerina, Passerella iliaca, Melospiza melodia" ~ "Terrestrial bird", #Song sparrow, could  be chipping or fox sparrow too though.
    
    Including_domestic.Local_species %in% "Ascaphus truei, Incilius alvarius, Anaxyrus boreas, Anaxyrus californicus, Anaxyrus canorus, Anaxyrus cognatus, Anaxyrus exsul, Anaxyrus microscaphus, Anaxyrus punctatus, Anaxyrus woodhousii, Pseudacris cadaverina, Pseudacris hypochondriaca, Pseudacris regilla, Pseudacris sierra, Scaphiopus couchii, Spea hammondii, Spea intermontana, Rana aurora, Lithobates berlandieri, Rana boylii, Rana cascadae, Lithobates catesbeianus, Rana draytonii, Rana luteiventris, Rana muscosa, Lithobates pipiens, Rana pretiosa, Rana sierrae, Lithobates sphenocephalus, Lithobates yavapaiensis, Xenopus laevis" ~ "Frog", #frog or toad
    
    Including_domestic.Local_species %in% "Phoca fasciata, Phoca vitulina" ~ "Harbor seal", #harbor seal
    
    Including_domestic.Local_species %in% "Nucifraga columbiana, Corvus brachyrhynchos, Corvus corax" ~ "Terrestrial bird", #corvidae -- Clark's nutcracker, American crow, Common raven
    
        Including_domestic.Local_species %in% "Pipilo maculatus" ~ "Terrestrial bird", #spotted towhee
    
    Including_domestic.Local_species %in% "Onychomys torridus" ~ "Small rodent", #southern grasshopper mouse
    
    Including_domestic.Local_species %in% "Arctocephalus townsendi" ~ "Guadelupe fur seal", #guadelupe fur seal
    
    Including_domestic.Local_species %in% "Callospermophilus lateralis" ~ "Small rodent", #golden-mantled ground squirrel
    
    Including_domestic.Local_species %in% "Marmota flaviventris" ~ "Small rodent", #yellow-bellied marmot
    
    Including_domestic.Local_species %in% "Microtus montanus" ~ "Small rodent" #montane vole
  ))

# for the 16s run - broad taxonomy
gaviota <- gaviota %>%
  mutate(broadtax = case_when(
    Barcode_region == "16" & Best_match_references %in% "Bos taurus" ~ "Cattle", # 100
    Barcode_region == "16" & Best_match_references %in% "Lynx rufus" ~ "Bobcat", # 100
    #Best_match_references %in% "Heliconoides inflatus" ~ "Heliconoides inflatus", # sea butterfly snail
    #Best_match_references %in% "Odocoileus virginianus" ~ "white tailed deer", # 100
    #Best_match_references %in% "Crotalus viridis" ~ "prairie rattlesnake", # 100
    Barcode_region == "16" & Best_match_references %in% "Puma concolor" ~ "Puma", # puma
    Barcode_region == "16" & Best_match_references %in% "Rattus rattus" ~ "Small rodent", # rat
    #Best_match_references %in% "Gallus sonneratii" ~ "chicken", # 100
    Barcode_region == "16" & Best_match_references %in% "Callipepla californica" ~ "Terrestrial bird", # California quail
    #Best_match_references %in% "Scomber japonicus" ~ "chub mackerel", # 100
    Barcode_region == "16" & Best_match_references %in% "Lamna ditropis" ~ "Salmon shark", #salmon shark
    Barcode_region == "16" & Best_match_references %in% "Magnisudis atlantica" ~ "Duckbill baracuda", # duckbill barracuda
    Barcode_region == "16" & Best_match_references %in% "Ophiodon elongatus" ~ "Lingcod", # lingcod
    Barcode_region == "16" & Best_match_references %in% "Thysanoessa inermis" ~ "Krill", # Krill
    #Best_match_references %in% "Sebastes serranoides" ~ "Sebastes sp", # olive rockfish
    Barcode_region == "16" & Best_match_references %in% "Clupea pallasii" ~ "Pacific herring", # pacific herring
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa" ~ "Wild boar", #wild boar
    #Best_match_references %in% "Canis rufus" ~ "canid", #
    Barcode_region == "16" & Best_match_references %in% "Pilaira anomala" ~ "Mold", #mold
    Barcode_region == "16" & Best_match_references %in% "Reithrodontomys megalotis" ~ "Small rodent", #western harvest mouse
    Barcode_region == "16" & Best_match_references %in% "Microtus californicus" ~ "Small rodent", #California vole
    Barcode_region == "16" & Best_match_references %in% "Chaetodipus californicus" ~ "Small rodent", #California pocket mouse
    Barcode_region == "16" & Best_match_references %in% "Zalophus californianus" ~ "California sea lion", #California sea lion
    Barcode_region == "16" & Best_match_references %in% "Canis latrans" ~ "Coyote", #coyote
    Barcode_region == "16" & Best_match_references %in% "Neotoma lepida" ~ "Small rodent", #desert woodrat
    #Best_match_references %in% "Meleagris gallopavo" ~ "turkey", #
    Barcode_region == "16" & Best_match_references %in% "Otospermophilus beecheyi" ~ "Small rodent", #California ground squirrel
    Barcode_region == "16" & Best_match_references %in% "Taxidea taxus" ~ "Other terrestrial mammal", #american badger
    Barcode_region == "16" & Best_match_references %in% "Neotoma macrotis" ~ "Small rodent", #big-eared woodrat
    #Best_match_references %in% "Ravinia vagabunda" ~ "flesh fly", #
    Barcode_region == "16" & Best_match_references %in% "Canis lupus" ~ "Coyote", # canid
    Barcode_region == "16" & Best_match_references %in% "Forficula auricularia" ~ "Terrestrial insect", #European earwig
    Barcode_region == "16" & Best_match_references %in% "Mephitis mephitis" ~ "Striped skunk", #striped skunk
    #Best_match_references %in% "Paralucilia pseudolyrcea" ~ "blow fly", #
    Barcode_region == "16" & Best_match_references %in% "Urocyon cinereoargenteus" ~ "Grey fox", #grey fox
    Barcode_region == "16" & Best_match_references %in% "Pandalus jordani" ~ "Ocean pink shrimp", #ocean pink shrimp
    #Best_match_references %in% "Merluccius productus" ~ "north pacific hake", #
    Barcode_region == "16" & Best_match_references %in% "Peromyscus maniculatus" ~ "Small rodent", #deer mouse
    Barcode_region == "16" & Best_match_references %in% "Pituophis catenifer" ~ "Reptile", #gopher snake
    Barcode_region == "16" & Best_match_references %in% "Sylvilagus bachmani" ~ "Brush rabbit", #brush rabbit
    Barcode_region == "16" & Best_match_references %in% "Plestiodon skiltonianus" ~ "Reptile", #western skink
    Barcode_region == "16" & Best_match_references %in% "Anchoa ischana" ~ "Sharpnose anchovy", #sharpnose anchovy
    Barcode_region == "16" & Best_match_references %in% "Gonatopsis borealis" ~ "Boreopacific armhook squid", #Boreopacific armhook squid
    Barcode_region == "16" & Best_match_references %in% "Leucocarbo chalconotus" ~ "Pelagic cormorant", # cormorant species, checked with BLAST. shag (stewart island shag, but those are just in NZ)
    Barcode_region == "16" & Best_match_references %in% "Pseudopaludicola boliviana" ~ "Anura", # frog, Leptodactylidae family, found in regions in south america???
    Barcode_region == "16" & Best_match_references %in% "Microtus chrotorrhinus" ~ "Small rodent", # rock vole, found in eastern US
    Barcode_region == "16" & Best_match_references %in% "Picoides tridactylus" ~ "Terrestrial bird", # eurasian three-toed woodpecker, not found in US --- changed to Dryobates pubescens, downy woodpecker
    Barcode_region == "16" & Best_match_references %in% "Thomomys bottae" ~ "Small rodent", # Botta's pocket gopher
    Barcode_region == "16" & Best_match_references %in% "Neodon bershulaensis" ~ "Small rodent", # Bershula mountain vole, found in China -- BLAST shows microtus species second match
    #Best_match_references %in% "Chloropicon sieburthii" ~ "phytoplankton", # 90
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus" ~ "Terrestrial insect", # hairy rove beetle -- only shows up once
    Barcode_region == "16" & Best_match_references %in% "Rhipicephalus sanguineus" ~ "Brown dog tick", # Brown dog tick
    Barcode_region == "16" & Best_match_references %in% "Phakopsora meibomiae" ~ "Fungi", # fungi
    Barcode_region == "16" & Best_match_references %in% "Tettigonia orientalis" ~ "Terrestrial insect", # bush cricket, but this species is found in Japan --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Isogenoides varians" ~ "Terrestrial insect", # stonefly
    Barcode_region == "16" & Best_match_references %in% "Setvena wahkeena" ~ "Terrestrial insect", # stonefly (found in socal?) -- BLAST - cricket match
    Barcode_region == "16" & Best_match_references %in% "Xiphidiopsis gurneyi" ~ "Terrestrial insect", #bush cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Abronia graminea" ~ "Reptile", # alligator lizard (found in socal?) -- BLAST - southern aligator lizard
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini" ~ "Terrestrial insect", # camel cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Chloropicon laureae" ~ "Green algae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae" ~ "Terrestrial insect", #Scarabaeidae (scarab) -- changed to species that came up with this species in results -- BLAST- cricket
    Barcode_region == "16" & Best_match_references %in% "Stenopelmatus fuscus" ~ "Terrestrial insect", # 80
    Barcode_region == "16" & Best_match_references %in% "Gymnaetoides acutilobatus" ~ "Small rodent", # endemic to china??? --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Microdipodops pallidus" ~ "Terrestrial insect", # pale kangaroo mouse
    Barcode_region == "16" & Best_match_references %in% "Onthophagus crinitus" ~ "Terrestrial insect", # dung beetle
    Barcode_region == "16" & Best_match_references %in% "Pheidole deima" ~ "Terrestrial insect", # ant (found in socal?) --- BLAST - Veromessor andrei - harvester ant
    Barcode_region == "16" & Best_match_references %in% "Hazenia capsulata" ~ "Green algae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Anadara broughtonii" ~ "Clam", # Inflated ark clam
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea" ~ "mold", # slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis" ~ "Green algae", # 70

    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos javanicus, Bos indicus, Bos grunniens, Bos grunniens x Bos taurus" ~ "Cattle",
    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos primigenius, Bos indicus" ~ "Cattle", #cow
    Barcode_region == "16" & Best_match_references %in% "Canis rufus, Canis lupus, Canis latrans" ~ "Coyote", # coyote
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea, Cavenderia pseudoaureostipes" ~ "mold", #slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Green algae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Chloropicon sieburthii, Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Green algae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus, Anthrenus scrophulariae" ~ "Terrestrial insect", # Common carpet beetle/Buffalo carpet beetle
    Barcode_region == "16" & Best_match_references %in% "Crotalus viridis, Crotalus oreganus" ~ "Reptile", #western rattlesnake
    Barcode_region == "16" & Best_match_references %in% "Gallus sonneratii, Gallus gallus" ~ "Terrestrial bird", # wild turkey probably
    Barcode_region == "16" & Best_match_references %in% "Heliconoides inflatus, Cathaica polystigma, Nesiohelix moreletiana, Karaftohelix maackii, Atlanta turriculata, Aceria guerreronis, Hahnia ngai, Grylloprimevala jilina, Aconurella prolixa, Processina dashahensis, Tapinoma indicum, Homo sapiens" ~ "Sea butterfly snail", # Heliconoides inflatus -- Sea butterfly (range is not Socal, but Mexico, and it's the closest species in the list) 
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae, Paratriodonta lentula" ~ "Terrestrial insect", #Scarabaeidae (scarab)
    Barcode_region == "16" & Best_match_references %in% "Meleagris gallopavo, Labidesthes sicculus" ~ "Terrestrial bird", #wild turkey
    Barcode_region == "16" & Best_match_references %in% "Merluccius productus, Merluccius angustimanus" ~ "North Pacific Hake", #North Pacific Hake
    Barcode_region == "16" & Best_match_references %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Mule deer", #mule deer
    Barcode_region == "16" & Best_match_references %in% "Paralucilia pseudolyrcea, Paralucilia paraensis, Hemilucilia semidiaphana, Calliphora vomitoria, Calliphora vicina, Calliphora sinensis, Calliphora nigribarbis, Calliphora calliphoroides" ~ "Terrestrial insect", # fly
    Barcode_region == "16" & Best_match_references %in% "Ravinia vagabunda, Ravinia stimulans, Ravinia derelicta" ~ "Terrestrial insect", #flesh fly
    Barcode_region == "16" & Best_match_references %in% "Scomber japonicus, Scomber colias, Scomber australasicus"~ "Pacific makerel", #pacific mackerel
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa, Thryonomys swinderianus" ~ "Wild boar", #wild pig
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini, Onthophagus obscurior" ~ "Terrestrial insect", #dung beetle
    Barcode_region == "16" & Best_match_references %in% "Sebastes serranoides, Sebastes melanops, Sebastes flavidus" ~ "Rockfish",  #rockfish
    Barcode_region == "16" & Best_match_references %in% "" ~ "no match",
    TRUE ~ broadtax
    ))



checkna <- gaviota[is.na(gaviota$broadtax), ] 
rm(checkna)
```

## Curate broader taxonomic class
grouping categories:
- coyote
- cattle
- mule deer
- reptiles/amphibians
- small rodents
- brandt's cormorant
- terrestrial bird
- wild boar
- other terrestrial mammal
- california sea lion
- puma
- American shad


```{r}
gaviota$broadtax1 = NA

gaviota <- gaviota %>%
  mutate(broadtax1 = case_when(
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus" ~ "Coyote", # coyote
    
    Including_domestic.Local_species %in% "Phalacrocorax penicillatus, Phalacrocorax brasilianus, Phalacrocorax pelagicus" 
      ~ "Brandt's cormorant", # Brant's cormorant
    
    Including_domestic.Local_species %in% "Phalacrocorax pelagicus, Phalacrocorax penicillatus" 
      ~ "Brandt's cormorant", # Brant's cormorant
    
    Including_domestic.Local_species %in% "Microtus californicus" 
      ~ "Small rodent", # Californian vole
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei" ~ "Small rodent", #Peromyscus - deer mouse
    
    Including_domestic.Local_species %in% "Sylvilagus bachmani" ~ "Brush rabbit", # Brush rabbit
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Arborimus albipes, Arborimus longicaudus, Arborimus pomo, Myodes californicus, Lemmiscus curtatus, Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii, Ondatra zibethicus, Phenacomys intermedius, Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis, Onychomys leucogaster, Onychomys torridus, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei, Reithrodontomys raviventris, Reithrodontomys megalotis, Sigmodon arizonae, Sigmodon hispidus" ~ "Small rodent", # 	Cricetidae (mouse or vole)
      
    Including_domestic.Local_species %in% "Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis" ~ "Small rodent", # Neotoma - woodrat
      
    Including_domestic.Local_species %in% "Bos taurus, Bos indicus, Bos javanicus, Bos frontalis" ~ "Cattle", # Bovine
      
    Including_domestic.Local_species %in% "Gallus gallus" ~ "Terrestrial bird", #chicken
      
    Including_domestic.Local_species %in% "Urocyon cinereoargenteus, Urocyon littoralis" ~ "Grey fox", # grey fox
      
    Including_domestic.Local_species %in% "Glaucomys oregonensis, Glaucomys sabrinus, Callospermophilus lateralis" ~ "Small rodent", # flying squirrel/golden mantle ground squirrel
      
    Including_domestic.Local_species %in% "Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii" ~ "Small rodent", # vole
    
    Including_domestic.Local_species %in% "Thomomys bottae" ~"Small rodent", # botta's pocket gopher
    
    Including_domestic.Local_species %in% "Lynx rufus, Puma concolor" ~ "Bobcat", # bobcat
      
    Including_domestic.Local_species %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Mule deer", # mule deer
      
    Including_domestic.Local_species %in% "Sus scrofa" ~ "Wild boar", #wild boar
      
    Including_domestic.Local_species %in% "Chloris sinica, Spinus pinus, Plectrophenax nivalis, Zonotrichia leucophrys, Zonotrichia atricapilla, Zonotrichia albicollis, Sturnella neglecta, Sturnella magna, Serinus canaria" ~ "Terrestrial bird", # Passeriformes
      
    Including_domestic.Local_species %in% "Peromyscus eremicus" ~ "Small rodent", # cactus mouse
      
    Including_domestic.Local_species %in% "Zalophus californianus" ~ "California sea lion", # sea lion
    
    Including_domestic.Local_species %in% "Canis lupus" ~ "Coyote", # coyote
      
    Including_domestic.Local_species %in% "Peromyscus crinitus, Peromyscus eremicus" ~ "Small rodent", # Peromyscus
      
    Including_domestic.Local_species %in% "Callipepla californica, Callipepla gambelii" ~ "Terrestrial bird", # CA quail 
      
    Including_domestic.Local_species %in% "Alosa sapidissima, Dorosoma petenense" ~ "American shad", # American shad
      
    Including_domestic.Local_species %in% "Leuconotopicus albolarvatus, Jynx torquilla, Melanerpes lewis, Melanerpes erythrocephalus, Melanerpes formicivorus, Melanerpes uropygialis, Sphyrapicus thyroideus, Sphyrapicus varius, Sphyrapicus nuchalis, Sphyrapicus ruber, Picoides arcticus, Dryobates pubescens, Dryobates nuttallii, Dryobates scalaris, Picoides villosus, Colaptes auratus, Colaptes chrysoides, Dryocopus pileatus" ~ "Terrestrial bird", # woodpecker
      
    Including_domestic.Local_species %in% "Plestiodon gilberti" ~ "Reptile/amphibian", # Gilbert's skink
      
    Including_domestic.Local_species %in% "Taxidea taxus" ~ "Other terrestrial mammal", # American badger
      
    Including_domestic.Local_species %in% "Microtus townsendii" ~ "Small rodent", # Microtus - townsend's vole
      
    Including_domestic.Local_species %in% "Neotoma macrotis" ~ "Small rodent", # big eared woodrat
    
    Including_domestic.Local_species %in% "Aphelocoma insularis, Aphelocoma californica, Aphelocoma woodhouseii" ~ "Terrestrial bird", # CA scrub jay
      
    Including_domestic.Local_species %in% "Puma concolor" ~ "Puma", # puma
      
    Including_domestic.Local_species %in% "Meleagris gallopavo, Gallus gallus" ~ "Terrestrial bird", # wild turkey
      
    Including_domestic.Local_species %in% "Mephitis mephitis" ~ "other terrestrial mammal", # striped skunk
      
    Including_domestic.Local_species %in% "Vulpes vulpes" ~ "Red fox", # red fox
      
    Including_domestic.Local_species %in% "Chaetodipus californicus" ~ "Small rodent", # California pocket mouse
      
    Including_domestic.Local_species %in% "Mus musculus" ~ "Small rodent", # house mouse
    
    Including_domestic.Local_species %in% "Rattus rattus" ~ "Small rodent", # black rat
      
    Including_domestic.Local_species %in% "Peromyscus maniculatus, Peromyscus keeni" ~ "Small rodent", # deer mice
      
    Including_domestic.Local_species %in% "Sylvilagus floridanus, Sylvilagus nuttallii" ~ "Brush rabbit", # brush rabbit
      
    Including_domestic.Local_species %in% "Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii" ~ "Small rodent", # vole
      
    Including_domestic.Local_species %in% "Lemmiscus curtatus, Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii, Ondatra zibethicus, Onychomys leucogaster, Synaptomys borealis, Peromyscus maniculatus, Peromyscus keeni" ~ "Small rodent", # cricetidae - some kind of rodent
    
    Including_domestic.Local_species %in% "Neotoma lepida" ~ "Small rodent", # desert woodrat
    
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus, Urocyon cinereoargenteus, Urocyon littoralis, Vulpes macrotis, Vulpes vulpes" ~ "Coyote", # coyote
    
            
        Including_domestic.Local_species %in% "Turdus obscurus, Turdus migratorius" ~ "Terrestrial bird", #American robin
    
    Including_domestic.Local_species %in% "Spizella passerina, Passerella iliaca, Melospiza melodia" ~ "Terrestrial bird", #Song sparrow, could  be chipping or fox sparrow too though.
    
    Including_domestic.Local_species %in% "Ascaphus truei, Incilius alvarius, Anaxyrus boreas, Anaxyrus californicus, Anaxyrus canorus, Anaxyrus cognatus, Anaxyrus exsul, Anaxyrus microscaphus, Anaxyrus punctatus, Anaxyrus woodhousii, Pseudacris cadaverina, Pseudacris hypochondriaca, Pseudacris regilla, Pseudacris sierra, Scaphiopus couchii, Spea hammondii, Spea intermontana, Rana aurora, Lithobates berlandieri, Rana boylii, Rana cascadae, Lithobates catesbeianus, Rana draytonii, Rana luteiventris, Rana muscosa, Lithobates pipiens, Rana pretiosa, Rana sierrae, Lithobates sphenocephalus, Lithobates yavapaiensis, Xenopus laevis" ~ "reptile/amphibian", #frog or toad
    
    Including_domestic.Local_species %in% "Phoca fasciata, Phoca vitulina" ~ "Harbor seal", #harbor seal
    
    Including_domestic.Local_species %in% "Nucifraga columbiana, Corvus brachyrhynchos, Corvus corax" ~ "Terrestrial bird", #corvidae -- Clark's nutcracker, American crow, Common raven
    
            Including_domestic.Local_species %in% "Pipilo maculatus" ~ "Terrestrial bird", #spotted towhee
    
    Including_domestic.Local_species %in% "Onychomys torridus" ~ "Small rodent", #southern grasshopper mouse
    
    Including_domestic.Local_species %in% "Arctocephalus townsendi" ~ "Guadelupe fur seal", #guadelupe fur seal
    
    Including_domestic.Local_species %in% "Callospermophilus lateralis" ~ "Small rodent", #golden-mantled ground squirrel
    
    Including_domestic.Local_species %in% "Marmota flaviventris" ~ "Small rodent", #yellow-bellied marmot
    
    Including_domestic.Local_species %in% "Microtus montanus" ~ "Small rodent" #montane vole
  ))

# for the 16s run - broad taxonomy 1
gaviota <- gaviota %>%
  mutate(broadtax1 = case_when(
    Barcode_region == "16" & Best_match_references %in% "Bos taurus" ~ "Cattle", # 100
    Barcode_region == "16" & Best_match_references %in% "Lynx rufus" ~ "Bobcat", # 100
    #Best_match_references %in% "Heliconoides inflatus" ~ "Heliconoides inflatus", # sea butterfly snail
    #Best_match_references %in% "Odocoileus virginianus" ~ "white tailed deer", # 100
    #Best_match_references %in% "Crotalus viridis" ~ "prairie rattlesnake", # 100
    Barcode_region == "16" & Best_match_references %in% "Puma concolor" ~ "Puma", # puma
    Barcode_region == "16" & Best_match_references %in% "Rattus rattus" ~ "Small rodent", # rat
    #Best_match_references %in% "Gallus sonneratii" ~ "chicken", # 100
    Barcode_region == "16" & Best_match_references %in% "Callipepla californica" ~ "Terrestrial bird", # California quail
    #Best_match_references %in% "Scomber japonicus" ~ "chub mackerel", # 100
    Barcode_region == "16" & Best_match_references %in% "Lamna ditropis" ~ "Salmon shark", #salmon shark
    Barcode_region == "16" & Best_match_references %in% "Magnisudis atlantica" ~ "Duckbill baracuda", # duckbill barracuda
    Barcode_region == "16" & Best_match_references %in% "Ophiodon elongatus" ~ "Lingcod", # lingcod
    Barcode_region == "16" & Best_match_references %in% "Thysanoessa inermis" ~ "Krill", # Krill
    #Best_match_references %in% "Sebastes serranoides" ~ "Sebastes sp", # olive rockfish
    Barcode_region == "16" & Best_match_references %in% "Clupea pallasii" ~ "Pacific herring", # pacific herring
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa" ~ "Wild boar", #wild boar
    #Best_match_references %in% "Canis rufus" ~ "canid", #
    Barcode_region == "16" & Best_match_references %in% "Pilaira anomala" ~ "Mold", #mold
    Barcode_region == "16" & Best_match_references %in% "Reithrodontomys megalotis" ~ "Small rodent", #western harvest mouse
    Barcode_region == "16" & Best_match_references %in% "Microtus californicus" ~ "Small rodent", #California vole
    Barcode_region == "16" & Best_match_references %in% "Chaetodipus californicus" ~ "Small rodent", #California pocket mouse
    Barcode_region == "16" & Best_match_references %in% "Zalophus californianus" ~ "California sea lion", #California sea lion
    Barcode_region == "16" & Best_match_references %in% "Canis latrans" ~ "Coyote", #coyote
    Barcode_region == "16" & Best_match_references %in% "Neotoma lepida" ~ "Small rodent", #desert woodrat
    #Best_match_references %in% "Meleagris gallopavo" ~ "turkey", #
    Barcode_region == "16" & Best_match_references %in% "Otospermophilus beecheyi" ~ "Small rodent", #California ground squirrel
    Barcode_region == "16" & Best_match_references %in% "Taxidea taxus" ~ "Other terrestrial mammal", #american badger
    Barcode_region == "16" & Best_match_references %in% "Neotoma macrotis" ~ "Small rodent", #big-eared woodrat
    #Best_match_references %in% "Ravinia vagabunda" ~ "flesh fly", #
    Barcode_region == "16" & Best_match_references %in% "Canis lupus" ~ "Coyote", # canid
    Barcode_region == "16" & Best_match_references %in% "Forficula auricularia" ~ "Terrestrial insect", #European earwig
    Barcode_region == "16" & Best_match_references %in% "Mephitis mephitis" ~ "Striped skunk", #striped skunk
    #Best_match_references %in% "Paralucilia pseudolyrcea" ~ "blow fly", #
    Barcode_region == "16" & Best_match_references %in% "Urocyon cinereoargenteus" ~ "Grey fox", #grey fox
    Barcode_region == "16" & Best_match_references %in% "Pandalus jordani" ~ "Ocean pink shrimp", #ocean pink shrimp
    #Best_match_references %in% "Merluccius productus" ~ "north pacific hake", #
    Barcode_region == "16" & Best_match_references %in% "Peromyscus maniculatus" ~ "Small rodent", #deer mouse
    Barcode_region == "16" & Best_match_references %in% "Pituophis catenifer" ~ "Reptile/amphibian", #gopher snake
    Barcode_region == "16" & Best_match_references %in% "Sylvilagus bachmani" ~ "Brush rabbit", #brush rabbit
    Barcode_region == "16" & Best_match_references %in% "Plestiodon skiltonianus" ~ "Reptile/amphibian", #western skink
    Barcode_region == "16" & Best_match_references %in% "Anchoa ischana" ~ "Sharpnose anchovy", #sharpnose anchovy
    Barcode_region == "16" & Best_match_references %in% "Gonatopsis borealis" ~ "Boreopacific armhook squid", #Boreopacific armhook squid
    Barcode_region == "16" & Best_match_references %in% "Leucocarbo chalconotus" ~ "Pelagic cormorant", # cormorant species, checked with BLAST. shag (stewart island shag, but those are just in NZ)
    Barcode_region == "16" & Best_match_references %in% "Pseudopaludicola boliviana" ~ "Reptile/amphibian", # frog, Leptodactylidae family, found in regions in south america???
    Barcode_region == "16" & Best_match_references %in% "Microtus chrotorrhinus" ~ "Small rodent", # rock vole, found in eastern US
    Barcode_region == "16" & Best_match_references %in% "Picoides tridactylus" ~ "Terrestrial bird", # eurasian three-toed woodpecker, not found in US --- changed to Dryobates pubescens, downy woodpecker
    Barcode_region == "16" & Best_match_references %in% "Thomomys bottae" ~ "Small rodent", # Botta's pocket gopher
    Barcode_region == "16" & Best_match_references %in% "Neodon bershulaensis" ~ "Small rodent", # Bershula mountain vole, found in China -- BLAST shows microtus species second match
    #Best_match_references %in% "Chloropicon sieburthii" ~ "phytoplankton", # 90
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus" ~ "Terrestrial insect", # hairy rove beetle -- only shows up once
    Barcode_region == "16" & Best_match_references %in% "Rhipicephalus sanguineus" ~ "Brown dog tick", # Brown dog tick
    Barcode_region == "16" & Best_match_references %in% "Phakopsora meibomiae" ~ "fungi", # fungi
    Barcode_region == "16" & Best_match_references %in% "Tettigonia orientalis" ~ "Terrestrial insect", # bush cricket, but this species is found in Japan --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Isogenoides varians" ~ "Terrestrial insect", # stonefly
    Barcode_region == "16" & Best_match_references %in% "Setvena wahkeena" ~ "Terrestrial insect", # stonefly (found in socal?) -- BLAST - cricket match
    Barcode_region == "16" & Best_match_references %in% "Xiphidiopsis gurneyi" ~ "Terrestrial insect", #bush cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Abronia graminea" ~ "Reptile/amphibian", # alligator lizard (found in socal?) -- BLAST - southern aligator lizard
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini" ~ "Terrestrial insect", # camel cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Chloropicon laureae" ~ "Green algae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae" ~ "Terrestrial insect", #Scarabaeidae (scarab) -- changed to species that came up with this species in results -- BLAST- cricket
    Barcode_region == "16" & Best_match_references %in% "Stenopelmatus fuscus" ~ "Terrestrial insect", # 80
    Barcode_region == "16" & Best_match_references %in% "Gymnaetoides acutilobatus" ~ "Small rodent", # endemic to china??? --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Microdipodops pallidus" ~ "Terrestrial insect", # pale kangaroo mouse
    Barcode_region == "16" & Best_match_references %in% "Onthophagus crinitus" ~ "Terrestrial insect", # dung beetle
    Barcode_region == "16" & Best_match_references %in% "Pheidole deima" ~ "Terrestrial insect", # ant (found in socal?) --- BLAST - Veromessor andrei - harvester ant
    Barcode_region == "16" & Best_match_references %in% "Hazenia capsulata" ~ "Green algae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Anadara broughtonii" ~ "Inflated ark clam", # Inflated ark clam
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea" ~ "mold", # slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis" ~ "Green algae", # 70

    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos javanicus, Bos indicus, Bos grunniens, Bos grunniens x Bos taurus" ~ "Cattle",
    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos primigenius, Bos indicus" ~ "Cattle", #cow
    Barcode_region == "16" & Best_match_references %in% "Canis rufus, Canis lupus, Canis latrans" ~ "Coyote", # coyote
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea, Cavenderia pseudoaureostipes" ~ "mold", #slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Green algae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Chloropicon sieburthii, Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Green algae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus, Anthrenus scrophulariae" ~ "Terrestrial insect", # Common carpet beetle/Buffalo carpet beetle
    Barcode_region == "16" & Best_match_references %in% "Crotalus viridis, Crotalus oreganus" ~ "Reptile/amphibian", #western rattlesnake
    Barcode_region == "16" & Best_match_references %in% "Gallus sonneratii, Gallus gallus" ~ "Terrestrial bird", # wild turkey probably
    Barcode_region == "16" & Best_match_references %in% "Heliconoides inflatus, Cathaica polystigma, Nesiohelix moreletiana, Karaftohelix maackii, Atlanta turriculata, Aceria guerreronis, Hahnia ngai, Grylloprimevala jilina, Aconurella prolixa, Processina dashahensis, Tapinoma indicum, Homo sapiens" ~ "Sea butterfly snail", # Heliconoides inflatus -- Sea butterfly (range is not Socal, but Mexico, and it's the closest species in the list) 
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae, Paratriodonta lentula" ~ "Terrestrial insect", #Scarabaeidae (scarab)
    Barcode_region == "16" & Best_match_references %in% "Meleagris gallopavo, Labidesthes sicculus" ~ "Terrestrial bird", #wild turkey
    Barcode_region == "16" & Best_match_references %in% "Merluccius productus, Merluccius angustimanus" ~ "North Pacific Hake", #North Pacific Hake
    Barcode_region == "16" & Best_match_references %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Mule deer", #mule deer
    Barcode_region == "16" & Best_match_references %in% "Paralucilia pseudolyrcea, Paralucilia paraensis, Hemilucilia semidiaphana, Calliphora vomitoria, Calliphora vicina, Calliphora sinensis, Calliphora nigribarbis, Calliphora calliphoroides" ~ "Terrestrial insect", # fly
    Barcode_region == "16" & Best_match_references %in% "Ravinia vagabunda, Ravinia stimulans, Ravinia derelicta" ~ "Terrestrial insect", #flesh fly
    Barcode_region == "16" & Best_match_references %in% "Scomber japonicus, Scomber colias, Scomber australasicus"~ "Pacific makerel", #pacific mackerel
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa, Thryonomys swinderianus" ~ "Wild boar", #wild pig
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini, Onthophagus obscurior" ~ "Terrestrial insect", #dung beetle
    Barcode_region == "16" & Best_match_references %in% "Sebastes serranoides, Sebastes melanops, Sebastes flavidus" ~ "Rockfish",  #rockfish
    Barcode_region == "16" & Best_match_references %in% "" ~ "no match",
    TRUE ~ broadtax1
    ))



checkna <- gaviota[is.na(gaviota$broadtax1), ] 
#rm(checkna)

#write.csv(alaska, "alaska.csv")
```



genus name
```{r}
gaviota$genus = NA

gaviota <- gaviota %>%
  mutate(genus = case_when(
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus" ~ "Canis", # coyote
    
    Including_domestic.Local_species %in% "Phalacrocorax penicillatus, Phalacrocorax brasilianus, Phalacrocorax pelagicus" 
      ~ "Urile", # Brandt's cormorant
    
    Including_domestic.Local_species %in% "Phalacrocorax pelagicus, Phalacrocorax penicillatus" 
      ~ "Urile", # Brandt's cormorant
    
    Including_domestic.Local_species %in% "Microtus californicus" 
      ~ "Microtus", # Californian vole
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei" ~ "Peromyscus", #Peromyscus - deer mouse
    
    Including_domestic.Local_species %in% "Sylvilagus bachmani" ~ "Sylvilagus", # Brush rabbit
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Arborimus albipes, Arborimus longicaudus, Arborimus pomo, Myodes californicus, Lemmiscus curtatus, Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii, Ondatra zibethicus, Phenacomys intermedius, Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis, Onychomys leucogaster, Onychomys torridus, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei, Reithrodontomys raviventris, Reithrodontomys megalotis, Sigmodon arizonae, Sigmodon hispidus" ~ "Peromyscus", # 	Cricetidae (mouse or vole)
      
    Including_domestic.Local_species %in% "Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis" ~ "Neotoma", # Neotoma - woodrat
      
    Including_domestic.Local_species %in% "Bos taurus, Bos indicus, Bos javanicus, Bos frontalis" ~ "Bos", # Bovine
      
    Including_domestic.Local_species %in% "Gallus gallus" ~ "Gallus", #chicken
      
    Including_domestic.Local_species %in% "Urocyon cinereoargenteus, Urocyon littoralis" ~ "Urocyon", # grey fox
      
    Including_domestic.Local_species %in% "Glaucomys oregonensis, Glaucomys sabrinus, Callospermophilus lateralis" ~ "Sciuridae", # flying squirrel/golden mantle ground squirrel
      
    Including_domestic.Local_species %in% "Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii" ~ "Microtus", # vole
    Including_domestic.Local_species %in% "Thomomys bottae" ~"Thomomys", # botta's pocket gopher
    
    Including_domestic.Local_species %in% "Lynx rufus, Puma concolor" ~ "Lynx", # bobcat
      
    Including_domestic.Local_species %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Odocoileus", # mule deer
      
    Including_domestic.Local_species %in% "Sus scrofa" ~ "Sus", #wild boar
      
    Including_domestic.Local_species %in% "Chloris sinica, Spinus pinus, Plectrophenax nivalis, Zonotrichia leucophrys, Zonotrichia atricapilla, Zonotrichia albicollis, Sturnella neglecta, Sturnella magna, Serinus canaria" ~ "Zonotrichia", # Sparrow
      
    Including_domestic.Local_species %in% "Peromyscus eremicus" ~ "Peromyscus", # cactus mouse
      
    Including_domestic.Local_species %in% "Zalophus californianus" ~ "Zalophus", # sea lion
    
    Including_domestic.Local_species %in% "Canis lupus" ~ "Canis", # coyote
      
    Including_domestic.Local_species %in% "Peromyscus crinitus, Peromyscus eremicus" ~ "Peromyscus", # Peromyscus
      
    Including_domestic.Local_species %in% "Callipepla californica, Callipepla gambelii" ~ "Callipepla", # CA quail 
      
    Including_domestic.Local_species %in% "Alosa sapidissima, Dorosoma petenense" ~ "Alosa", # American shad
      
    Including_domestic.Local_species %in% "Leuconotopicus albolarvatus, Jynx torquilla, Melanerpes lewis, Melanerpes erythrocephalus, Melanerpes formicivorus, Melanerpes uropygialis, Sphyrapicus thyroideus, Sphyrapicus varius, Sphyrapicus nuchalis, Sphyrapicus ruber, Picoides arcticus, Dryobates pubescens, Dryobates nuttallii, Dryobates scalaris, Picoides villosus, Colaptes auratus, Colaptes chrysoides, Dryocopus pileatus" ~ "Melanerpes", # acorn woodpecker
      
    Including_domestic.Local_species %in% "Plestiodon gilberti" ~ "Plestiodon", # Gilbert's skink
      
    Including_domestic.Local_species %in% "Taxidea taxus" ~ "Taxidea", # American badger
      
    Including_domestic.Local_species %in% "Microtus townsendii" ~ "Microtus", # Microtus - townsend's vole
      
    Including_domestic.Local_species %in% "Neotoma macrotis" ~ "Neotoma", # big eared woodrat
    
    Including_domestic.Local_species %in% "Aphelocoma insularis, Aphelocoma californica, Aphelocoma woodhouseii" ~ "Aphelocoma", # CA scrub jay
      
    Including_domestic.Local_species %in% "Puma concolor" ~ "Puma", # puma
      
    Including_domestic.Local_species %in% "Meleagris gallopavo, Gallus gallus" ~ "Meleagris", # wild turkey
      
    Including_domestic.Local_species %in% "Mephitis mephitis" ~ "Mephitis", # striped skunk
      
    Including_domestic.Local_species %in% "Vulpes vulpes" ~ "Vulpes", # red fox
      
    Including_domestic.Local_species %in% "Chaetodipus californicus" ~ "Chaetodipus", # California pocket mouse
      
    Including_domestic.Local_species %in% "Mus musculus" ~ "Mus", # house mouse
    
    Including_domestic.Local_species %in% "Rattus rattus" ~ "Rattus", # black rat
      
    Including_domestic.Local_species %in% "Peromyscus maniculatus, Peromyscus keeni" ~ "Peromyscus", # deer mice
      
    Including_domestic.Local_species %in% "Sylvilagus floridanus, Sylvilagus nuttallii" ~ "Sylvilagus", # brush rabbit
      
    Including_domestic.Local_species %in% "Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii" ~ "Microtus", # vole
      
    Including_domestic.Local_species %in% "Lemmiscus curtatus, Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii, Ondatra zibethicus, Onychomys leucogaster, Synaptomys borealis, Peromyscus maniculatus, Peromyscus keeni" ~ "Microtus", # cricetidae - some kind of rodent (choosing microtus)
   
     Including_domestic.Local_species %in% "Neotoma lepida" ~ "Neotoma", # desert woodrat
    
     Including_domestic.Local_species %in% "Canis latrans, Canis lupus, Urocyon cinereoargenteus, Urocyon littoralis, Vulpes macrotis, Vulpes vulpes" ~ "Canis", # coyote
    
    Including_domestic.Local_species %in% "Turdus obscurus, Turdus migratorius" ~ "Turdus", #American robin
    
    Including_domestic.Local_species %in% "Spizella passerina, Passerella iliaca, Melospiza melodia" ~ "Melospiza", #Song sparrow, could  be chipping or fox sparrow too though.
    
    Including_domestic.Local_species %in% "Ascaphus truei, Incilius alvarius, Anaxyrus boreas, Anaxyrus californicus, Anaxyrus canorus, Anaxyrus cognatus, Anaxyrus exsul, Anaxyrus microscaphus, Anaxyrus punctatus, Anaxyrus woodhousii, Pseudacris cadaverina, Pseudacris hypochondriaca, Pseudacris regilla, Pseudacris sierra, Scaphiopus couchii, Spea hammondii, Spea intermontana, Rana aurora, Lithobates berlandieri, Rana boylii, Rana cascadae, Lithobates catesbeianus, Rana draytonii, Rana luteiventris, Rana muscosa, Lithobates pipiens, Rana pretiosa, Rana sierrae, Lithobates sphenocephalus, Lithobates yavapaiensis, Xenopus laevis" ~ "Rana", #frog or toad -- choosing Rana as genus
    
    Including_domestic.Local_species %in% "Phoca fasciata, Phoca vitulina" ~ "Phoca", #harbor seal
    
    Including_domestic.Local_species %in% "Nucifraga columbiana, Corvus brachyrhynchos, Corvus corax" ~ "Corvus", #corvidae -- Clark's nutcracker, American crow, Common raven
    
    Including_domestic.Local_species %in% "Pipilo maculatus" ~ "Pipilo", #spotted towhee
    
    Including_domestic.Local_species %in% "Onychomys torridus" ~ "Onychomys", #southern grasshopper mouse
    
    Including_domestic.Local_species %in% "Arctocephalus townsendi" ~ "Arctocephalus", #guadelupe fur seal
    
    Including_domestic.Local_species %in% "Callospermophilus lateralis" ~ "Callospermophilus", #golden-mantled ground squirrel
    
    Including_domestic.Local_species %in% "Marmota flaviventris" ~ "Marmota", #yellow-bellied marmot
    
    Including_domestic.Local_species %in% "Microtus montanus" ~ "Microtus" #montane vole
  ))

# for the 16s run
gaviota <- gaviota %>%
  mutate(genus = case_when(
    Barcode_region == "16" & Best_match_references %in% "Bos taurus" ~ "Bos", # 100
    Barcode_region == "16" & Best_match_references %in% "Lynx rufus" ~ "Lynx", # 100
    #Best_match_references %in% "Heliconoides inflatus" ~ "Heliconoides inflatus", # sea butterfly snail
    #Best_match_references %in% "Odocoileus virginianus" ~ "white tailed deer", # 100
    #Best_match_references %in% "Crotalus viridis" ~ "prairie rattlesnake", # 100
    Barcode_region == "16" & Best_match_references %in% "Puma concolor" ~ "Puma", # puma
    Barcode_region == "16" & Best_match_references %in% "Rattus rattus" ~ "Rattus", # rat
    #Best_match_references %in% "Gallus sonneratii" ~ "chicken", # 100
    Barcode_region == "16" & Best_match_references %in% "Callipepla californica" ~ "Callipepla", # California quail
    #Best_match_references %in% "Scomber japonicus" ~ "chub mackerel", # 100
    Barcode_region == "16" & Best_match_references %in% "Lamna ditropis" ~ "Lamna", #salmon shark
    Barcode_region == "16" & Best_match_references %in% "Magnisudis atlantica" ~ "Magnisudis", # duckbill barracuda
    Barcode_region == "16" & Best_match_references %in% "Ophiodon elongatus" ~ "Ophiodon", # lingcod
    Barcode_region == "16" & Best_match_references %in% "Thysanoessa inermis" ~ "Thysanoessa", # Krill
    #Best_match_references %in% "Sebastes serranoides" ~ "Sebastes sp", # olive rockfish
    Barcode_region == "16" & Best_match_references %in% "Clupea pallasii" ~ "Clupea", # pacific herring
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa" ~ "Sus", #wild boar
    #Best_match_references %in% "Canis rufus" ~ "canid", #
    Barcode_region == "16" & Best_match_references %in% "Pilaira anomala" ~ "Pilaira", #mold
    Barcode_region == "16" & Best_match_references %in% "Reithrodontomys megalotis" ~ "Reithrodontomys", #western harvest mouse
    Barcode_region == "16" & Best_match_references %in% "Microtus californicus" ~ "Microtus", #California vole
    Barcode_region == "16" & Best_match_references %in% "Chaetodipus californicus" ~ "Chaetodipus", #California pocket mouse
    Barcode_region == "16" & Best_match_references %in% "Zalophus californianus" ~ "Zalophus", #California sea lion
    Barcode_region == "16" & Best_match_references %in% "Canis latrans" ~ "Canis", #coyote
    Barcode_region == "16" & Best_match_references %in% "Neotoma lepida" ~ "Neotoma", #desert woodrat
    #Best_match_references %in% "Meleagris gallopavo" ~ "turkey", #
    Barcode_region == "16" & Best_match_references %in% "Otospermophilus beecheyi" ~ "Otospermophilus", #California ground squirrel
    Barcode_region == "16" & Best_match_references %in% "Taxidea taxus" ~ "Taxidea", #american badger
    Barcode_region == "16" & Best_match_references %in% "Neotoma macrotis" ~ "Neotoma", #big-eared woodrat
    #Best_match_references %in% "Ravinia vagabunda" ~ "flesh fly", #
    Barcode_region == "16" & Best_match_references %in% "Canis lupus" ~ "Canis", # canid
    Barcode_region == "16" & Best_match_references %in% "Forficula auricularia" ~ "Forficula", #European earwig
    Barcode_region == "16" & Best_match_references %in% "Mephitis mephitis" ~ "Mephitis", #striped skunk
    #Best_match_references %in% "Paralucilia pseudolyrcea" ~ "blow fly", #
    Barcode_region == "16" & Best_match_references %in% "Urocyon cinereoargenteus" ~ "Urocyon", #grey fox
    Barcode_region == "16" & Best_match_references %in% "Pandalus jordani" ~ "Pandalus", #ocean pink shrimp
    #Best_match_references %in% "Merluccius productus" ~ "north pacific hake", #
    Barcode_region == "16" & Best_match_references %in% "Peromyscus maniculatus" ~ "Peromyscus", #deer mouse
    Barcode_region == "16" & Best_match_references %in% "Pituophis catenifer" ~ "Pituophis", #gopher snake
    Barcode_region == "16" & Best_match_references %in% "Sylvilagus bachmani" ~ "Sylvilagus", #brush rabbit
    Barcode_region == "16" & Best_match_references %in% "Plestiodon skiltonianus" ~ "Plestiodon", #western skink
    Barcode_region == "16" & Best_match_references %in% "Anchoa ischana" ~ "Anchoa", #sharpnose anchovy
    Barcode_region == "16" & Best_match_references %in% "Gonatopsis borealis" ~ "Gonatopsis", #Boreopacific armhook squid
    Barcode_region == "16" & Best_match_references %in% "Leucocarbo chalconotus" ~ "Urile", # cormorant species, checked with BLAST. shag (stewart island shag, but those are just in NZ)
    Barcode_region == "16" & Best_match_references %in% "Pseudopaludicola boliviana" ~ "Pseudopaludicola", # frog, Leptodactylidae family, found in regions in south america???
    Barcode_region == "16" & Best_match_references %in% "Microtus chrotorrhinus" ~ "Microtus", # rock vole, found in eastern US
    Barcode_region == "16" & Best_match_references %in% "Picoides tridactylus" ~ "Dryobates", # eurasian three-toed woodpecker, not found in US --- changed to Dryobates pubescens, downy woodpecker
    Barcode_region == "16" & Best_match_references %in% "Thomomys bottae" ~ "Thomomys", # Botta's pocket gopher
    Barcode_region == "16" & Best_match_references %in% "Neodon bershulaensis" ~ "Microtus", # Bershula mountain vole, found in China -- BLAST shows microtus species second match
    #Best_match_references %in% "Chloropicon sieburthii" ~ "phytoplankton", # 90
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus" ~ "Creophilus", # hairy rove beetle -- only shows up once
    Barcode_region == "16" & Best_match_references %in% "Rhipicephalus sanguineus" ~ "Rhipicephalus", # Brown dog tick
    Barcode_region == "16" & Best_match_references %in% "Phakopsora meibomiae" ~ "Phakopsora", # fungi
    Barcode_region == "16" & Best_match_references %in% "Tettigonia orientalis" ~ "Stenopelmatus", # bush cricket, but this species is found in Japan ---BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Isogenoides varians" ~ "Isogenoides", # stonefly
    Barcode_region == "16" & Best_match_references %in% "Setvena wahkeena" ~ "Stenopelmatus", # stonefly (found in socal?) -- BLAST - cricket match
    Barcode_region == "16" & Best_match_references %in% "Xiphidiopsis gurneyi" ~ "Stenopelmatus", #bush cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Abronia graminea" ~ "Elgaria", # alligator lizard (found in socal?) -- BLAST - southern aligator lizard
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini" ~ "Stenopelmatus", # camel cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Chloropicon laureae" ~ "Chloropicon", # green algae
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae" ~ "Stenopelmatus", #Scarabaeidae (scarab) -- changed to species that came up with this species in results -- BLAST- cricket
    Barcode_region == "16" & Best_match_references %in% "Stenopelmatus fuscus" ~ "Stenopelmatus", # jerusalem cricket
    Barcode_region == "16" & Best_match_references %in% "Gymnaetoides acutilobatus" ~ "Stenopelmatus", # endemic to china??? cave cricket --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Microdipodops pallidus" ~ "Microdipodops", # pale kangaroo mouse
    Barcode_region == "16" & Best_match_references %in% "Onthophagus crinitus" ~ "Onthophagus", # dung beetle
    Barcode_region == "16" & Best_match_references %in% "Pheidole deima" ~ "Veromessor", # ant (found in socal?) --- BLAST - Veromessor andrei - harvestor ant
    Barcode_region == "16" & Best_match_references %in% "Hazenia capsulata" ~ "Hazenia", # green algae
    Barcode_region == "16" & Best_match_references %in% "Anadara broughtonii" ~ "Anadara", # Inflated ark clam
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea" ~ "Cavenderia", # slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis" ~ "Cloropicon", # 70

    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos javanicus, Bos indicus, Bos grunniens, Bos grunniens x Bos taurus" ~ "Bos",
    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos primigenius, Bos indicus" ~ "Bos", #cow
    Barcode_region == "16" & Best_match_references %in% "Canis rufus, Canis lupus, Canis latrans" ~ "Canis", # coyote
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea, Cavenderia pseudoaureostipes" ~ "Cavenderia", #slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Chloropicon", #green algae
    Barcode_region == "16" & Best_match_references %in% "Chloropicon sieburthii, Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Chloropicon", #green algae
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus, Anthrenus scrophulariae" ~ "Anthrenus", # Common carpet beetle/Buffalo carpet beetle
    Barcode_region == "16" & Best_match_references %in% "Crotalus viridis, Crotalus oreganus" ~ "Crotalus", #western rattlesnake
    Barcode_region == "16" & Best_match_references %in% "Gallus sonneratii, Gallus gallus" ~ "Gallus", # wild turkey probably, or chicken
    Barcode_region == "16" & Best_match_references %in% "Heliconoides inflatus, Cathaica polystigma, Nesiohelix moreletiana, Karaftohelix maackii, Atlanta turriculata, Aceria guerreronis, Hahnia ngai, Grylloprimevala jilina, Aconurella prolixa, Processina dashahensis, Tapinoma indicum, Homo sapiens" ~ "Heliconoides", # Heliconoides inflatus -- Sea butterfly (range is not Socal, but Mexico, and it's the closest species in the list) 
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae, Paratriodonta lentula" ~ "Paratriodonta", #Scarabaeidae (scarab)
    Barcode_region == "16" & Best_match_references %in% "Meleagris gallopavo, Labidesthes sicculus" ~ "Meleagris", #wild turkey
    Barcode_region == "16" & Best_match_references %in% "Merluccius productus, Merluccius angustimanus" ~ "Merluccius", #North Pacific Hake
    Barcode_region == "16" & Best_match_references %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Odocoileus", #mule deer
    Barcode_region == "16" & Best_match_references %in% "Paralucilia pseudolyrcea, Paralucilia paraensis, Hemilucilia semidiaphana, Calliphora vomitoria, Calliphora vicina, Calliphora sinensis, Calliphora nigribarbis, Calliphora calliphoroides" ~ "Paralucilia", # fly -- choosing Paralucilia
    Barcode_region == "16" & Best_match_references %in% "Ravinia vagabunda, Ravinia stimulans, Ravinia derelicta" ~ "Ravinia", #flesh fly
    Barcode_region == "16" & Best_match_references %in% "Scomber japonicus, Scomber colias, Scomber australasicus"~ "Scomber", #pacific mackerel
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa, Thryonomys swinderianus" ~ "Sus", #wild pig
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini, Onthophagus obscurior" ~ "Onthophagus", #dung beetle
    Barcode_region == "16" & Best_match_references %in% "Sebastes serranoides, Sebastes melanops, Sebastes flavidus" ~ "Sebastes",  #rockfish
    Barcode_region == "16" & Best_match_references %in% "" ~ "no match",
    TRUE ~ genus
    ))


 

checkna <- gaviota[is.na(gaviota$genus), ] 
```



family name
```{r}
gaviota$family = NA

gaviota <- gaviota %>%
  mutate(family = case_when(
    Including_domestic.Local_species %in% "Canis latrans, Canis lupus" ~ "Canidae", # coyote
    
    Including_domestic.Local_species %in% "Phalacrocorax penicillatus, Phalacrocorax brasilianus, Phalacrocorax pelagicus" 
      ~ "Phalacrocoracidae", # Brandt's cormorant
    
    Including_domestic.Local_species %in% "Phalacrocorax pelagicus, Phalacrocorax penicillatus" 
      ~ "Phalacrocoracidae", # Brandt's cormorant
    
    Including_domestic.Local_species %in% "Microtus californicus" 
      ~ "Cricetidae", # Californian vole
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei" ~ "Cricetidae", #Peromyscus - deer mouse
    
    Including_domestic.Local_species %in% "Sylvilagus bachmani" ~ "Leporidae", # Brush rabbit
    
    Including_domestic.Local_species %in% "Peromyscus gambelii, Arborimus albipes, Arborimus longicaudus, Arborimus pomo, Myodes californicus, Lemmiscus curtatus, Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii, Ondatra zibethicus, Phenacomys intermedius, Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis, Onychomys leucogaster, Onychomys torridus, Peromyscus boylii, Peromyscus californicus, Peromyscus crinitus, Peromyscus eremicus, Peromyscus fraterculus, Peromyscus sonoriensis, Peromyscus truei, Reithrodontomys raviventris, Reithrodontomys megalotis, Sigmodon arizonae, Sigmodon hispidus" ~ "Cricetidae", # 	Cricetidae (mouse or vole)
      
    Including_domestic.Local_species %in% "Neotoma albigula, Neotoma cinerea, Neotoma fuscipes, Neotoma lepida, Neotoma macrotis" ~ "Cricetidae", # Neotoma - woodrat
      
    Including_domestic.Local_species %in% "Bos taurus, Bos indicus, Bos javanicus, Bos frontalis" ~ "Bovidae", # Bovine
      
    Including_domestic.Local_species %in% "Gallus gallus" ~ "Phasianidae", #chicken
      
    Including_domestic.Local_species %in% "Urocyon cinereoargenteus, Urocyon littoralis" ~ "Canidae", # grey fox
      
    Including_domestic.Local_species %in% "Glaucomys oregonensis, Glaucomys sabrinus, Callospermophilus lateralis" ~ "Sciuridae", # flying squirrel/golden mantle ground squirrel
      
    Including_domestic.Local_species %in% "Microtus californicus, Microtus longicaudus, Microtus montanus, Microtus oregoni, Microtus townsendii" ~ "Cricetidae", # vole
    Including_domestic.Local_species %in% "Thomomys bottae" ~"Cricetidae", # botta's pocket gopher
    
    Including_domestic.Local_species %in% "Lynx rufus, Puma concolor" ~ "Felidae", # bobcat
      
    Including_domestic.Local_species %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Cervidae", # mule deer
      
    Including_domestic.Local_species %in% "Sus scrofa" ~ "Suidae", #wild boar
      
    Including_domestic.Local_species %in% "Chloris sinica, Spinus pinus, Plectrophenax nivalis, Zonotrichia leucophrys, Zonotrichia atricapilla, Zonotrichia albicollis, Sturnella neglecta, Sturnella magna, Serinus canaria" ~ "Passerellidae", # Sparrow
      
    Including_domestic.Local_species %in% "Peromyscus eremicus" ~ "Cricetidae", # cactus mouse
      
    Including_domestic.Local_species %in% "Zalophus californianus" ~ "Otariidae", # sea lion
    
    Including_domestic.Local_species %in% "Canis lupus" ~ "Canidae", # coyote
      
    Including_domestic.Local_species %in% "Peromyscus crinitus, Peromyscus eremicus" ~ "Cricetidae", # Peromyscus
      
    Including_domestic.Local_species %in% "Callipepla californica, Callipepla gambelii" ~ "Odontophoridae", # CA quail 
      
    Including_domestic.Local_species %in% "Alosa sapidissima, Dorosoma petenense" ~ "Alosidae", # American shad
      
    Including_domestic.Local_species %in% "Leuconotopicus albolarvatus, Jynx torquilla, Melanerpes lewis, Melanerpes erythrocephalus, Melanerpes formicivorus, Melanerpes uropygialis, Sphyrapicus thyroideus, Sphyrapicus varius, Sphyrapicus nuchalis, Sphyrapicus ruber, Picoides arcticus, Dryobates pubescens, Dryobates nuttallii, Dryobates scalaris, Picoides villosus, Colaptes auratus, Colaptes chrysoides, Dryocopus pileatus" ~ "Picidae", # acorn woodpecker
      
    Including_domestic.Local_species %in% "Plestiodon gilberti" ~ "Scincidae", # Gilbert's skink
      
    Including_domestic.Local_species %in% "Taxidea taxus" ~ "Mustelidae", # American badger
      
    Including_domestic.Local_species %in% "Microtus townsendii" ~ "Cricetidae", # Microtus - townsend's vole
      
    Including_domestic.Local_species %in% "Neotoma macrotis" ~ "Cricetidae", # big eared woodrat
    
    Including_domestic.Local_species %in% "Aphelocoma insularis, Aphelocoma californica, Aphelocoma woodhouseii" ~ "Corvidae", # CA scrub jay
      
    Including_domestic.Local_species %in% "Puma concolor" ~ "Felidae", # puma
      
    Including_domestic.Local_species %in% "Meleagris gallopavo, Gallus gallus" ~ "Phasianidae", # wild turkey
      
    Including_domestic.Local_species %in% "Mephitis mephitis" ~ "Mephitidae", # striped skunk
      
    Including_domestic.Local_species %in% "Vulpes vulpes" ~ "Canidae", # red fox
      
    Including_domestic.Local_species %in% "Chaetodipus californicus" ~ "Heteromyidae", # California pocket mouse
      
    Including_domestic.Local_species %in% "Mus musculus" ~ "Muridae", # house mouse
    
    Including_domestic.Local_species %in% "Rattus rattus" ~ "Muridae", # black rat
      
    Including_domestic.Local_species %in% "Peromyscus maniculatus, Peromyscus keeni" ~ "Cricetidae", # deer mice
      
    Including_domestic.Local_species %in% "Sylvilagus floridanus, Sylvilagus nuttallii" ~ "Leporidae", # brush rabbit
      
    Including_domestic.Local_species %in% "Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii" ~ "Cricetidae", # vole
      
    Including_domestic.Local_species %in% "Lemmiscus curtatus, Microtus canicaudus, Microtus pennsylvanicus, Microtus townsendii, Ondatra zibethicus, Onychomys leucogaster, Synaptomys borealis, Peromyscus maniculatus, Peromyscus keeni" ~ "Cricetidae", # cricetidae - some kind of rodent (choosing microtus)
   
     Including_domestic.Local_species %in% "Neotoma lepida" ~ "Cricetidae", # desert woodrat
    
     Including_domestic.Local_species %in% "Canis latrans, Canis lupus, Urocyon cinereoargenteus, Urocyon littoralis, Vulpes macrotis, Vulpes vulpes" ~ "Canidae", # coyote
    
    Including_domestic.Local_species %in% "Turdus obscurus, Turdus migratorius" ~ "Turdidae", #American robin
    
    Including_domestic.Local_species %in% "Spizella passerina, Passerella iliaca, Melospiza melodia" ~ "Passerellidae", #Song sparrow, could  be chipping or fox sparrow too though.
    
    Including_domestic.Local_species %in% "Ascaphus truei, Incilius alvarius, Anaxyrus boreas, Anaxyrus californicus, Anaxyrus canorus, Anaxyrus cognatus, Anaxyrus exsul, Anaxyrus microscaphus, Anaxyrus punctatus, Anaxyrus woodhousii, Pseudacris cadaverina, Pseudacris hypochondriaca, Pseudacris regilla, Pseudacris sierra, Scaphiopus couchii, Spea hammondii, Spea intermontana, Rana aurora, Lithobates berlandieri, Rana boylii, Rana cascadae, Lithobates catesbeianus, Rana draytonii, Rana luteiventris, Rana muscosa, Lithobates pipiens, Rana pretiosa, Rana sierrae, Lithobates sphenocephalus, Lithobates yavapaiensis, Xenopus laevis" ~ "Ranidae", #frog or toad -- choosing Rana as genus
    
    Including_domestic.Local_species %in% "Phoca fasciata, Phoca vitulina" ~ "Phocidae", #harbor seal
    
    Including_domestic.Local_species %in% "Nucifraga columbiana, Corvus brachyrhynchos, Corvus corax" ~ "Corvidae", #corvidae -- Clark's nutcracker, American crow, Common raven
    
    Including_domestic.Local_species %in% "Pipilo maculatus" ~ "Passerellidae", #spotted towhee
    
    Including_domestic.Local_species %in% "Onychomys torridus" ~ "Cricetidae", #southern grasshopper mouse
    
    Including_domestic.Local_species %in% "Arctocephalus townsendi" ~ "Otariidae", #guadelupe fur seal
    
    Including_domestic.Local_species %in% "Callospermophilus lateralis" ~ "Sciuridae", #golden-mantled ground squirrel
    
    Including_domestic.Local_species %in% "Marmota flaviventris" ~ "Sciuridae", #yellow-bellied marmot
    
    Including_domestic.Local_species %in% "Microtus montanus" ~ "Cricetidae" #montane vole
  ))

# for the 16s run
gaviota <- gaviota %>%
  mutate(family = case_when(
    Barcode_region == "16" & Best_match_references %in% "Bos taurus" ~ "Bovidae", # 100
    Barcode_region == "16" & Best_match_references %in% "Lynx rufus" ~ "Felidae", # 100
    #Best_match_references %in% "Heliconoides inflatus" ~ "Heliconoides inflatus", # sea butterfly snail
    #Best_match_references %in% "Odocoileus virginianus" ~ "white tailed deer", # 100
    #Best_match_references %in% "Crotalus viridis" ~ "prairie rattlesnake", # 100
    Barcode_region == "16" & Best_match_references %in% "Puma concolor" ~ "Felidae", # puma
    Barcode_region == "16" & Best_match_references %in% "Rattus rattus" ~ "Muridae", # rat
    #Best_match_references %in% "Gallus sonneratii" ~ "chicken", # 100
    Barcode_region == "16" & Best_match_references %in% "Callipepla californica" ~ "Odontophoridae", # California quail
    #Best_match_references %in% "Scomber japonicus" ~ "chub mackerel", # 100
    Barcode_region == "16" & Best_match_references %in% "Lamna ditropis" ~ "Lamnidae", #salmon shark
    Barcode_region == "16" & Best_match_references %in% "Magnisudis atlantica" ~ "Paralepididae", # duckbill barracuda
    Barcode_region == "16" & Best_match_references %in% "Ophiodon elongatus" ~ "Hexagrammidae", # lingcod
    Barcode_region == "16" & Best_match_references %in% "Thysanoessa inermis" ~ "Euphausiidae", # Krill
    #Best_match_references %in% "Sebastes serranoides" ~ "Sebastes sp", # olive rockfish
    Barcode_region == "16" & Best_match_references %in% "Clupea pallasii" ~ "Clupeidae", # pacific herring
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa" ~ "Suidae", #wild boar
    #Best_match_references %in% "Canis rufus" ~ "canid", #
    Barcode_region == "16" & Best_match_references %in% "Pilaira anomala" ~ "Pilobolaceae", #mold
    Barcode_region == "16" & Best_match_references %in% "Reithrodontomys megalotis" ~ "Cricetidae", #western harvest mouse
    Barcode_region == "16" & Best_match_references %in% "Microtus californicus" ~ "Cricetidae", #California vole
    Barcode_region == "16" & Best_match_references %in% "Chaetodipus californicus" ~ "Heteromyidae", #California pocket mouse
    Barcode_region == "16" & Best_match_references %in% "Zalophus californianus" ~ "Otariidae", #California sea lion
    Barcode_region == "16" & Best_match_references %in% "Canis latrans" ~ "Canidae", #coyote
    Barcode_region == "16" & Best_match_references %in% "Neotoma lepida" ~ "Cricetidae", #desert woodrat
    #Best_match_references %in% "Meleagris gallopavo" ~ "turkey", #
    Barcode_region == "16" & Best_match_references %in% "Otospermophilus beecheyi" ~ "Sciuridae", #California ground squirrel
    Barcode_region == "16" & Best_match_references %in% "Taxidea taxus" ~ "Mustelidae", #american badger
    Barcode_region == "16" & Best_match_references %in% "Neotoma macrotis" ~ "Cricetidae", #big-eared woodrat
    #Best_match_references %in% "Ravinia vagabunda" ~ "flesh fly", #
    Barcode_region == "16" & Best_match_references %in% "Canis lupus" ~ "Canidae", # canid
    Barcode_region == "16" & Best_match_references %in% "Forficula auricularia" ~ "Forficulidae", #European earwig
    Barcode_region == "16" & Best_match_references %in% "Mephitis mephitis" ~ "MephMephitidaeitis", #striped skunk
    #Best_match_references %in% "Paralucilia pseudolyrcea" ~ "blow fly", #
    Barcode_region == "16" & Best_match_references %in% "Urocyon cinereoargenteus" ~ "Canidae", #grey fox
    Barcode_region == "16" & Best_match_references %in% "Pandalus jordani" ~ "Pandalidae", #ocean pink shrimp
    #Best_match_references %in% "Merluccius productus" ~ "north pacific hake", #
    Barcode_region == "16" & Best_match_references %in% "Peromyscus maniculatus" ~ "Cricetidae", #deer mouse
    Barcode_region == "16" & Best_match_references %in% "Pituophis catenifer" ~ "Colubridae", #gopher snake
    Barcode_region == "16" & Best_match_references %in% "Sylvilagus bachmani" ~ "Leporidae", #brush rabbit
    Barcode_region == "16" & Best_match_references %in% "Plestiodon skiltonianus" ~ "Scincidae", #western skink
    Barcode_region == "16" & Best_match_references %in% "Anchoa ischana" ~ "Engraulidae", #sharpnose anchovy
    Barcode_region == "16" & Best_match_references %in% "Gonatopsis borealis" ~ "Gonatidae", #Boreopacific armhook squid
    Barcode_region == "16" & Best_match_references %in% "Leucocarbo chalconotus" ~ "Phalacrocoracidae", # cormorant species, checked with BLAST. shag (stewart island shag, but those are just in NZ)
    Barcode_region == "16" & Best_match_references %in% "Pseudopaludicola boliviana" ~ "Leptodactylidae", # frog, Leptodactylidae family, found in regions in south america???
    Barcode_region == "16" & Best_match_references %in% "Microtus chrotorrhinus" ~ "Cricetidae", # rock vole, found in eastern US
    Barcode_region == "16" & Best_match_references %in% "Picoides tridactylus" ~ "Picidae", # eurasian three-toed woodpecker, not found in US --- changed to Dryobates pubescens, downy woodpecker
    Barcode_region == "16" & Best_match_references %in% "Thomomys bottae" ~ "Geomyidae", # Botta's pocket gopher
    Barcode_region == "16" & Best_match_references %in% "Neodon bershulaensis" ~ "Cricetidae", # Bershula mountain vole, found in China -- BLAST shows microtus species second match
    #Best_match_references %in% "Chloropicon sieburthii" ~ "phytoplankton", # 90
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus" ~ "Staphylinidae", # hairy rove beetle -- only shows up once
    Barcode_region == "16" & Best_match_references %in% "Rhipicephalus sanguineus" ~ "Ixodidae", # Brown dog tick
    Barcode_region == "16" & Best_match_references %in% "Phakopsora meibomiae" ~ "Phakopsoraceae", # fungi
    Barcode_region == "16" & Best_match_references %in% "Tettigonia orientalis" ~ "Stenopelmatidae", # bush cricket, but this species is found in Japan ---BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Isogenoides varians" ~ "Perlodidae", # stonefly
    Barcode_region == "16" & Best_match_references %in% "Setvena wahkeena" ~ "Stenopelmatidae", # stonefly (found in socal?) -- BLAST - cricket match
    Barcode_region == "16" & Best_match_references %in% "Xiphidiopsis gurneyi" ~ "Stenopelmatidae", #bush cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Abronia graminea" ~ "Anguidae", # alligator lizard (found in socal?) -- BLAST - southern aligator lizard
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini" ~ "Stenopelmatidae", # camel cricket (found in socal?) --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Chloropicon laureae" ~ "Chloropicaceae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae" ~ "Stenopelmatidae", #Scarabaeidae (scarab) -- changed to species that came up with this species in results -- BLAST- cricket
    Barcode_region == "16" & Best_match_references %in% "Stenopelmatus fuscus" ~ "Stenopelmatidae", # jerusalem cricket
    Barcode_region == "16" & Best_match_references %in% "Gymnaetoides acutilobatus" ~ "Stenopelmatidae", # endemic to china??? cave cricket --- BLAST - cricket
    Barcode_region == "16" & Best_match_references %in% "Microdipodops pallidus" ~ "Heteromyidae", # pale kangaroo mouse
    Barcode_region == "16" & Best_match_references %in% "Onthophagus crinitus" ~ "Scarabaeidae", # dung beetle
    Barcode_region == "16" & Best_match_references %in% "Pheidole deima" ~ "Formicidae", # ant (found in socal?) --- BLAST - Veromessor andrei - harvestor ant
    Barcode_region == "16" & Best_match_references %in% "Hazenia capsulata" ~ "Gayraliaceae", # green algae
    Barcode_region == "16" & Best_match_references %in% "Anadara broughtonii" ~ "Arcidae", # Inflated ark clam
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea" ~ "Cavenderiaceae", # slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis" ~ "Chloropicaceae", # 70

    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos javanicus, Bos indicus, Bos grunniens, Bos grunniens x Bos taurus" ~ "Bovidae",
    Barcode_region == "16" & Best_match_references %in% "Bos taurus, Bos primigenius, Bos indicus" ~ "Bovidae", #cow
    Barcode_region == "16" & Best_match_references %in% "Canis rufus, Canis lupus, Canis latrans" ~ "Canidae", # coyote
    Barcode_region == "16" & Best_match_references %in% "Cavenderia subdiscoidea, Cavenderia pseudoaureostipes" ~ "Cavenderiaceae", #slime mold
    Barcode_region == "16" & Best_match_references %in% "Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Chloropicaceae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Chloropicon sieburthii, Chloropicon roscoffensis, Chloropicon primus, Chloropicon laureae" ~ "Chloropicaceae", #green algae
    Barcode_region == "16" & Best_match_references %in% "Creophilus maxillosus, Anthrenus scrophulariae" ~ "Dermestidae", # Common carpet beetle/Buffalo carpet beetle
    Barcode_region == "16" & Best_match_references %in% "Crotalus viridis, Crotalus oreganus" ~ "Viperidae", #western rattlesnake
    Barcode_region == "16" & Best_match_references %in% "Gallus sonneratii, Gallus gallus" ~ "Phasianidae", # wild turkey probably, or chicken
    Barcode_region == "16" & Best_match_references %in% "Heliconoides inflatus, Cathaica polystigma, Nesiohelix moreletiana, Karaftohelix maackii, Atlanta turriculata, Aceria guerreronis, Hahnia ngai, Grylloprimevala jilina, Aconurella prolixa, Processina dashahensis, Tapinoma indicum, Homo sapiens" ~ "Heliconoididae", # Heliconoides inflatus -- Sea butterfly (range is not Socal, but Mexico, and it's the closest species in the list) 
    Barcode_region == "16" & Best_match_references %in% "Johnmannia powerae, Paratriodonta lentula" ~ "Scarabaeidae", #Scarabaeidae (scarab)
    Barcode_region == "16" & Best_match_references %in% "Meleagris gallopavo, Labidesthes sicculus" ~ "Phasianidae", #wild turkey
    Barcode_region == "16" & Best_match_references %in% "Merluccius productus, Merluccius angustimanus" ~ "MerlMerlucciidaeuccius", #North Pacific Hake
    Barcode_region == "16" & Best_match_references %in% "Odocoileus virginianus, Odocoileus hemionus" ~ "Cervidae", #mule deer
    Barcode_region == "16" & Best_match_references %in% "Paralucilia pseudolyrcea, Paralucilia paraensis, Hemilucilia semidiaphana, Calliphora vomitoria, Calliphora vicina, Calliphora sinensis, Calliphora nigribarbis, Calliphora calliphoroides" ~ "Tachinidae", # fly -- choosing Paralucilia
    Barcode_region == "16" & Best_match_references %in% "Ravinia vagabunda, Ravinia stimulans, Ravinia derelicta" ~ "Sarcophagidae", #flesh fly
    Barcode_region == "16" & Best_match_references %in% "Scomber japonicus, Scomber colias, Scomber australasicus"~ "Scombridae", #pacific mackerel
    Barcode_region == "16" & Best_match_references %in% "Sus scrofa, Thryonomys swinderianus" ~ "Suidae", #wild pig
    Barcode_region == "16" & Best_match_references %in% "Tachycines zorzini, Onthophagus obscurior" ~ "Scarabaeidae", #dung beetle
    Barcode_region == "16" & Best_match_references %in% "Sebastes serranoides, Sebastes melanops, Sebastes flavidus" ~ "Scorpaenidae",  #rockfish
    Barcode_region == "16" & Best_match_references %in% "" ~ "no match",
    TRUE ~ family
    ))


 

checkna <- gaviota[is.na(gaviota$family), ] 
```




# label predator
```{r}
gaviota$host = NA

gaviota <- gaviota %>%
  mutate(host = case_when(
    common_name %in% c("Bobcat", "Coyote", "Wild boar", "Mule deer", "Puma") ~ common_name,
    TRUE ~ "no"  # Assign "no" if not in the specified list
  ))


```

```{r}
ggplot(gaviota, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Common Names", x = "Common Name", y = "Count") +
  theme_minimal()
```
```{r}
ggplot(gaviota, aes(x = host)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Common Names", x = "Common Name", y = "Count") +
  theme_minimal()
```

```{r}
# add site names
gaviota <- gaviota %>%
  mutate(
    site = case_when(
      substr(Sample, 1, 1) == "D" ~ "Dangermond",
      substr(Sample, 1, 1) == "C" ~ "Coal Oil Point",
      substr(Sample, 1, 1) == "N" ~ "North Campus Open Space",
      substr(Sample, 1, 1) == "H" ~ "Hollister",
      TRUE ~ "Unknown Site"  # Default for unmatched cases
    )
  )

dangermond_coyote_samples <- gaviota %>% 
  filter(common_name == "Coyote", 
         site == "Dangermond") 

dangermond_coyote_list = list(dangermond_coyote_samples$Sample)

dangermond_bobcat_samples <- gaviota %>% 
  filter(common_name == "Bobcat", 
         site == "Dangermond")

dangermond_bobcat_list <- list(dangermond_bobcat_samples$Sample)

dangermond_boar_samples <- gaviota %>% 
  filter(common_name == "Wild boar", 
         site == "Dangermond") 

dangermond_boar_list <- list(dangermond_boar_samples$Sample)

dangermond_deer_samples <- gaviota %>% 
  filter(common_name == "Mule deer", 
         site == "Dangermond")

dangermond_deer_list <- list(dangermond_deer_samples$Sample)

dangermond_puma_samples <- gaviota %>% 
  filter(common_name == "Puma", 
         site == "Dangermond")

dangermond_puma_list <- list(dangermond_puma_samples$Sample)


cop_coyote_samples <- gaviota %>% 
  filter(common_name == "Coyote", 
         site == "Coal Oil Point")

cop_coyote_list = list(cop_coyote_samples$Sample)

ncos_coyote_samples <- gaviota %>% 
  filter(common_name == "Coyote", 
         site == "North Campus Open Space")

ncos_coyote_list = list(ncos_coyote_samples$Sample)

hollister_coyote_samples <- gaviota %>% 
  filter(common_name == "Coyote",
         site == "Hollister")

hollister_coyote_list = list(hollister_coyote_samples$Sample)

hollister_boar_samples <- gaviota %>% 
  filter(common_name == "Wild boar",
         site == "Hollister")

hollister_boar_list = list(hollister_boar_samples$Sample)


```







```{r}
#Dangermond
dangermond_coyote <- gaviota[gaviota$Sample %in% c("DEC1", "DEC3", "DEC4", "DEC5", "DEC6", "DEC7", "DEC8", "DEC10", "DEC11", "DEC12", "DEC13", "DAB1", "DABC2", "DABC3", "DABC4", "DABC5", "DABC6", "DABC7", "DABF1", "DAC1", "DAC2", "DAC3", "DAD1", "DADC2", "DADP1", "DAF1", "DAF2", "DAFB1", "DAP1", "DBC2", "DCC1", "DCC3", "DCC5", "DCC6", "DCC7", "DCC8", "DCF1", "DDB1", "DDB2", "DDB3", "DDC10", "DDC2", "DDC3", "DDC4", "DDC5", "DDC5", "DDC6", "DDC7", "DDC8", "DDC9", "DDC9", "DEB1", "DEB2", "DEB3", "DEF1", "DEP2", "DFB1", "DFB2", "DFB3", "DGB5", "DGB6", "DGC11", "DGC2", "DGC3", "DGC4", "DGC5", "DGC6", "DGC7", "DGF1", "DGP1", "DJB2", "DJC3", "DJC4", "DJC5", "DJC6", "DJC7", "DOC1", "DSB1", "DSC1", "DSC2", "DUC1", "DZB6", "DZC2", "DDB4", "DDC1", "DDC12", "DDC13", "DDC15", "DDK2", "DDP2", "DJC1", "DLB2", "DOC2"),]

dangermond_bobcat <- gaviota[gaviota$Sample %in% c("DABB1", "DABB2", "DABB3", "DABB4", "DABC1", "DAFB2", "DFB4" , "DGB1", "DGB10", "DGB2", "DGB3" , "DGB4", "DGB7", "DGB8", "DGB9", "DGC1", "DGC10", "DGC12", "DJB1", "DJB3", "DJB4", "DOB1", "DQB1", "DSB2", "DUB1", "DWB1", "DWB1", "DWB2", "DWP1", "DZB1", "DZB2", "DZB3", "DZB4", "DZB5", "DZC1", "DZP1"),] 

dangermond_boar <- gaviota[gaviota$Sample %in% c("DABH1", "DABH2", "DABH3", "DAFH1", "DAGH1", "DAHH1", "DAHH2", "DAHH2" , "DAHH3", "DAHH4", "DAHH5", "DAHH6", "DAHH7" , "DAIH1", "DAIH2", "DAIH3", "DAIH5", "DAIH6", "DAJH1", "DAJH2", "DAKH1", "DAKH2", "DAKH3", "DAKH4","DALH1", "DALH2", "DALH3", "DALH4", "DALH5", "DALH6", "DALH7", "DALH8", "DAMH1", "DAMH3", "DANH1", "DAOH1", "DAOH2", "DAQH1", "DAQH2", "DFK1", "DFU1", "DLU1", "DLU2", "DQC2", "DQU1", "DTU1", "DTU2", "DTU3", "DUKH1", "DUKH10", "DUKH11", "DUKH12", "DUKH13", "DUKH14", "DUKH2", "DUKH3", "DUKH4", "DUKH6", "DUKH7", "DUKH8", "DZB7", "DLU3"),] 


dangermond_deer <- gaviota[gaviota$Sample %in% c("DJD1", "DJD2", "DLD1", "DLD2", "DLD3", "DLD4", "DZD1", "DZH1"),] 

dangermond_puma <- gaviota[gaviota$Sample %in% c("DJC9"),]


cop_coyote <- gaviota[gaviota$Sample %in% c("CABC1", "CAEC2", "CAEC3", "CBF1", "CHC1", "CHC2", "CIB1", "CIC1", "CKC1", "CKC2", "CKC3", "CKC4", "CKC5", "CMC1", "CMC2", "CNC1", "CSC1", "CYC1", "CYC2"),]


ncos_coyote <- gaviota[gaviota$Sample %in% c("NAAC1", "NABC2", "NABC2", "NACC1", "NHC1", "NHC2", "NKC1", "NMC1", "NMC2", "NRC1", "NRC2", "NXC1", "NXC2", "NYC1", "NPB1"),]


hollister_coyote <- gaviota[gaviota$Sample %in% c("HVC1"),] 

hollister_boar <- gaviota[gaviota$Sample %in% c("HVH1", "HVH2"),]
```

plots
```{r}
# dangermond coyote
dangermond_coyote_plot <- ggplot(dangermond_coyote, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# cop coyote
cop_coyote_plot <- ggplot(cop_coyote, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Coal Oil Point Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip()+
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# ncos coyote
ncos_coyote_plot <- ggplot(ncos_coyote, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "NCOS Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip()+
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond bobcat
dangermond_bobcat_plot <- ggplot(dangermond_bobcat, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Bobcat Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond boar
dangermond_boar_plot <- ggplot(dangermond_boar, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Wild Boar Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond deer
dangermond_deer_plot <- ggplot(dangermond_deer, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Deer Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond puma
dangermond_puma_plot <- ggplot(dangermond_puma, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Puma Diet Items", x = "Common Name", y = "Count") +
  theme_minimal()+
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# hollister coyote
hollister_coyote_plot <- ggplot(hollister_coyote, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Hollister Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal()+
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

#hollister boar
hollister_boar_plot <- ggplot(hollister_boar, aes(x = common_name)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Hollister Wild Boar Diet Items", x = "Common Name", y = "Count") +
  theme_minimal()+
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)
```

plots
```{r}
dangermond_coyote_plot

dangermond_bobcat_plot

dangermond_boar_plot

dangermond_deer_plot

dangermond_puma_plot

cop_coyote_plot

ncos_coyote_plot

hollister_coyote_plot

hollister_boar_plot
```


broadtax plots
```{r}
# dangermond coyote
dangermond_coyote_plot <- ggplot(dangermond_coyote, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# cop coyote
cop_coyote_plot <- ggplot(cop_coyote, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Coal Oil Point Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip()+
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# ncos coyote
ncos_coyote_plot <- ggplot(ncos_coyote, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "NCOS Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip()+
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond bobcat
dangermond_bobcat_plot <- ggplot(dangermond_bobcat, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Bobcat Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond boar
dangermond_boar_plot <- ggplot(dangermond_boar, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Wild Boar Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond deer
dangermond_deer_plot <- ggplot(dangermond_deer, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Deer Diet Items", x = "Common Name", y = "Count") +
  theme_minimal() +
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# dangermond puma
dangermond_puma_plot <- ggplot(dangermond_puma, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Dangermond Puma Diet Items", x = "Common Name", y = "Count") +
  theme_minimal()+
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

# hollister coyote
hollister_coyote_plot <- ggplot(hollister_coyote, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Hollister Coyote Diet Items", x = "Common Name", y = "Count") +
  theme_minimal()+
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)

#hollister boar
hollister_boar_plot <- ggplot(hollister_boar, aes(x = broadtax)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Hollister Wild Boar Diet Items", x = "Common Name", y = "Count") +
  theme_minimal()+
  coord_flip() +
  geom_text(stat='count', aes(label=..count..), hjust=-0.5)
```

plots
```{r}
dangermond_coyote_plot

dangermond_bobcat_plot

dangermond_boar_plot

dangermond_deer_plot

dangermond_puma_plot

cop_coyote_plot

ncos_coyote_plot

hollister_coyote_plot

hollister_boar_plot
```



```{r}
unique(ncos_coyote$Sample)
```


make a column for what I thought the species was when I collected it.
```{r}
#put sample id's into gaviota
scat_and_lab_ids <- read_csv(here("data", "IDs.scat.lab.correcteddateformat.csv"))

colnames(gaviota)[1] ="LabID"

gaviota <- gaviota %>% left_join(scat_and_lab_ids, by = 'LabID', relationship = 'many-to-many') 

guessed_sp = NA

gaviota <- gaviota |>
  mutate(
    sp = substr(ScatID, 12, 12),
    guessed_sp = case_when(
      sp == "B" ~ "bobcat",
      sp == "C" ~ "coyote",
      sp == "F" ~ "fox",
      sp == "U" ~ "bear",
      sp == "P" ~ "puma",
      sp == "D" ~ "deer",
      sp == "H" ~ "wild pig",
      sp == "u" ~ "unknown",
      TRUE ~ NA_character_
    )
  )
```

make gaviota a csv

```{r}
#write_csv(gaviota, "gaviota.csv")
```



DABC1 has both bobcat mule deer in it, but was guessed as coyote
DAHH2 has 2 wild boar

```{r}
# Convert common_name to lowercase
gaviota$common_name <- tolower(gaviota$common_name)
gaviota$guessed_sp <- tolower(gaviota$guessed_sp)

gaviota <- gaviota %>% 
  mutate(target = case_when(common_name %in% c("coyote", "bobcat", "mule deer", "wild boar", "puma") ~ "target",
        TRUE ~ "No"))

target <- gaviota %>% 
  filter(target == "target") %>% 
  select(LabID, ScatID, curated_species, common_name, guessed_sp, target)

match <- target %>% 
  filter(common_name == guessed_sp)

match_count <- sum(target$common_name == target$guessed_sp, na.rm = TRUE)
mismatch_count <- sum(target$common_name != target$guessed_sp, na.rm = TRUE)

# List differences
differences <- gaviota[gaviota$common_name != gaviota$guessed_sp, c("common_name", "guessed_sp")]
same <- gaviota[gaviota$common_name == gaviota$guessed_sp, c("common_name", "guessed_sp")]

# Print results
cat("Matches:", match_count, "\n")
cat("Mismatches:", mismatch_count, "\n")
print(differences)
```



make gaviota a csv

```{r}
write_csv(gaviota, "gaviota_taxonomy.csv")
```

