---
title: "Model Selections"
author: 
date: "2025-09-08"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(betareg)
library(MuMIn)
library(knitr)
library(kableExtra)
library(here)
library(pROC)
library(flextable)
library(patchwork)
```


Import data

```{r, results = 'hide', message = F}
# dangermond samples
gav_dangermond_prop <- read_csv(here("metabarcoding_scripts_GL","outputs", "gav_dangermond_dist_marine_prop.csv"))

# Coal Oil Point and NCOS samples
gav_cop_ncos_prop <- read_csv(here("metabarcoding_scripts_GL","outputs", "gav_cop_ncos_dist_marine_prop.csv"))
```

# Beta Regression Dangermond
## Response variable: Proportion of Marine diet items

Transform prop_marine to avoid exact 0 or 1

Take out bobcat and puma samples, as they don't have any marine prey in them.
```{r}
epsilon <- 1e-4
gav_dangermond_prop <- gav_dangermond_prop |> 
  mutate(prop_marine_beta = pmin(pmax(prop_marine, epsilon), 1 - epsilon)) |> 
  filter(host == 'coyote') |> 
  filter(guadelupe_fur_seal != 1)
```



2. Fit beta regression candidate models
```{r}
beta0null <- betareg(prop_marine_beta ~ 1, data = gav_dangermond_prop)  # intercept-only
beta1coast <- betareg(prop_marine_beta ~ dist_to_coast_m, data = gav_dangermond_prop)
#beta2predator <- betareg(prop_marine_beta ~ host, data = gav_dangermond_prop)
#beta3additive <- betareg(prop_marine_beta ~ dist_to_coast_m + host, data = gav_dangermond_prop)
#beta4interaction <- betareg(prop_marine_beta ~ dist_to_coast_m * host, data = gav_dangermond_prop)
```

3. Model selection using model.sel()
```{r}
#model_sel <- model.sel(beta0null, beta1coast, beta2predator, beta3additive, beta4interaction, rank = "BIC")

#with only 0 and 1 model 
model_sel <- model.sel(beta0null, beta1coast, rank = "BIC")

model_sel_AIC <- model.sel(beta0null, beta1coast, rank = "AIC")

model_sel
```


## add equation and biological explanation to model selection table

Convert to data.frame
```{r}
model_sel_df <- as.data.frame(model_sel)
model_sel_df$Model <- rownames(model_sel_df)

# model_sel_df <- model_sel_df |> 
#   rename(DistCoast = dist_to_coast_m,
#          `DistCoast:host` = `dist_to_coast_m:host`)
```


Add LaTeX-style equations
```{r}
model_sel_df <- model_sel_df |> 
  mutate(Equation = case_when(
    Model == "beta0null" ~ "\\(\\mathrm{MarineProp} = \\beta_0\\)",
    Model == "beta1coast" ~ "\\(\\mathrm{MarineProp} = \\beta_0 + \\beta_1 \\cdot dist\\_coast\\)"))

# model_sel_df <- model_sel_df |> 
#   mutate(Equation = case_when(
#     Model == "beta0null" ~ "\\(\\mathrm{MarineProp} = \\beta_0\\)",
#     Model == "beta1coast" ~ "\\(\\mathrm{MarineProp} = \\beta_0 + \\beta_1 \\cdot dist\\_coast\\)",
#     Model == "beta2predator" ~ "\\(\\mathrm{MarineProp} = \\beta_0 + \\beta_2 \\cdot predator\\)",
#     Model == "beta3additive" ~ "\\(\\mathrm{MarineProp} = \\beta_0 + \\beta_1 \\cdot dist\\_coast + \\beta_2 \\cdot predator\\)",
#     Model == "beta4interaction" ~ "\\(\\mathrm{MarineProp} = \\beta_0 + \\beta_1 \\cdot dist\\_coast + \\beta_2 \\cdot predator + \\beta_3 \\cdot dist\\_coast \\cdot predator\\)"
#   ))



```

Add biological explanation
```{r}
model_sel_df <- model_sel_df |> 
  mutate(Explanation = case_when(
  Model == "beta0null" ~ "The proportion of marine prey is constant across all distances from the coastline",
  Model == "beta1coast" ~ "Proportion of marine prey is spatially structured"))

# model_sel_df <- model_sel_df |> 
#   mutate(Explanation = case_when(
#   Model == "beta0null" ~ "The proportion of marine prey is constant across all predators and distances from the coastline",
#   Model == "beta1coast" ~ "Proportion of marine prey is spatially structured, not predator specific",
#   Model == "beta2predator" ~ "Feeding behavior on marine prey differs among predators, but is independent of location relative to coastline",
#   Model == "beta3additive" ~ "The effect of distance to coast on the proportion of marine prey in a predator's diet is the same across predator species",
#   Model == "beta4interaction" ~ "The proportion of marine prey in diet depends on both predator identity and distance to the coast"
# ))
```



 --- Step 3: Select and reorder columns ---
```{r}

model_sel_df <- model_sel_df %>% 
  mutate(
    Equation = case_when(
      Model == "beta0null"  ~ "MarineProp = β₀",
      Model == "beta1coast" ~ "MarineProp = β₀ + β₁ × dist to coast(m)"
    )
  )

plot_df <- model_sel_df %>%
  select(Model, Equation, Explanation, delta) %>%
  arrange(delta) %>%
  rename(`ΔBIC` = delta) %>% # took out , BICWeight = weight
  mutate(Model = case_when(
    Model == "beta0null" ~ "Null",
    Model == "beta1coast" ~ "Coast",
    TRUE ~ Model
  ))

plot_df <- plot_df %>%
  mutate(
    #BIC = round(BIC, 2),
    `ΔBIC` = round(`ΔBIC`, 2)
  )

# plot_df <- model_sel_df %>%
#   select(Model, Equation, Explanation, BIC, delta, weight) %>%
#   arrange(BIC) %>%
#   rename(DeltaBIC = delta, BICWeight = weight) %>%
#   mutate(Model = case_when(
#     Model == "beta0null" ~ "Null",
#     Model == "beta1coast" ~ "Coast",
#     Model == "beta2predator" ~ "Predator",
#     Model == "beta3additive" ~ "Additive",
#     Model == "beta4interaction" ~ "Interaction",
#     TRUE ~ Model
#   ))




```

create a flextable that I can output to word
```{r}
beta_ft <- flextable(plot_df) %>%
  autofit() %>%                        # auto-adjust column widths
  theme_box() %>%                       # boxed style
  bold(part = "header") %>%             # bold header
  set_caption("Beta Regression Model selection table with ΔBIC") %>%
  colformat_num(j = c("ΔBIC"), digits = 2) # format numbers

beta_ft

```



Step 4: Make LaTeX kable table
```{r, echo=T, results='asis'}
kable(plot_df, format = "latex", booktabs = TRUE, escape = FALSE, row.names = FALSE,
      caption = "Beta Regression Model Selection Table - Dangermond") %>%
  kable_styling(latex_options = c("HOLD_position", "striped")) %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3, width = "3cm")
```

```{r}
# Compare models with BIC
#BIC(beta0null, beta1coast)

# -----------------------------
# 2. Prediction grid
# -----------------------------
predgrid_beta <- data.frame(
  dist_to_coast_m = seq(
    min(gav_dangermond_prop$dist_to_coast_m),
    max(gav_dangermond_prop$dist_to_coast_m),
    length.out = 200
  )
)

# Predicted proportions
predgrid_beta$pred_coast <- predict(beta1coast, newdata = predgrid_beta, type = "response")
predgrid_beta$pred_null <- predict(beta0null, newdata = predgrid_beta, type = "response")

# Convert to long format for ggplot
pred_long_beta <- predgrid_beta %>%
  pivot_longer(
    cols = starts_with("pred_"),
    names_to = "model",
    values_to = "prob"
  ) %>%
  mutate(
    model = recode(model,
                   "pred_null" = "Constant-only model",
                   "pred_coast" = "Distance to coast"
                   )
  )

# Create a named vector for colors matching models
model_colors <- c("beta0null" = "red", "beta1coast" = "blue")

# Plot with delta BIC annotations
beta_plot<- ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = prop_marine_beta)) +
  geom_jitter(height = 0, width = 0, alpha = 0.3, color = "gray30") +
  geom_line(data = pred_long_beta, aes(x = dist_to_coast_m, y = prob, color = model),
            linewidth = 1.2) +
  scale_color_manual(values = c("Distance to coast" = "blue",
                                "Constant-only model" = "red")) +
  labs(
    x = "Distance to coast (m)",
    y = "Proportion marine prey",
    color = "Beta Regression Model",
    title = "Beta regression: Modeled Marine Prey Proportion"
  ) +
  # # Annotate ΔBIC using model_sel_df
  # annotate("text",
  #          x = max(gav_dangermond_prop$dist_to_coast_m),
  #          y = 0.95,
  #          hjust = 1,
  #          label = paste0("Null model ΔBIC = ", round(model_sel_df$delta[model_sel_df$Model == "beta0null"], 2)),
  #          size = 4.5,
  #          color = model_colors["beta0null"]) +
  # annotate("text",
  #          x = max(gav_dangermond_prop$dist_to_coast_m),
  #          y = 0.88,
  #          hjust = 1,
  #          label = paste0("Distance to coast model ΔBIC = ", round(model_sel_df$delta[model_sel_df$Model == "beta1coast"], 2)),
  #          size = 4.5,
  #          color = model_colors["beta1coast"]) +
  theme_minimal(base_size = 14)

beta_plot

ggsave(filename = here("figs", "beta_plot.png"), plot = beta_plot, dpi = 600, units = "in", width = 8, height = 6)

```


The best fitting model is the null model, which means that the proportion of marine prey is constant across all predators and distances from the coastline. However, the model that is second best fit with a small delta BIC is the model that only takes into account the predator species. In this model, feeding behavior or marine prey differs among predators, but is independent of the location relative to the coastline. 


**************************************************

# Logistic Regression - Dangermond
## Response variable: Presence of Marine Diet Items (1 = at least one marine diet item, 0 = terrestrial diet items only)

```{r}
gav_dangermond_prop <- gav_dangermond_prop |> 
  mutate(marine_present = case_when(
    marine_species_count > 0 ~ 1,
    marine_species_count == 0 ~ 0
  ))
```


Filter out puma and bobcat predators (no marine prey)
Don't need to transform response variable to avoid exact 1 or 0 for logistic regression.
```{r}
#epsilon <- 1e-4
gav_dangermond_prop <- gav_dangermond_prop |> 
  #mutate(marine_present_beta = pmin(pmax(marine_present, epsilon), 1 - epsilon)) |> 
  filter(host!="puma") |> 
  filter(host != "bobcat") |> 
  filter(host != "wild boar")
```



2. Fit logistic regression candidate models
```{r}
glm0null <- glm(marine_present ~ 1, family = binomial, data = gav_dangermond_prop)  # intercept-only
glm1coast <- glm(marine_present ~ dist_to_coast_m, family = binomial, data = gav_dangermond_prop)
#glm2predator <- glm(marine_present ~ host, family = binomial, data = gav_dangermond_prop)
#glm3additive <- glm(marine_present ~ dist_to_coast_m + host, family = binomial, data = gav_dangermond_prop)
#glm4interaction <- glm(marine_present ~ dist_to_coast_m * host, family = binomial, data = gav_dangermond_prop)


# summary(glm2predator)
# 
# library(DHARMa)
# 
# simulateResiduals(glm2predator, plot = T)
# 
# plot()
# 
# plot(beta1coast)
# 
# plot(beta0null)
# 
# summary(beta2predator)
# 
# summary(beta1coast)
# 
# summary(beta0null)
# 
# summary(glm2predator)

#log likeliehoods compare to each other


```

```{r}
summary(glm1coast)
```


3. Model selection using model.sel()
```{r}
#model_sel <- model.sel(glm0null, glm1coast, glm2predator, glm3additive, glm4interaction, rank = "BIC")

#with only model 0 and 1
model_sel <- model.sel(glm0null, glm1coast, rank = "BIC")

model_sel_AIC <- model.sel(glm0null, glm1coast, rank = "AIC")


model_sel

summary(glm1coast)$coefficients

coef(glm1coast)
```


## add equation and biological explanation to model selection table

Convert to data.frame
```{r}
model_sel_df <- as.data.frame(model_sel)
model_sel_df$Model <- rownames(model_sel_df)

# model_sel_df <- model_sel_df |> 
#   rename(DistCoast = dist_to_coast_m,
#          `DistCoast:host` = `dist_to_coast_m:host`)
```


Add LaTeX-style equations
```{r}
model_sel_df <- model_sel_df |> 
  mutate(Equation = case_when(
    Model == "glm0null" ~ "\\(\\mathrm{MarinePres} = \\beta_0\\)",
    Model == "glm1coast" ~ "\\(\\mathrm{MarinePres} = \\beta_0 + \\beta_1 \\cdot dist\\_coast\\)"
  ))

# model_sel_df <- model_sel_df |> 
#   mutate(Equation = case_when(
#     Model == "glm0null" ~ "\\(\\mathrm{MarinePres} = \\beta_0\\)",
#     Model == "glm1coast" ~ "\\(\\mathrm{MarinePres} = \\beta_0 + \\beta_1 \\cdot dist\\_coast\\)",
#     Model == "glm2predator" ~ "\\(\\mathrm{MarinePres} = \\beta_0 + \\beta_2 \\cdot predator\\)",
#     Model == "glm3additive" ~ "\\(\\mathrm{MarinePres} = \\beta_0 + \\beta_1 \\cdot dist\\_coast + \\beta_2 \\cdot predator\\)",
#     Model == "glm4interaction" ~ "\\(\\mathrm{MarinePres} = \\beta_0 + \\beta_1 \\cdot dist\\_coast + \\beta_2 \\cdot predator + \\beta_3 \\cdot dist\\_coast \\cdot predator\\)"
#   ))

```

Add biological explanation
```{r}
model_sel_df <- model_sel_df |> 
  mutate(Explanation = case_when(
  Model == "glm0null" ~ "The presence of marine prey is constant across all distances from the coastline",
  Model == "glm1coast" ~ "Presence of marine prey is spatially structured"
))

# model_sel_df <- model_sel_df |> 
#   mutate(Explanation = case_when(
#   Model == "glm0null" ~ "The presence of marine prey is constant across all predators and distances from the coastline",
#   Model == "glm1coast" ~ "Presence of marine prey is spatially structured, not predator specific",
#   Model == "glm2predator" ~ "Feeding behavior on marine prey differs among predators, but is independent of location relative to coastline",
#   Model == "glm3additive" ~ "The effect of distance to coast on the presence of marine prey in a predator's diet is the same across predator species",
#   Model == "glm4interaction" ~ "The presence of marine prey in diet depends on both predator identity and distance to the coast"
# ))
```

 --- Step 3: Select and reorder columns ---
```{r}
plot_df <- model_sel_df %>%
  select(Model, Equation, Explanation, delta) %>% # took out BICWeight
  arrange(delta) %>%
  rename(`ΔBIC` = delta) %>%
  mutate(Model = case_when(
    Model == "beta0null" ~ "Null",
    Model == "beta1coast" ~ "Coast",
    TRUE ~ Model
  ))



# plot_df <- model_sel_df %>%
#   select(Model, Equation, Explanation, BIC, delta) %>% # took out BICWeight
#   arrange(BIC) %>%
#   rename(DeltaBIC = delta) %>%
#   mutate(Model = case_when(
#     Model == "beta0null" ~ "Null",
#     Model == "beta1coast" ~ "Coast",
#     Model == "beta2predator" ~ "Predator",
#     Model == "beta3additive" ~ "Additive",
#     Model == "beta4interaction" ~ "Interaction",
#     TRUE ~ Model
#   ))



```

```{r}
# model_sel_df <- model_sel_df %>% 
#   mutate(
#     Equation = case_when(
#       Model == "glm0null"  ~ "MarineProp = β₀",
#       Model == "glm1coast" ~ "MarineProp = β₀ + β₁ × dist to coast(m)"
#     )
#   )

model_sel_df <- model_sel_df %>% 
  mutate(
    Equation = case_when(
      Model == "glm0null"  ~ "p = 1 / (1 + e^(-β₀))",
      Model == "glm1coast" ~ "p = 1 / (1 + e^(-(β₀ + β₁ × Distance to coast (m))))"
    )
  )


plot_df <- model_sel_df %>%
  select(Model, Equation, Explanation, delta) %>% # took out BICWeight
  arrange(delta) %>%
  rename(`ΔBIC` = delta) %>%
  mutate(Model = case_when(
    Model == "glm0null" ~ "Null",
    Model == "glm1coast" ~ "Coast",
    TRUE ~ Model
  ))


log_ft <- flextable(plot_df) %>%
  autofit() %>%                        # auto-adjust column widths
  theme_box() %>%                       # boxed style
  bold(part = "header") %>%             # bold header
  set_caption("Logistic Regression Model selection table with ΔBIC and BIC") %>%
  colformat_num(j = c("`ΔBIC`"), digits = 2) # format numbers

log_ft
```


Step 4: Make LaTeX kable table
```{r}
kable(plot_df, format = "latex", booktabs = TRUE, escape = FALSE, row.names = FALSE,
      caption = "Logistic Regression Model Selection Table - Dangermond") %>%
  kable_styling(latex_options = c("HOLD_position", "striped")) %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3, width = "3cm")
```

The best fitting logistic regression model is the one modeling distance to coast only. In this model, presence of marine prey is spatially structured, but is not predator specific.









# Plot probility curve and solve for distances at various percentages of marine detections. Using the best fitting logistic regression, *glm1coast*,

Solve for distance where probability = 0.5

```{r}
coefs <- coef(glm1coast)
intercept <- coefs[1]
slope <- coefs[2]

dist_50 <- -intercept / slope
dist_50

# Extract coefficients
b0 <- coef(glm1coast)[1]
b1 <- coef(glm1coast)[2]

# Solve for distance at 75% probability
p_target <- 0.75
dist_75 <- (log(p_target / (1 - p_target)) - b0) / b1
dist_75


# Function: distance at probability p
get_distance <- function(p, model) {
  stopifnot(p > 0 & p < 1)  # probability must be between 0 and 1
  
  coefs <- coef(model)
  intercept <- coefs[1]
  slope <- coefs[2]
  
  # logit of p
  #logit_p <- log(p / (1 - p))
  
  # solve for distance
  #dist <- (logit_p - intercept) / slope
  
  dist <- -(((log((1 - p)/p)) + intercept) / slope)
  return(dist)
}
                                                                                                                                                                                    get_distance(.501, glm1coast)
                                                                                                 dist_25 <- get_distance(.25, glm1coast)                                                                                   
                                                                                                                                                                                    
                                                                                                                                                                                    
                                                                                              coefs <- coef(glm1coast)
  intercept <- coefs[1]
  slope <- coefs[2]

```





```{r}
# Make prediction grid
predgrid <- data.frame(dist_to_coast_m = seq(min(gav_dangermond_prop$dist_to_coast_m),
                                           max(gav_dangermond_prop$dist_to_coast_m),
                                           length.out = 200))

predgrid$pred <- predict(glm1coast, newdata = predgrid, type = "response")


#calculate AUC for coast model
# get fitted probabilities for the observed data
prob <- predict(glm1coast, type = "response")

# calculate AUC using observed response vs fitted prob
roc_obj <- roc(gav_dangermond_prop$marine_present, prob)
auc_val <- auc(roc_obj)


print(auc_val)


# get ROC data for ggplot
roc_df <- data.frame(
  specificity = rev(roc_obj$specificities),
  sensitivity = rev(roc_obj$sensitivities)
)

#calculate AUC for null model
# get fitted probabilities for the observed data
prob_null <- predict(glm0null, type = "response")

# calculate AUC using observed response vs fitted prob
roc_obj_null <- roc(gav_dangermond_prop$marine_present, prob_null)
auc_val_null <- auc(roc_obj_null)


print(auc_val_null)

# plot with label in top-right
ggplot(roc_df, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "blue", size = 1.2) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
  annotate("text", 
           x = 0.95, y = 0.05,   # top-right corner in ROC space
           label = paste0("AUC = ", round(auc(roc_obj), 3)),
           color = "black", size = 5) +
  labs(x = "1 - Specificity", y = "Sensitivity") +
  theme_minimal()

ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = marine_present)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.3) +
  geom_line(data = predgrid, aes(x = dist_to_coast_m, y = pred),
            color = "blue", size = 1.2) +
  labs(y = "Probability of marine detection",
       x = "Distance to coast (m)") +
    annotate("text", x = max(gav_dangermond_prop$dist_to_coast_m), 
           y = 1,  # top of y-axis
           label = "AUC = 0.6473", color = "black", hjust = 1, vjust = 1)



ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = marine_present)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.3) +
  geom_line(data = predgrid, aes(x = dist_to_coast_m, y = pred),
            color = "blue", size = 1.2) +
  geom_vline(xintercept = dist_50, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = dist_50, y = 0.5, label = "50% probability of finding sample with marine detection: 6.4m from coast", color = "red", vjust = -0.5, hjust = -.025) +
  labs(y = "Probability of marine detection",
       x = "Distance to coast (m)") +
  annotate("text", x = max(gav_dangermond_prop$dist_to_coast_m), 
           y = 1,  # top of y-axis
           label = "AUC = 0.6473", color = "black", hjust = 1, vjust = 1)

```
plot 75%

```{r}

ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = marine_present)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.3) +
  geom_line(data = predgrid, aes(x = dist_to_coast_m, y = pred),
            color = "blue", size = 1.2) +
  geom_vline(xintercept = dist_50, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = dist_50, y = 0.5, label = "50% probability of marine detection: 6.4m", color = "red", vjust = -0.5, hjust = -.025) +
  labs(y = "Probability of marine detection",
       x = "Distance to coast (m)") +
  geom_vline(xintercept = dist_75, linetype = "dashed", color = "red") +
  annotate("text", x = dist_75, y = 0.5, label = "75% probability \nof marine detection: \n-1715.691m", color = "red", vjust = -1.5, hjust = -.025)
  # labs(x = "Distance to coast (m)", 
  #      y = "Probability of marine detection") +
  # theme_minimal()

```


# Get distance within which 50% of ACTUAL marine detections occur (not modeled probability)

```{r}
# Subset only detections
marine_samples <- gav_dangermond_prop[gav_dangermond_prop$marine_present == 1, ]

# Median distance
median_distance <- median(marine_samples$dist_to_coast_m)
median_distance


# plot

ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = marine_present)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.3) +
  geom_line(data = predgrid, aes(x = dist_to_coast_m, y = pred),
            color = "blue", size = 1.2) +
  geom_vline(xintercept = median_distance, linetype = "dashed", color = "darkgreen", size = 1) +
  annotate("text", x = median_distance, y = 0.5, label = "50% of actual sampled detections", color = "darkgreen", vjust = -0.5, hjust = -.025)

```

plot predicted probabilities with probability distance and actual 50% detections distance
```{r}
ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = marine_present)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.3) +
  geom_line(data = predgrid, aes(x = dist_to_coast_m, y = pred),
            color = "blue", size = 1.2) +
  geom_vline(xintercept = median_distance, linetype = "dotted", color = "darkgreen", size = 1) +
  annotate("text", x = median_distance, y = 0.5, label = "50% (median) of actual \nsampled detections (258.6m)", color = "darkgreen", vjust = -0.5, hjust = -.025) +
  geom_vline(xintercept = dist_50, linetype = "dotdash", color = "red", size = 1) +
  annotate("text", x = dist_50, y = 0.5, label = "50% probability marine item detected (6.4m)", color = "red", vjust = -6.5, hjust = -0.025) +
  labs(y = "Probability of marine detection",
       x = "Distance to coast (m)",
       title = "Modeled Probability of Marine Detection") +
  geom_vline(xintercept = dist_25, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = dist_25, y = 0.5, label = "25% probability marine item detected (1728.5)", color = "red", vjust = 5, hjust = -.025) +
  theme_bw()

```

plot both models predictions on one plot
```{r}
# Create a prediction grid (for glm1coast, we need distance)
predgrid <- data.frame(
  dist_to_coast_m = seq(
    min(gav_dangermond_prop$dist_to_coast_m),
    max(gav_dangermond_prop$dist_to_coast_m),
    length.out = 200
  )
)

# Predicted probabilities for glm1coast (with slope)
predgrid$pred_coast <- predict(glm1coast, newdata = predgrid, type = "response")

# Predicted probabilities for glmnull0 (intercept only model)
# Same probability across all distances
predgrid$pred_null <- predict(glm0null, newdata = predgrid, type = "response")

# Convert to long format for easier plotting
pred_long <- predgrid |>
  tidyr::pivot_longer(
    cols = starts_with("pred_"),
    names_to = "model",
    values_to = "prob"
  ) |>
  dplyr::mutate(
    model = dplyr::recode(model,
      "pred_null"  = "Null model",
      "pred_coast" = "Distance to coast"
    )
  )

# Plot observed points + both model predictions
log_plot <- ggplot(gav_dangermond_prop, aes(x = dist_to_coast_m, y = marine_present)) +
  geom_jitter(height = 0.05, width = 0, alpha = 0.3, color = "gray30", size = 3) +
  geom_line(data = pred_long, aes(x = dist_to_coast_m, y = prob, color = model),
            linewidth = 1.2) +
  scale_color_manual(values = c("Distance to coast" = "blue",
                                "Null model" = "red")) +
  labs(
    x = "Distance to coast (m)",
    y = "Probability of marine detection",
    color = "Logistic Regression Model"
    #title = "Logistic Regresion: Modeled Marine Prey Detection"
  ) +
  annotate("text",
  x = max(gav_dangermond_prop$dist_to_coast_m),
  y = .4, hjust = 1, vjust = 1,
  label = paste0("Constant only model"),
  size = 4.5, color = "red") +
    annotate("text",
  x = max(gav_dangermond_prop$dist_to_coast_m),
  y = .20, hjust = 1, vjust = 1,
  label = paste0("Distance to coast model"),
  size = 4.5, color = "blue") +
 # annotate("text",
  #x = max(gav_dangermond_prop$dist_to_coast_m),
  #y = .93, hjust = 1, vjust = 1,
  #label = paste0("Null model \u0394BIC = ", round(model_sel$delta[1], 3)),
  #size = 4.5, color = "red") +
#annotate("text",
 # x = max(gav_dangermond_prop$dist_to_coast_m),
  #y = 0.86, hjust = 1, vjust = 1,
  #label = paste0("Distance to coast model \u0394BIC = ", round(model_sel$delta[2], 3)),
  #size = 4.5, color = "blue") +
theme_minimal(base_size = 14) +
theme(legend.position = "none")
# annotate("text",
#   x = max(gav_dangermond_prop$dist_to_coast_m),
#   y = .72, hjust = 1, vjust = 1,
#   label = paste0("Null model AUC = ", round(auc_val_null, 3)),
#   size = 4.5, color = "red") +
# annotate("text",
#   x = max(gav_dangermond_prop$dist_to_coast_m),
#   y = 0.64, hjust = 1, vjust = 1,
#   label = paste0("Distance to coast model AUC = ", round(auc_val, 3)),
#   size = 4.5, color = "blue") +
# theme_minimal(base_size = 14)

log_plot

ggsave(filename = here("figs", "log_plot.png"), plot = log_plot, dpi = 600, units = "in", width = 8, height = 6)

```

NOTE on second field site: 

I had planned to do a beta and logistic regression model selection with the samples at Coal Oil Point and North campus open space as well, but there were only 2 samples at this site that had marine prey in them. This is too small a sample size to do these models with.


```{r}
library(officer)

beta_doc <- read_docx() %>%
  body_add_par("Table 1: Beta Regression Model Results") %>%
  body_add_flextable(beta_ft)

print(beta_doc, target = here("figs", "beta_table.docx"))

#for logistic table
log_doc <- read_docx() %>%
  body_add_par("Table 1: Logistic Regression Model Results") %>%
  body_add_flextable(log_ft)

print(log_doc, target = here("figs", "log_table.docx"))

```

```{r}
both_figs <- log_plot / beta_plot +
  plot_annotation(
    tag_levels = "A",  # labels A, B
    tag_prefix = "",    # no extra character
    tag_suffix = "",
    theme = theme(
      plot.tag = element_text(face = "bold", size = 16)  # bold labels
    )
  )

both_figs

log_plot_tagged <- log_plot + labs(tag = "A") +
  theme(plot.tag = element_text(face = "bold", size = 16))

beta_plot_tagged <- beta_plot + labs(tag = "B") +
  theme(plot.tag = element_text(face = "bold", size = 16))

both_figs <- log_plot_tagged / beta_plot_tagged
both_figs

ggsave(filename = here("figs", "both_model_figs.png"), plot = both_figs, dpi = 600, units = "in", width = 8, height = 8)
```

