---
title: "Rarefaction curves"
author: 
date: "2025-10-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

packages
```{r}
library(tidyverse)
library(vegan)
library(here)
library(patchwork)
```


Import data
```{r}
gaviota_pres_abs <- read_csv(here("metabarcoding_scripts_GL", "outputs", "gaviota_pres_abs_host.csv"))
```

split up by host
```{r}
coyote <- gaviota_pres_abs |> 
  filter(host == "coyote")

bobcat <- gaviota_pres_abs |> 
  filter(host == "bobcat")

boar <- gaviota_pres_abs |> 
  filter(host == "wild boar")
```


convert to a matrix for vegan
```{r}
coyote_diet_mat <- coyote |> select(-sample_id, -host) %>% as.data.frame()

bobcat_diet_mat <- bobcat|> select(-sample_id, -host) %>% as.data.frame()

boar_diet_mat <- boar |> select(-sample_id, -host) %>% as.data.frame()
```


# Make Incidence Based Rarefaction Curves

how diet richness (species accumulation) increases as more scats are sampled (not within-sample accumulation)

compute sample-based rarefaction
```{r}
# for coyotes
coyote_accum <- specaccum(coyote_diet_mat, method = "random")

#bobcat
bobcat_accum <- specaccum(bobcat_diet_mat, method = "random")

#wild boar
boar_accum <- specaccum(boar_diet_mat, method = "random")


# Plot coyote (at both sites)
coyote_plot <- plot(coyote_accum, 
     ci.type = "polygon", 
     ci.col = "lightgreen", 
     col = "darkgreen", 
     lwd = 2, 
     ylab = "Cumulative diet richness",
     xlab = "Number of scats sampled")

#plot bobcat
bobcat_plot <- plot(bobcat_accum, 
     ci.type = "polygon", 
     ci.col = "lightblue", 
     col = "blue", 
     lwd = 2, 
     ylab = "Cumulative diet richness",
     xlab = "Number of scats sampled")

#plot boar
boar_plot <- plot(boar_accum, 
     ci.type = "polygon", 
     ci.col = "pink", 
     col = "red", 
     lwd = 2, 
     ylab = "Cumulative diet richness",
     xlab = "Number of scats sampled")

```
Convert to dataframe to plot ggplot object

```{r}

# Make a function to convert to dataframe for ggplot
accum_to_df <- function(accum_obj) {
  data.frame(
    sites = accum_obj$sites,
    richness = accum_obj$richness,
    sd = accum_obj$sd
  )
}


coyote_accum_df <- accum_to_df(coyote_accum)
bobcat_accum_df <- accum_to_df(bobcat_accum)
boar_accum_df <- accum_to_df(boar_accum)


coyote_plot <- ggplot(coyote_accum_df, aes(x = sites, y = richness)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = richness - sd, ymax = richness + sd),
              fill = "lightblue", alpha = 0.4) +
  labs(x = "Number of scat samples", y = "", title = "Coyote") +
  theme_bw()


bobcat_plot <- ggplot(bobcat_accum_df, aes(x = sites, y = richness)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = richness - sd, ymax = richness + sd),
              fill = "lightblue", alpha = 0.4) +
  labs(x = "Number of scat samples", y = "Diet richness", title = "Bobcat") +
  theme_bw()

boar_plot <- ggplot(boar_accum_df, aes(x = sites, y = richness)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = richness - sd, ymax = richness + sd),
              fill = "lightblue", alpha = 0.4) +
  labs(x = "Number of scat samples", y = "", title = "Wild Boar") +
  theme_bw()


# Combine them
species_richness_plots <- coyote_plot / bobcat_plot / boar_plot
species_richness_plots

```


Estimate Unobserved Diet Items (Using Chao2 for Incidence-Based data)
```{r}
#coyote
coyote_pool_est <- specpool(coyote_diet_mat)
coyote_pool_est

#calculate unobserved component
coyote_unobserved <- coyote_pool_est$chao - coyote_pool_est$Species
coyote_unobserved

#bobcat
bobcat_pool_est <- specpool(bobcat_diet_mat)
bobcat_pool_est

#calculate unobserved component
bobcat_unobserved <- bobcat_pool_est$chao - bobcat_pool_est$Species
bobcat_unobserved

#boar
boar_pool_est <- specpool(boar_diet_mat)
boar_pool_est

#calculate unobserved component
boar_unobserved <- boar_pool_est$chao - boar_pool_est$Species
boar_unobserved

```

# Sample Coverage (Using Good's Coverage for Incidence)

Good's Coverage:
Compute Good’s coverage (Ĉ):

[
\hat{C} = 1 - \frac{f_1}{N}
]

where

( f_1 ) = number of diet items found in exactly one scat (singletons)

( N ) = total number of incidences across all diet items

For incidence data, coverage can still be approximated using the frequency of unique occurrences across samples
```{r}
# how many scats each diet item occurred in

# coyote
coyote_item_freq <- colSums(coyote_diet_mat)
coyote_item_freq

# bobcat
bobcat_item_freq <- colSums(bobcat_diet_mat)
bobcat_item_freq

# boar
boar_item_freq <- colSums(boar_diet_mat)
boar_item_freq

#coverage function
coverage <- function(x) {
  f1 <- sum(x == 1)
  N <- sum(x)
  C <- 1 - (f1 / N)
  return(C)
}

#compute coverage

#coyote
coyote_coverage <- coverage(coyote_item_freq)

#bobcat
bobcat_coverage <- coverage(bobcat_item_freq)

#boar
boar_coverage <- coverage(boar_item_freq)

```

# Combine Results

```{r}
#coyote
coyote_results <- tibble(
  Observed_richness = coyote_pool_est$Species,
  Chao2_estimate = coyote_pool_est$chao,
  SE_Chao2 = coyote_pool_est$chao.se,
  Coverage = coyote_coverage
)

coyote_results

#bobcat
bobcat_results <- tibble(
  Observed_richness = bobcat_pool_est$Species,
  Chao2_estimate = bobcat_pool_est$chao,
  SE_Chao2 = bobcat_pool_est$chao.se,
  Coverage = bobcat_coverage
)

bobcat_results

#boar
boar_results <- tibble(
  Observed_richness = boar_pool_est$Species,
  Chao2_estimate = boar_pool_est$chao,
  SE_Chao2 = boar_pool_est$chao.se,
  Coverage = boar_coverage
)

boar_results
```

```{r}
# Combine Chao2 + Coverage results into a dataframe for annotation
summary_df <- tibble(
  host = c("Coyote", "Bobcat", "Wild Boar"),
  observed_richness = c(coyote_results$Observed_richness,
                        bobcat_results$Observed_richness,
                        boar_results$Observed_richness),
  chao2 = c(coyote_results$Chao2_estimate, 
            bobcat_results$Chao2_estimate, 
            boar_results$Chao2_estimate),
  se = c(coyote_results$SE_Chao2, 
         bobcat_results$SE_Chao2, 
         boar_results$SE_Chao2),
  coverage = c(coyote_results$Coverage, 
               bobcat_results$Coverage, 
               boar_results$Coverage)
)

summary_df <- summary_df %>%
  mutate(
    chao2_lower = chao2 - 1.96 * se,
    chao2_upper = chao2 + 1.96 * se,
    chao2_CI_label = sprintf("95%% CI: %.1f–%.1f", chao2_lower, chao2_upper)
  )

# Round for clean display
# summary_df <- summary_df %>%
#   mutate(
#     chao2_label = sprintf("Chao2 = %.1f ± %.1f", chao2, se),
#     coverage_label = sprintf("Coverage = %.1f%%", coverage * 100)
#   )



summary_df

```

extrapolate for species needed for chao2
```{r}
# Function to estimate samples needed to reach Chao2
samples_needed_to_chao <- function(accum_obj, chao_target) {
  # Current max richness
  max_richness <- max(accum_obj$richness)
  
  if(chao_target > max_richness) {
    # Linear extrapolation using last two points
    x0 <- accum_obj$sites[length(accum_obj$sites)-1]
    x1 <- accum_obj$sites[length(accum_obj$sites)]
    y0 <- accum_obj$richness[length(accum_obj$sites)-1]
    y1 <- accum_obj$richness[length(accum_obj$sites)]
    
    samples_needed <- x1 + (chao_target - y1) * (x1 - x0) / (y1 - y0)
  } else {
    # Find first sample where richness >= Chao2
    samples_needed <- accum_obj$sites[which(accum_obj$richness >= chao_target)[1]]
  }
  
  return(samples_needed)
}

# Apply to all three predators
coyote_samples_needed <- samples_needed_to_chao(coyote_accum, summary_df$chao2[1])
bobcat_samples_needed <- samples_needed_to_chao(bobcat_accum, summary_df$chao2[2])
boar_samples_needed   <- samples_needed_to_chao(boar_accum, summary_df$chao2[3])

# Print results
samples_needed <- tibble(
  host = c("Coyote", "Bobcat", "Wild Boar"),
  samples_needed = c(coyote_samples_needed,
                               bobcat_samples_needed,
                               boar_samples_needed)
)
```



```{r}
#function with samples needed
add_chao_coverage <- function(p, observed_richness, chao_est, se, coverage) {
  # Calculate 95% CI
  lower_CI <- chao_est - 1.96 * se
  upper_CI <- chao_est + 1.96 * se
  
  # Get x range from the plot
  x_range <- ggplot_build(p)$layout$panel_params[[1]]$x.range
  
  p +
    # Observed richness (blue dashed line)
    geom_hline(yintercept = observed_richness, linetype = "dashed", color = "blue") +
    
    # Observed richness label
    annotate("text", x = Inf, y = observed_richness + 1,
             label = sprintf("Observed = %.1f", observed_richness),
             hjust = 1.1, vjust = 0, size = 3.3, color = "blue") +
    
    # Chao2 95% CI shaded region
    annotate("rect",
             xmin = x_range[1], xmax = x_range[2],
             ymin = lower_CI, ymax = upper_CI,
             fill = "red", alpha = 0.15) +
    
    # Expected richness (red dashed line)
    geom_hline(yintercept = chao_est, linetype = "dashed", color = "red") +
    
    # Expected richness label
    annotate("text", x = Inf, y = upper_CI + 1,
             label = sprintf("Expected = %.1f (95%% CI: %.1f–%.1f)",
                             chao_est, lower_CI, upper_CI),
             hjust = 1.1, vjust = 0, size = 3.3, color = "red") +
    
    # Coverage text (black, at bottom)
    annotate("text", x = Inf, y = 1,
             label = sprintf("Coverage = %.1f%%", coverage * 100),
             hjust = 1.1, vjust = 0, size = 3.3, color = "black")
}


summary_full <- summary_df %>%
  left_join(samples_needed, by = "host")

coyote_plot_annot <- add_chao_coverage(
  p = coyote_plot,
  observed_richness = summary_full$observed_richness[summary_full$host == "Coyote"],
  chao_est = summary_full$chao2[summary_full$host == "Coyote"],
  se = summary_full$se[summary_full$host == "Coyote"],
  coverage = summary_full$coverage[summary_full$host == "Coyote"]
)

bobcat_plot_annot <- add_chao_coverage(
  p = bobcat_plot,
  observed_richness = summary_full$observed_richness[summary_full$host == "Bobcat"],
  chao_est = summary_full$chao2[summary_full$host == "Bobcat"],
  se = summary_full$se[summary_full$host == "Bobcat"],
  coverage = summary_full$coverage[summary_full$host == "Bobcat"]
)

boar_plot_annot <- add_chao_coverage(
  p = boar_plot,
  observed_richness = summary_full$observed_richness[summary_full$host == "Wild Boar"],
  chao_est = summary_full$chao2[summary_full$host == "Wild Boar"],
  se = summary_full$se[summary_full$host == "Wild Boar"],
  coverage = summary_full$coverage[summary_full$host == "Wild Boar"]
)

```

```{r}
# Example check
ggplot() +
  geom_line(aes(x = coyote_accum$sites, y = coyote_accum$richness)) +
  geom_point(aes(x = coyote_accum$sites, y = coyote_accum$richness)) +
  geom_vline(xintercept = 508, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 124.1, linetype = "dashed", color = "red") +
  labs(x = "Samples", y = "Prey richness", 
       title = "Extrapolated Chao2 Target for Coyote") +
  theme_minimal()
```



```{r}
# Function to add Chao2 + coverage info to plots
add_chao_coverage <- function(p, observed_richness, chao_est, se, coverage) {
  # Calculate 95% confidence interval
  lower_CI <- chao_est - 1.96 * se
  upper_CI <- chao_est + 1.96 * se
  
  # Get x range from the plot
  x_range <- ggplot_build(p)$layout$panel_params[[1]]$x.range
  
  p +
    # Observed richness line (blue, dashed)
    geom_hline(yintercept = observed_richness, linetype = "dashed", color = "blue") +
    annotate("text", x = Inf, y = observed_richness + 1,
             label = sprintf("Observed = %.1f", observed_richness),
             hjust = 1.1, vjust = 0, size = 3.3, color = "blue") +
    
    # Expected richness (Chao2) line (red, dashed)
    geom_hline(yintercept = chao_est, linetype = "dashed", color = "red") +
    
    # 95% CI shaded area for Chao2
    annotate("rect",
             xmin = x_range[1], xmax = x_range[2],
             ymin = lower_CI, ymax = upper_CI,
             fill = "red", alpha = 0.15) +
    
    # Chao2 label
    annotate("text", x = Inf, y = upper_CI + 1,
             label = sprintf("Expected Diet Richness = %.1f (95%% CI: %.1f–%.1f)", chao_est, lower_CI, upper_CI),
             hjust = 1.1, vjust = 0, size = 3.3, color = "red") +
    
    # Coverage label
    annotate("text", x = Inf, y = 1,
             label = sprintf("Coverage = %.1f%%", coverage * 100),
             hjust = 1.1, vjust = 0, size = 3.3, color = "black")
}



coyote_plot_annot <- add_chao_coverage(
  coyote_plot,
  summary_df$observed_richness[1],
  summary_df$chao2[1],
  summary_df$se[1],
  summary_df$coverage[1]
)

bobcat_plot_annot <- add_chao_coverage(
  bobcat_plot,
  summary_df$observed_richness[2],
  summary_df$chao2[2],
  summary_df$se[2],
  summary_df$coverage[2]
)

boar_plot_annot <- add_chao_coverage(
  boar_plot,
  summary_df$observed_richness[3],
  summary_df$chao2[3],
  summary_df$se[3],
  summary_df$coverage[3]
)



species_richness_plots_final <- coyote_plot_annot / bobcat_plot_annot / boar_plot_annot

ggsave(
  here("figs", "species_richness_rarefaction_with_chao2_CI.png"),
  species_richness_plots_final,
  width = 6,
  height = 10,
  dpi = 600
)


```

estimate the number of samples needed to reach the Chao2 estimated richness — sample-based rarefaction extrapolation.
```{r}
coyote_accum

# Number of scats needed to reach Chao2 estimated richness
chao_target <- summary_df$chao2[1]  # coyote Chao2

# Find first sample size where richness >= Chao2
samples_needed <- coyote_accum$sites[which(coyote_accum$richness >= chao_target)[1]]
samples_needed
```


```{r}
library(dplyr)

# create dataframe from specaccum
coyote_df <- data.frame(
  sites = coyote_accum$sites,
  richness = coyote_accum$richness
)

# find bracketing points
lower_idx <- max(which(coyote_df$richness < chao_target))
upper_idx <- min(which(coyote_df$richness >= chao_target))

# linear interpolation
x0 <- coyote_df$sites[lower_idx]
x1 <- coyote_df$sites[upper_idx]
y0 <- coyote_df$richness[lower_idx]
y1 <- coyote_df$richness[upper_idx]

samples_needed_interp <- x0 + (chao_target - y0) * (x1 - x0) / (y1 - y0)
samples_needed_interp

```


Have to extrapolate because my datasets don't reach the expected total richness.
This is just linear extrapolation between the last 2 points. -- this isn't very robust
```{r}
# Function to estimate samples needed to reach Chao2
samples_needed_to_chao <- function(accum_obj, chao_target) {
  # Current max richness
  max_richness <- max(accum_obj$richness)
  
  if(chao_target > max_richness) {
    # Linear extrapolation using last two points
    x0 <- accum_obj$sites[length(accum_obj$sites)-1]
    x1 <- accum_obj$sites[length(accum_obj$sites)]
    y0 <- accum_obj$richness[length(accum_obj$sites)-1]
    y1 <- accum_obj$richness[length(accum_obj$sites)]
    
    samples_needed <- x1 + (chao_target - y1) * (x1 - x0) / (y1 - y0)
  } else {
    # Find first sample where richness >= Chao2
    samples_needed <- accum_obj$sites[which(accum_obj$richness >= chao_target)[1]]
  }
  
  return(samples_needed)
}

# Apply to all three predators
coyote_samples_needed <- samples_needed_to_chao(coyote_accum, summary_df$chao2[1])
bobcat_samples_needed <- samples_needed_to_chao(bobcat_accum, summary_df$chao2[2])
boar_samples_needed   <- samples_needed_to_chao(boar_accum, summary_df$chao2[3])

# Print results
samples_needed <- tibble(
  Predator = c("Coyote", "Bobcat", "Wild Boar"),
  Samples_needed_for_Chao2 = c(coyote_samples_needed,
                               bobcat_samples_needed,
                               boar_samples_needed)
)

```

```{r}
coyote_plot_annot <- coyote_plot_annot 
  
  
bobcat_plot_annot 
  
  
boar_plot_annot
```









Use iNEXT for more robust extrapolation -- honestly couldn't get this to work properly....

```{r}
library(iNEXT)

iNEXT_list <- list(
  coyote = c(130, colSums(coyote_diet_mat)),
  bobcat = c(34, colSums(bobcat_diet_mat)),
  boar   = c(60, colSums(boar_diet_mat))
)

iNEXT_list <- list(
  Coyote = c(130, colSums(coyote_diet_mat)),
  Bobcat = c(34, colSums(bobcat_diet_mat)),
  "Wild Boar" = c(60, colSums(boar_diet_mat))
)


iNEXT_out <- iNEXT(iNEXT_list, q = 0, datatype = "incidence_freq")

iNEXT_out$DataInfo

coverage_perc <- iNEXT_out$DataInfo %>%
  mutate(SC_percent = SC * 100)

ggiNEXT(iNEXT_out) +
  theme_bw() +
  labs( x = "Number of Scat Samples",
        y = "Prey Species Diversity")

iNEXT_out$DataInfo
```


```{r}
# t = estimated number of samples needed to reach the target coverage of 90%
asymp <- estimateD(iNEXT_list, q = 0, datatype = "incidence_freq", base = "cover", level = 0.90)


# Suppose max scats for any predator = 131
level_vec <- seq(10, 150, by = 10)  # estimates every 10 scats

richness_by_size <- estimateD(
  iNEXT_list, 
  q = 0, 
  datatype = "incidence_freq",
  base = "size",
  level = level_vec
)

richness_by_size
```




```{r}


ggiNEXT(iNEXT_out) + 
  labs(#title = "Diet Richness",
       x = "Number of scat samples",
       y = "Diet richness") +
  theme_bw()

# Combine results
summary_df <- tibble(
  host = c("Coyote", "Bobcat", "Wild Boar"),
  observed_richness = c(coyote_results$Observed_richness,
                        bobcat_results$Observed_richness,
                        boar_results$Observed_richness),
  chao2 = c(coyote_results$Chao2_estimate, 
            bobcat_results$Chao2_estimate, 
            boar_results$Chao2_estimate),
  se = c(coyote_results$SE_Chao2, 
         bobcat_results$SE_Chao2, 
         boar_results$SE_Chao2),
  coverage = c(coyote_results$Coverage, 
               bobcat_results$Coverage, 
               boar_results$Coverage)
)

estimate_samples_to_chao <- function(accum_obj, chao_est) {
  sites <- accum_obj$sites
  richness <- accum_obj$richness
  if(chao_est > max(richness)) {
    x0 <- sites[length(sites)-1]
    x1 <- sites[length(sites)]
    y0 <- richness[length(sites)-1]
    y1 <- richness[length(sites)]
    x1 + (chao_est - y1) * (x1 - x0) / (y1 - y0)
  } else {
    sites[which(richness >= chao_est)[1]]
  }
}

samples_needed <- tibble(
  host = c("Coyote", "Bobcat", "Wild Boar"),
  samples_needed = c(
    estimate_samples_to_chao(coyote_accum, coyote_results$Chao2_estimate),
    estimate_samples_to_chao(bobcat_accum, bobcat_results$Chao2_estimate),
    estimate_samples_to_chao(boar_accum, boar_results$Chao2_estimate)
  )
)
samples_needed


iNEXT_out$AsyEst

shannon_diversity <- iNEXT_out$AsyEst |> 
  filter(Diversity == 'Shannon diversity')

species_richness <- iNEXT_out$AsyEst |> 
  filter(Diversity == 'Species richness')



```



```{r}
coverage_df <- iNEXT_out$iNextEst

coverage_summary <- iNEXT_out$iNextEst$size_based %>%
  filter(Method %in% c("Observed", "Extrapolation")) %>%
  group_by(Assemblage) %>%
  summarise(
    Observed_Coverage = round(SC[Method == "Observed"][1] * 100, 1),
    Expected_Coverage = round(max(SC[Method == "Extrapolation"], na.rm = TRUE) * 100, 1),
    Percent_of_Expected = round((Observed_Coverage / Expected_Coverage) * 100, 1)
  )

coverage_summary


ggiNEXT(iNEXT_out, type = 3) +
  ggtitle("Sample coverage vs. number of samples")

```

```{r}
sest_summary <- iNEXT_out$iNextEst$size_based %>%
  filter(Method %in% c("Observed", "Extrapolation")) %>%
  group_by(Assemblage) %>%
  summarise(
    Observed_Sest = qD[Method == "Observed"][1],
    Total_Sest = max(qD[Method == "Extrapolation"], na.rm = TRUE),
    Percent_Species_Sampled = round((Observed_Sest / Total_Sest) * 100, 1)
  )

sest_summary
```


could not figure out how to use iNEXT to extrapolate sample numbers needed for chao2 (expected prey density) estimate
```{r}
# # Extrapolate to 2x current sample size
# iNEXT_out <- iNEXT(iNEXT_list, q = 0, datatype = "incidence_freq", size = 100 * nrow(coyote_diet_mat))
# 
# # Extract the iNEXT output dataframe
# iNEXT_df_coyote <- iNEXT_out$iNextEst[[1]]  # first element = coyote
# 
# head(iNEXT_df)
# 
# 
# 
# 
# #Find the sample size where richness reaches Chao2
# chao2_target <- coyote_results$Chao2_estimate
# 
# # Find the smallest sample size where predicted richness >= Chao2
# coyote_samples_needed <- min(iNEXT_df$t[iNEXT_df_coyote$qD >= chao2_target])
# coyote_samples_needed
# 


```

Do this for bobcat
```{r}
# # Extrapolate to 2x current sample size
# iNEXT_out <- iNEXT(iNEXT_list, q = 0, datatype = "incidence_freq", size = 100 * nrow(coyote_diet_mat))
# 
# # Extract the iNEXT output dataframe
# iNEXT_df_bobcat <- iNEXT_out$iNextEst[[2]]  
# 
# head(iNEXT_df_bobcat)
# 
# 
# 
# 
# # Chao2 target
# chao2_target <- bobcat_results$Chao2_estimate
# 
# # Extract bobcat predicted richness vs sample size
# iNEXT_df_bobcat <- iNEXT_out$iNextEst[[2]]
# 
# # Find smallest t where predicted richness >= Chao2
# bobcat_samples_needed <- min(iNEXT_df_bobcat$t[iNEXT_df_bobcat$qD >= chao2_target])
# 
# # Check result
# bobcat_samples_needed


```

