---
title: "ANOSIM"
author: "Grace Lewin"
date: "2025-10-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(vegan)
library(here)
library(patchwork)
```


Import data
```{r}
gav <- read_csv(here("data", "grouped_prey_sites_host.csv"))
```


Step 1: Prepare data for ANOSIM
We’ll extract the prey community matrix and compute Jaccard dissimilarity.

```{r}
reordered_gav <- gav %>%
  mutate(count = rowSums(select(., coyote:black_rat))) |> 
  select(
    sample_id, host, marine_present, terrestrial_present, diet_category, site, lon, lat, count,
    coyote:black_rat  # move all prey species columns to the end
  ) |> 
  filter(count > 0) 

prey_matrix <- reordered_gav %>%
  select(coyote:black_rat)

prey_matrix <- prey_matrix[rowSums(prey_matrix) > 0, ]


# Compute Jaccard distance
diet_dist <- vegdist(prey_matrix, method = "jaccard", binary = TRUE)

```

Step 2: Run ANOSIM
```{r}
# Run ANOSIM using predator as grouping factor
anosim_result <- anosim(diet_dist, grouping = reordered_gav$host)

# View output
summary(anosim_result)
plot(anosim_result, col = viridis::viridis(3, option = "D"))
```


```{r}
boxplot(anosim_result$dis.rank ~ anosim_result$class.vec,
        notch = FALSE, 
        col = viridis::viridis(3, option = "C"),
        ylab = "Ranked dissimilarities",
        xlab = "Predator Species",
        main = "ANOSIM Rank Dissimilarity Plot")
```

```{r}
# Extract and tidy up data from ANOSIM object
anosim_df <- data.frame(
  Rank = anosim_result$dis.rank,
  GroupComparison = anosim_result$class.vec
)

# Reorder factor levels for nicer plot
anosim_df$GroupComparison <- factor(anosim_df$GroupComparison,
                                    levels = unique(anosim_df$GroupComparison))


# Extract R statistic and significance
R_value <- anosim_result$statistic
p_value <- anosim_result$signif

cat("Global R =", round(R_value, 3), " | p =", p_value, "\n")

base_colors <- brewer.pal(5, "Set2")
fills <- sapply(base_colors, function(col) adjustcolor(col, alpha.f = 0.5))

# Create ggplot version of the ANOSIM boxplot
anosim_plot <- ggplot(anosim_df, aes(x = GroupComparison, y = Rank, fill = GroupComparison)) +
  geom_boxplot(outlier.shape = 21, color = "black", width = 0.7) +
  #scale_fill_viridis_d(option = "D") +
  #scale_fill_manual(values = c("#FF9999", "#99CCFF", "#99FF99", "red"))+
  scale_fill_manual(values = c("#FC8D62", "#E78AC3", "#66C2A5", "#8DA0CB"))+
  labs(
    title = paste0("R = ", round(R_value, 3), ", p = ", p_value),
    x = "",
    y = "Ranked Dissimilarity"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

anosim_plot

ggsave(plot = anosim_plot, filename = here("figs", "anosim.png"), dpi = 600, units = "in", width = 5, height = 4)
```

```{r}



# Extract the rank distributions
anosim_df <- data.frame(
  Dissimilarity = c(anosim_result$dis.rank[anosim_result$class.vec == 1], 
                    anosim_result$dis.rank[anosim_result$class.vec == 2]),
  Type = rep(c("Within", "Between"), 
             c(sum(anosim_result$class.vec == 1), sum(anosim_result$class.vec == 2)))
)

ggplot(anosim_df, aes(x = Type, y = Dissimilarity, fill = Type)) +
  geom_boxplot(alpha = 0.7, width = 0.5) +
  theme_minimal(base_size = 14) +
  labs(
    title = paste0("ANOSIM: R = ", round(R_value, 3), ", p = ", p_value),
    x = NULL,
    y = "Ranked Dissimilarity"
  ) +
  scale_fill_manual(values = c("skyblue3", "tomato3")) +
  theme(legend.position = "none")
```



```{r}
# Check dispersion homogeneity
disp <- betadisper(diet_dist, reordered_gav$host)
anova(disp)
plot(disp)

```



try it with NDMS
```{r}
nmds <- metaMDS(diet_dist, k = 2, trymax = 100)

nmds_points <- as.data.frame(scores(nmds, display = "sites"))
nmds_points$host <- reordered_gav$host

ggplot(nmds_points, aes(x = NMDS1, y = NMDS2, color = host)) +
  geom_point(size = 3, alpha = 0.2) +
  stat_ellipse(level = 0.68, linetype = 2) +  # 68% confidence ellipse per predator
  theme_minimal(base_size = 14) +
  labs(
    title = "NMDS of Predator Diet Composition (Jaccard Distance)",
    subtitle = paste("Stress =", round(nmds$stress, 3)),
    color = "Predator"
  )

```


### try anosim with just coyote data at different sites. then try just coytote and boar to see if i get a result

```{r}
reordered_gav2_coyote <- gav %>%
  mutate(count = rowSums(select(., coyote:black_rat))) |> 
  select(
    sample_id, host, marine_present, terrestrial_present, diet_category, site, lon, lat, count,
    coyote:black_rat  # move all prey species columns to the end
  ) |> 
  filter(count > 0) |> 
  mutate(site = case_when(
    site == "Coal Oil Point" ~ "COP/NCOS",
    site == "North Campus Open Space" ~ "COP/NCOS",
    TRUE ~ site
  )) |> 
  filter(host == "coyote")

prey_matrix_coyote <- reordered_gav2_coyote %>%
  select(coyote:black_rat)

prey_matrix_coyote <- prey_matrix_coyote[rowSums(prey_matrix_coyote) > 0, ]


# Compute Jaccard distance
diet_dist_coyote <- vegdist(prey_matrix_coyote, method = "jaccard", binary = TRUE)

```

Step 2: Run ANOSIM
```{r}
# Run ANOSIM using predator as grouping factor
anosim_result_coyote <- anosim(diet_dist_coyote, grouping = reordered_gav2_coyote$site)

# View output
summary(anosim_result_coyote)
plot(anosim_result_coyote, col = viridis::viridis(3, option = "D"))

boxplot(anosim_result_coyote$dis.rank ~ anosim_result_coyote$class.vec,
        notch = FALSE, 
        col = viridis::viridis(3, option = "C"),
        ylab = "Ranked dissimilarities",
        xlab = "Predator Species",
        main = "ANOSIM Rank Dissimilarity Plot")

# Extract and tidy up data from ANOSIM object
anosim_df_coyote <- data.frame(
  Rank = anosim_result_coyote$dis.rank,
  GroupComparison = anosim_result_coyote$class.vec
)

# Recode GroupComparison for plotting
anosim_df_coyote <- anosim_df_coyote %>%
  mutate(GroupComparison = recode(GroupComparison,
                                  "Dangermond" = "Remote",
                                  "COP/NCOS" = "Suburban"))

# Reorder factor levels for nicer plot
anosim_df_coyote$GroupComparison <- factor(anosim_df_coyote$GroupComparison,
                                    levels = unique(anosim_df_coyote$GroupComparison))


# Extract R statistic and significance
R_value <- anosim_result_coyote$statistic
p_value <- anosim_result_coyote$signif

cat("Global R =", round(R_value, 3), " | p =", p_value, "\n")

base_colors <- brewer.pal(5, "Set2")
fills <- sapply(base_colors, function(col) adjustcolor(col, alpha.f = 0.5))

# Create ggplot version of the ANOSIM boxplot
anosim_plot_coyote <- ggplot(anosim_df_coyote, aes(x = GroupComparison, y = Rank, fill = GroupComparison)) +
  geom_boxplot(outlier.shape = 21, color = "black", width = 0.7) +
  #scale_fill_viridis_d(option = "D") +
  #scale_fill_manual(values = c("#FF9999", "#99CCFF", "#99FF99", "red"))+
  scale_fill_manual(values = c("#FC8D62", "#e8df5a", "#cc592d"))+
  labs(
    title = paste0("R = ", round(R_value, 3), "\np = ", p_value),
    x = "",
    y = "Ranked Dissimilarity of Coyote Diets"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

anosim_plot_coyote

ggsave(filename = here("figs", "coyote_anosim.png"), plot = anosim_plot_coyote, dpi = 600, units = "in", width = 2, height = 7)
```


This is just for coyote and bobcat
```{r}
reordered_gav2_coyote_bobcat <- gav %>%
  mutate(count = rowSums(select(., coyote:black_rat))) |> 
  select(
    sample_id, host, marine_present, terrestrial_present, diet_category, site, lon, lat, count,
    coyote:black_rat  # move all prey species columns to the end
  ) |> 
  filter(count > 0) |> 
  mutate(site = case_when(
    site == "Coal Oil Point" ~ "COP/NCOS",
    site == "North Campus Open Space" ~ "COP/NCOS",
    TRUE ~ site
  )) |> 
  filter(host %in% c("coyote", "bobcat"))

prey_matrix_coyote_bobcat <- reordered_gav2_coyote_bobcat %>%
  select(coyote:black_rat)

prey_matrix_coyote_bobcat <- prey_matrix_coyote_bobcat[rowSums(prey_matrix_coyote_bobcat) > 0, ]


# Compute Jaccard distance
diet_dist_coyote_bobcat <- vegdist(prey_matrix_coyote_bobcat, method = "jaccard", binary = TRUE)


```

Step 2: Run ANOSIM
```{r}
# Run ANOSIM using predator as grouping factor
anosim_result_coyote_bobcat <- anosim(diet_dist_coyote_bobcat, grouping = reordered_gav2_coyote_bobcat$host)

# View output
summary(anosim_result_coyote_bobcat)
plot(anosim_result_coyote_bobcat, col = viridis::viridis(3, option = "D"))

boxplot(anosim_result_coyote_bobcat$dis.rank ~ anosim_result_coyote_bobcat$class.vec,
        notch = FALSE, 
        col = viridis::viridis(3, option = "C"),
        ylab = "Ranked dissimilarities",
        xlab = "Predator Species",
        main = "ANOSIM Rank Dissimilarity Plot")

# Extract and tidy up data from ANOSIM object
anosim_df_coyote_bobcat <- data.frame(
  Rank = anosim_result_coyote_bobcat$dis.rank,
  GroupComparison = anosim_result_coyote_bobcat$class.vec
)

# # Recode GroupComparison for plotting
# anosim_df_coyote_bobcat <- anosim_df_coyote_bobcat %>%
#   mutate(GroupComparison = recode(GroupComparison,
#                                   "Dangermond" = "Coyote\nWild",
#                                   "COP/NCOS" = "Coyote\nUrban"))

# Reorder factor levels for nicer plot
anosim_df_coyote_bobcat$GroupComparison <- factor(anosim_df_coyote_bobcat$GroupComparison,
                                    levels = unique(anosim_df_coyote_bobcat$GroupComparison))


# Extract R statistic and significance
R_value <- anosim_result_coyote_bobcat$statistic
p_value <- anosim_result_coyote_bobcat$signif

cat("Global R =", round(R_value, 3), " | p =", p_value, "\n")

base_colors <- brewer.pal(5, "Set2")
fills <- sapply(base_colors, function(col) adjustcolor(col, alpha.f = 0.5))

# Create ggplot version of the ANOSIM boxplot
anosim_plot_coyote_bobcat <- ggplot(anosim_df_coyote_bobcat, aes(x = GroupComparison, y = Rank, fill = GroupComparison)) +
  geom_boxplot(outlier.shape = 21, color = "black", width = 0.7) +
  #scale_fill_viridis_d(option = "D") +
  #scale_fill_manual(values = c("#FF9999", "#99CCFF", "#99FF99", "red"))+
  scale_fill_manual(values = c("#FC8D62", "#E78AC3", "#66C2A5"))+
  labs(
    title = paste0("R = ", round(R_value, 3), ", p = ", p_value),
    x = "",
    y = "Ranked Dissimilarity"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

anosim_plot_coyote_bobcat

```


Between just coyote and boar:
```{r}
reordered_gav2_coyote_boar <- gav %>%
  mutate(count = rowSums(select(., coyote:black_rat))) |> 
  select(
    sample_id, host, marine_present, terrestrial_present, diet_category, site, lon, lat, count,
    coyote:black_rat  # move all prey species columns to the end
  ) |> 
  filter(count > 0) |> 
  mutate(site = case_when(
    site == "Coal Oil Point" ~ "COP/NCOS",
    site == "North Campus Open Space" ~ "COP/NCOS",
    TRUE ~ site
  )) |> 
  filter(host %in% c("coyote", "wild boar"))

prey_matrix_coyote_boar <- reordered_gav2_coyote_boar %>%
  select(coyote:black_rat)

prey_matrix_coyote_boar <- prey_matrix_coyote_boar[rowSums(prey_matrix_coyote_boar) > 0, ]


# Compute Jaccard distance
diet_dist_coyote_boar <- vegdist(prey_matrix_coyote_boar, method = "jaccard", binary = TRUE)

```

Step 2: Run ANOSIM
```{r}
# Run ANOSIM using predator as grouping factor
anosim_result_coyote_boar <- anosim(diet_dist_coyote_boar, grouping = reordered_gav2_coyote_boar$host)

# View output
summary(anosim_result_coyote_boar)
plot(anosim_result_coyote_boar, col = viridis::viridis(3, option = "D"))

boxplot(anosim_result_coyote_boar$dis.rank ~ anosim_result_coyote_boar$class.vec,
        notch = FALSE, 
        col = viridis::viridis(3, option = "C"),
        ylab = "Ranked dissimilarities",
        xlab = "Predator Species",
        main = "ANOSIM Rank Dissimilarity Plot")


# Extract and tidy up data from ANOSIM object
anosim_df_coyote_boar <- data.frame(
  Rank = anosim_result_coyote_boar$dis.rank,
  GroupComparison = anosim_result_coyote_boar$class.vec
)

# # Recode GroupComparison for plotting
# anosim_df_coyote_bobcat <- anosim_df_coyote_bobcat %>%
#   mutate(GroupComparison = recode(GroupComparison,
#                                   "Dangermond" = "Coyote\nWild",
#                                   "COP/NCOS" = "Coyote\nUrban"))

# Reorder factor levels for nicer plot
anosim_df_coyote_boar$GroupComparison <- factor(anosim_df_coyote_boar$GroupComparison,
                                    levels = unique(anosim_df_coyote_boar$GroupComparison))


# Extract R statistic and significance
R_value <- anosim_result_coyote_boar$statistic
p_value <- anosim_result_coyote_boar$signif

cat("Global R =", round(R_value, 3), " | p =", p_value, "\n")

base_colors <- brewer.pal(5, "Set2")
fills <- sapply(base_colors, function(col) adjustcolor(col, alpha.f = 0.5))

# Create ggplot version of the ANOSIM boxplot
anosim_plot_coyote_boar <- ggplot(anosim_df_coyote_boar, aes(x = GroupComparison, y = Rank, fill = GroupComparison)) +
  geom_boxplot(outlier.shape = 21, color = "black", width = 0.7) +
  #scale_fill_viridis_d(option = "D") +
  #scale_fill_manual(values = c("#FF9999", "#99CCFF", "#99FF99", "red"))+
  scale_fill_manual(values = c("#FC8D62", "#E0A8B8", "#8DA0CB"))+
  labs(
    title = paste0("R = ", round(R_value, 3), ", p = ", p_value),
    x = "",
    y = "Ranked Dissimilarity"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

anosim_plot_coyote_boar

```


Between just bobcat and boar:
```{r}
reordered_gav2_bobcat_boar <- gav %>%
  mutate(count = rowSums(select(., coyote:black_rat))) |> 
  select(
    sample_id, host, marine_present, terrestrial_present, diet_category, site, lon, lat, count,
    coyote:black_rat  # move all prey species columns to the end
  ) |> 
  filter(count > 0) |> 
  mutate(site = case_when(
    site == "Coal Oil Point" ~ "COP/NCOS",
    site == "North Campus Open Space" ~ "COP/NCOS",
    TRUE ~ site
  )) |> 
  filter(host %in% c("bobcat", "wild boar"))

prey_matrix_bobcat_boar <- reordered_gav2_bobcat_boar %>%
  select(coyote:black_rat)

#prey_matrix_bobcat_boar <- reordered_gav2_bobcat_boar[rowSums(reordered_gav2_bobcat_boar) > 0, ]


# Compute Jaccard distance
diet_dist_bobcat_boar <- vegdist(prey_matrix_bobcat_boar, method = "jaccard", binary = TRUE)

```

Step 2: Run ANOSIM
```{r}
# Run ANOSIM using predator as grouping factor
anosim_result_bobcat_boar <- anosim(diet_dist_bobcat_boar, grouping = reordered_gav2_bobcat_boar$host)

# View output
summary(anosim_result_bobcat_boar)
plot(anosim_result_bobcat_boar, col = viridis::viridis(3, option = "D"))

boxplot(anosim_result_bobcat_boar$dis.rank ~ anosim_result_bobcat_boar$class.vec,
        notch = FALSE, 
        col = viridis::viridis(3, option = "C"),
        ylab = "Ranked dissimilarities",
        xlab = "Predator Species",
        main = "ANOSIM Rank Dissimilarity Plot")

# Extract and tidy up data from ANOSIM object
anosim_df_bobcat_boar <- data.frame(
  Rank = anosim_result_bobcat_boar$dis.rank,
  GroupComparison = anosim_result_bobcat_boar$class.vec
)

# # Recode GroupComparison for plotting
# anosim_df_coyote_bobcat <- anosim_df_coyote_bobcat %>%
#   mutate(GroupComparison = recode(GroupComparison,
#                                   "Dangermond" = "Coyote\nWild",
#                                   "COP/NCOS" = "Coyote\nUrban"))

# Reorder factor levels for nicer plot
anosim_df_bobcat_boar$GroupComparison <- factor(anosim_df_bobcat_boar$GroupComparison,
                                    levels = unique(anosim_df_bobcat_boar$GroupComparison))


# Extract R statistic and significance
R_value <- anosim_result_bobcat_boar$statistic
p_value <- anosim_result_bobcat_boar$signif

cat("Global R =", round(R_value, 3), " | p =", p_value, "\n")

base_colors <- brewer.pal(5, "Set2")
fills <- sapply(base_colors, function(col) adjustcolor(col, alpha.f = 0.5))

# Create ggplot version of the ANOSIM boxplot
anosim_plot_bobcat_boar <- ggplot(anosim_df_bobcat_boar, aes(x = GroupComparison, y = Rank, fill = GroupComparison)) +
  geom_boxplot(outlier.shape = 21, color = "black", width = 0.7) +
  #scale_fill_viridis_d(option = "D") +
  #scale_fill_manual(values = c("#FF9999", "#99CCFF", "#99FF99", "red"))+
  scale_fill_manual(values = c("#FC8D62", "#E0A8B8", "#8DA0CB"))+
  labs(
    title = paste0("R = ", round(R_value, 3), ", p = ", p_value),
    x = "",
    y = "Ranked Dissimilarity"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

anosim_plot_bobcat_boar
```

```{r}
all_anosim_plots <- (anosim_plot) / (anosim_plot_coyote_bobcat + anosim_plot_coyote_boar)

all_anosim_plots <- (anosim_plot_coyote_bobcat / anosim_plot_coyote_boar)

ggsave(plot = all_anosim_plots, filename = here("figs", "all_anosim_plots.png"), dpi = 600, width = 6, height = 6)


```

