---
title: "1_import_and_clean"
output:
  pdf_document: default
  html_document: default
date: "2023-08-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This section of the pipeline starts with the illumina sequencing dataframes that come out of Kenen's filtering pipeline and merges them together. The ending dataframe is a .csv called "alaska".

TO-DO
CHECK KNP SORTED INTO ANIAKCHAK SAMPS

## Import illumina run data (.csv)

```{r}
## Katmai samples collected in 2019 by Kelsey and Gretchen

# HiSeq1527
run19 <- read.csv("Run19-Ellen.csv", header = T)
#unique(run19$Sample) #171

## Katmai samples collected in 2021

# Nextseq46
run23 <- read.csv("Run23-Ellen.csv", header = T)
#unique(run23$Sample) #215

# Miseq874
run24 <- read.csv("Run24-Ellen.csv", header = T)
#unique(run24$Sample) #106

# Nextseq127
run25 <- read.csv("Run25-Ellen.csv", header = T)
#unique(run25$Sample) #157

# Miseq938
run26 <- read.csv("Run26-Ellen.csv", header = T)
#unique(run26$Sample) #168

# Nextseq146 (libs K and M)
run27 <- read.csv("Run27-Ellen.csv", header = T)
#unique(run27$Sample) #221

# Nextseq146 (libs L and N)
run27LN <- read.csv("Run27_LN_Alaska_1pct.csv")
#unique(run27LN$Sample) #220

## Lake Clark and Glacier Bay samples collected in 2022

# Miseq994 (run35)
run35 <- read.csv("Run35_Miseq994_DymitResults.csv", header = T)
run35$Run = "35"
#unique(run35$Sample) #85

# NextSeq216 (run36)
run36 <- read.csv("Run36_NextSeq216_Dymit_Results.csv", header = T)
run36$Run = "36"
#unique(run36$Sample) #214

# NextSeq223 (run37)
run37 <- read.csv("Run37_12S_Vertebrates_Alaska_1pct.csv", header = T)
#unique(run37$Sample) #204

# NextSeq239
run40 <- read.csv("DYMIT_Run40_NextSeq239_Results.csv", header = T)
run40$Run = "40"
#unique(run40$Sample) #216

# NextSeq248
run42 <- read.csv("NextSeq248_Run42_Dymit_Results.csv", header = T)
run42$Run = "42"
#unique(run42$Sample) #10

# NextSeq315
run52 <- read.csv("IlluminaRun52_Nextseq315_DYMIT.csv", header = T)
run52$Run = "52"
#unique(run52$Sample) #22

# Glacier Bay Samples collected in 2023

# NextSeq350 
run57 <- read.csv("IlluminaRun57_Nextseq350_DYMIT.csv", header = T)
run57$Run = "57"
#unique(run57$Sample) #78

# NextSeq357
run58 <- read.csv("Run58_12S_Vertebrates_Alaska_1pct.csv")
#unique(run58$Sample) #26

# NextSeq373
run59 <- read.csv("IlluminaRun59_Nextseq373_DYMIT_Results.csv")
run59$Run = "59"
#unique(run59$Sample) #123

# NextSeq400
run68 <- read.csv("IlluminaRun68_Nextseq400_Results_Dymit.csv")
run68$Run = "68"
#unique(run68$Sample) #58

```


## Standardize columns so that dataframes can be merged

```{r}
## Check column names for each dataframe
colnames(run19)
colnames(run23)
colnames(run24)
colnames(run25)
colnames(run26)
colnames(run27)
colnames(run27LN)

colnames(run58)
run58$Run = "58"
run58$Barcode_region = "12S" 
run58$Taxa = "Vertebrates"
run58$Geographic_region = "Alaska"
run58$Project = "KNP"
run58$Percentage_filter = "1"

colnames(run37)
run37$Run = "37"
run37$Barcode_region = "12S" 
run37$Taxa = "Vertebrates"
run37$Geographic_region = "Alaska"
run37$Project = "KNP"
run37$Percentage_filter = "1"

nextseq <- rbind(run23, run25, run27, run27LN, run37, run58)
miseq <- rbind(run19, run24, run26)

## Split PCR rep read counts column into three columns

nextseq[c('a', 'b', 'c')] <- str_split_fixed(nextseq$PCR_rep_read_counts, ',', 3)
nextseq$a_rep = NA
nextseq$b_rep = NA
nextseq$c_rep = NA
nextseq$bb_rep = NA
nextseq$cb_rep = NA
nextseq <- mutate(nextseq, 
                 a_rep = ifelse(grepl("a:",a), a, NA), 
                 b_rep = ifelse(grepl("b:",a), a, NA),
                 bb_rep = ifelse(grepl("b:",b), b, NA),
                 cb_rep = ifelse(grepl("c:",b), b, NA),
                 c_rep = ifelse(grepl("c:",c), c, NA))
nextseq$b_rep <- with(nextseq, coalesce(b_rep,bb_rep))
nextseq$c_rep <- with(nextseq, coalesce(c_rep,cb_rep))
nextseq <- nextseq[ , ! names(nextseq) %in% c("bb_rep", "cb_rep", "a", "b", "c")]

nextseq$a_rep <- parse_number(nextseq$a_rep)
nextseq$b_rep <- parse_number(nextseq$b_rep)
nextseq$c_rep <- parse_number(nextseq$c_rep)

nextseq <- subset(nextseq, select = -PCR_rep_read_counts)

miseq[c('a', 'b', 'c')] <- str_split_fixed(miseq$PCR_rep_read_counts, ',', 3)
miseq$a_rep = NA
miseq$b_rep = NA
miseq$c_rep = NA
miseq$bb_rep = NA
miseq$cb_rep = NA
miseq <- mutate(miseq, 
                 a_rep = ifelse(grepl("a:",a), a, NA), 
                 b_rep = ifelse(grepl("b:",a), a, NA),
                 bb_rep = ifelse(grepl("b:",b), b, NA),
                 cb_rep = ifelse(grepl("c:",b), b, NA),
                 c_rep = ifelse(grepl("c:",c), c, NA))
miseq$b_rep <- with(miseq, coalesce(b_rep,bb_rep))
miseq$c_rep <- with(miseq, coalesce(c_rep,cb_rep))
miseq <- miseq[ , ! names(miseq) %in% c("bb_rep", "cb_rep", "a", "b", "c")]

miseq$a_rep <- parse_number(miseq$a_rep)
miseq$b_rep <- parse_number(miseq$b_rep)
miseq$c_rep <- parse_number(miseq$c_rep)

miseq <- subset(miseq, select = -PCR_rep_read_counts)

colnames(miseq)
colnames(nextseq)

run35 <- run35 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run36 <- run36 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run40 <- run40 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run42 <- run42 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run52 <- run52 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run57 <- run57 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run59 <- run59 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")
run68 <- run68 %>% 
        rename("a_rep" = "RepA",
               "b_rep" = "RepB",
               "c_rep" = "RepC")

miseq <- rbind(miseq, run35)
nextseq <- rbind(nextseq, run36, run40, run42, run52, run57, run59, run68)

# remove plant reads
nextseq <- nextseq %>% filter(!Barcode_region == "trnL")
```

## Remove common lab contaminants & BLAST errors
## TO DO: CHECK REMOVES TO MAKE SURE THEY ARENT REAL STUFF

```{r}
#unique(miseq$Sample) #526
#unique(nextseq$Sample) #1727

## Human contamination
miseq <- miseq[!grepl("Homo sapiens", miseq$Local_taxa),]
nextseq <- nextseq[!grepl("Homo sapiens", nextseq$Local_taxa),]
miseq <- miseq[!grepl("Hominidae", miseq$Local_taxa),]
nextseq <- nextseq[!grepl("Hominidae", nextseq$Local_taxa),]
miseq <- miseq[!grepl("Primates", miseq$Local_taxa),]
nextseq <- nextseq[!grepl("Primates", nextseq$Local_taxa),]

## Weird shit
#giant flying squirrel
miseq <- miseq[!grepl("Petaurista philippensis", miseq$Best_match_references),]
nextseq <- nextseq[!grepl("Petaurista philippensis", nextseq$Best_match_references),]
#sun bear
miseq <- miseq[!grepl("Helarctos malayanus", miseq$Best_match_references),]
nextseq <- nextseq[!grepl("Helarctos malayanus", nextseq$Best_match_references),]
#polar bear
miseq <- miseq[!grepl("Ursus maritimus", miseq$Best_match_references),]
nextseq <- nextseq[!grepl("Ursus maritimus", nextseq$Best_match_references),]
#elk
miseq <- miseq[!grepl("Cervus canadensis", miseq$Best_match_references),]
nextseq <- nextseq[!grepl("Cervus canadensis", nextseq$Best_match_references),]

## Unknown sequences
miseq_NSSF <- miseq[grepl("No significant similarity found", miseq$Local_taxa),]
nextseq_NSSF <- nextseq[grepl("No significant similarity found", nextseq$Local_taxa),]
Alaska_NSSF <- rbind(miseq_NSSF, nextseq_NSSF)
rm(miseq_NSSF)
rm(nextseq_NSSF)

miseq <- miseq[!grepl("No significant similarity found", miseq$Local_taxa),]
nextseq <- nextseq[!grepl("No significant similarity found", nextseq$Local_taxa),]

## Domestic cow
miseq <- miseq[!grepl("Bovidae", miseq$Local_taxa),]
nextseq <- nextseq[!grepl("Bovidae", nextseq$Local_taxa),]

#unique(miseq$Local_species)
#unique(nextseq$Local_species)

```

## Create new read variables

```{r}
# readcount = total number of reads for all reps per match
nextseq <- nextseq %>% mutate(readcount = select(., a_rep:c_rep) %>% rowSums(na.rm=TRUE))
miseq <- miseq %>% mutate(readcount = select(., a_rep:c_rep) %>% rowSums(na.rm=TRUE))

# totalsum = total number of reads per sample
nextseq <- nextseq %>% 
  group_by(Sample, Percentage_filter) %>% 
  mutate(totalsum = sum(readcount)) %>%
  ungroup()
miseq <- miseq %>% 
  group_by(Sample, Percentage_filter) %>% 
  mutate(totalsum = sum(readcount)) %>%
  ungroup()

# propreads = proportion of total reads per sample comprised by one species
nextseq <- nextseq %>% 
  group_by(Sample, Percentage_filter) %>% 
  mutate(propreads  = readcount/totalsum) %>%
  ungroup()
miseq <- miseq %>% 
  group_by(Sample, Percentage_filter) %>% 
  mutate(propreads  = readcount/totalsum) %>%
  ungroup()

```


## Filter out low-read samples from nextseq data

```{r}
## only for nextseq!!!
# remove samples with fewer than 1000 reads
#nextseq_lowreads <- subset(nextseq, totalsum < 1000)
#unique(lowreads$Sample) #10
#write.csv(lowreads,"lowreads.csv", row.names=F)
#nextseq <- subset(nextseq, totalsum > 1000)
#unique(KATMnextseq$Sample) #582
#write.csv(KATMnextseq,"KATMnextseq.csv", row.names=F)
```

## Merge miseq df with nextseq df

```{r}
alaska <- rbind(nextseq, miseq)
unique(alaska$Sample) #2228 samples
```

## Filter out species with low reads
```{r}
#alaska_lowreads <- subset(alaska, readcount < 500)
#alaska <- subset(alaska, readcount > 500)
```


## Create column for sites

```{r}
alaska <- alaska %>% drop_na(Sample)

## Merge in site IDs for 2019 Katmai samples
KATM19sites <- read.csv("KATM19sites.csv", header = T)
alaska <- merge(alaska, KATM19sites, by = "Sample", all = TRUE)

GLBAsites <- read.csv("GLBAsites_merge.csv", header = T)
alaska <- merge(alaska, GLBAsites, by = "Sample", all = TRUE)

alaska <- alaska %>%
  mutate(site = coalesce(site.x, site.y))
alaska <- subset(alaska, select = -site.x)
alaska <- subset(alaska, select = -site.y)

alaska$region = NA
alaska$park = NA

# Katmai field sites
alaska$region[grep("Becharof", alaska$site)] = as.character("coast")
alaska$park[grep("Becharof", alaska$site)] = as.character("KATM")

alaska$region[grep("Fure's Cabin", alaska$site)] = as.character("interior")
alaska$park[grep("Fure's Cabin", alaska$site)] = as.character("KATM")

alaska$site[grep("DB", alaska$Sample)] = as.character("Dakavak Bay")
alaska$region[grep("Dakavak Bay", alaska$site)] = as.character("coast")
alaska$park[grep("Dakavak Bay", alaska$site)] = as.character("KATM")

alaska$site[grep("WL", alaska$Sample)] = as.character("Hallo Bay")
alaska$site[grep("HB", alaska$Sample)] = as.character("Hallo Bay")
alaska$region[grep("Hallo Bay", alaska$site)] = as.character("coast")
alaska$park[grep("Hallo Bay", alaska$site)] = as.character("KATM")

alaska$site[grep("RT", alaska$Sample)] = as.character("Ridge Trail")
alaska$region[grep("Ridge Trail", alaska$site)] = as.character("interior")
alaska$park[grep("Ridge Trail", alaska$site)] = as.character("KATM")

alaska$site[grep("BT", alaska$Sample)] = as.character("Bluff Trail")
alaska$region[grep("Bluff Trail", alaska$site)] = as.character("interior")
alaska$park[grep("Bluff Trail", alaska$site)] = as.character("KATM")

alaska$site[grep("SW", alaska$Sample)] = as.character("Swikshak")
alaska$region[grep("Swikshak", alaska$site)] = as.character("coast")
alaska$park[grep("Swikshak", alaska$site)] = as.character("KATM")

alaska$site[grep("CD", alaska$Sample)] = as.character("Cape Douglas")
alaska$region[grep("Cape Douglas", alaska$site)] = as.character("coast")
alaska$park[grep("Cape Douglas", alaska$site)] = as.character("KATM")

alaska$site[grep("KB", alaska$Sample)] = as.character("Katmai Bay")
alaska$region[grep("Katmai Bay", alaska$site)] = as.character("coast")
alaska$park[grep("Katmai Bay", alaska$site)] = as.character("KATM")

alaska$site[grep("BC", alaska$Sample)] = as.character("Brooks Camp")
alaska$region[grep("Brooks Camp", alaska$site)] = as.character("interior")
alaska$park[grep("Brooks Camp", alaska$site)] = as.character("KATM")

alaska$site[grep("KS", alaska$Sample)] = as.character("Kamishak")
alaska$site[grep("K5", alaska$Sample)] = as.character("Kamishak")
alaska$region[grep("Kamishak", alaska$site)] = as.character("coast")
alaska$park[grep("Kamishak", alaska$site)] = as.character("KATM")

alaska$site[grep("VR", alaska$Sample)] = as.character("Valley Road")
alaska$region[grep("Valley Road", alaska$site)] = as.character("interior")
alaska$park[grep("Valley Road", alaska$site)] = as.character("KATM")

alaska$site[grep("UNK", alaska$Sample)] = as.character("Unknown")
alaska$region[grep("Unknown", alaska$site)] = as.character("coast")
alaska$park[grep("UNK", alaska$Sample)] = as.character("KATM")

alaska$site[grep("YW", alaska$Sample)] = as.character("Kaguyak")
alaska$region[grep("Kaguyak", alaska$site)] = as.character("coast")
alaska$park[grep("Kaguyak", alaska$site)] = as.character("KATM")

# Kenai Fjords
alaska$site[grep("KEFJ", alaska$Sample)] = as.character("Kenai Fjords")
alaska$site[grep("68", alaska$Run)] = as.character("Kenai Fjords")
alaska$region[grep("Kenai Fjords", alaska$site)] = as.character("coast")
alaska$park[grep("Kenai Fjords", alaska$site)] = as.character("KEFJ")

# Aniakchak
alaska$site[grep("AN", alaska$Sample)] = as.character("Aniakchak")
alaska$region[grep("Aniakchak", alaska$site)] = as.character("coast")
alaska$park[grep("Aniakchak", alaska$site)] = as.character("ANIA")

# Lake Clark field sites
alaska$site[grep("SH", alaska$Sample)] = as.character("Shelter Creek")
alaska$region[grep("Shelter Creek", alaska$site)] = as.character("coast")
alaska$park[grep("Shelter Creek", alaska$site)] = as.character("LACL")

alaska$site[grep("KJ", alaska$Sample)] = as.character("Kijik")
alaska$region[grep("Kijik", alaska$site)] = as.character("interior")
alaska$park[grep("Kijik", alaska$site)] = as.character("LACL")

alaska$site[grep("CH", alaska$Sample)] = as.character("Chinitna Bay")
alaska$region[grep("Chinitna Bay", alaska$site)] = as.character("coast")
alaska$park[grep("Chinitna Bay", alaska$site)] = as.character("LACL")

alaska$site[grep("ACB", alaska$Sample)] = as.character("Unknown")
alaska$region[grep("ACB", alaska$Sample)] = as.character("coast")
alaska$park[grep("ACB", alaska$Sample)] = as.character("LACL")

alaska$site[grep("KO", alaska$Sample)] = as.character("Kontrash Lake")
alaska$region[grep("Kontrash Lake", alaska$site)] = as.character("interior")
alaska$park[grep("Kontrash Lake", alaska$site)] = as.character("LACL")

alaska$site[grep("SS", alaska$Sample)] = as.character("Silver Salmon Creek")
alaska$region[grep("Silver Salmon Creek", alaska$site)] = as.character("coast")
alaska$park[grep("Silver Salmon Creek", alaska$site)] = as.character("LACL")

alaska$site[grep("TX", alaska$Sample)] = as.character("Tuxedni Bay")
alaska$site[grep("NT", alaska$Sample)] = as.character("Tuxedni Bay")
alaska$region[grep("Tuxedni Bay", alaska$site)] = as.character("coast")
alaska$park[grep("Tuxedni Bay", alaska$site)] = as.character("LACL")

# Glacier Bay field sites

# Glacier Bay
alaska$park[grep("GLBA", alaska$Sample)] = as.character("GLBA")
alaska$region[grep("GLBA", alaska$Sample)] = as.character("coast")

checkna <- alaska[is.na(alaska$site), ] 
unique(checkna$Sample)
rm(checkna)

alaska <- alaska[complete.cases(alaska[,c("site")]),] #this drops blanks/NTCs

# isolate raw GLBA data
#GLBAraw <- alaska[grepl("Glacier Bay", alaska$site),]
#write.csv(GLBAraw, "GLBAraw.csv", row.names = FALSE)

# isolate raw KEFJ data
KEFJraw <- alaska[grepl("Kenai Fjords", alaska$site),]
write.csv(KEFJraw, "KEFJraw.csv", row.names = FALSE)
```

## Split by % filter
```{r}
alaska0.5 <- alaska[alaska$Percentage_filter == "0.5", ]
write.csv(alaska0.5,"alaska0.5.csv", row.names=F)
alaska1 <- alaska[alaska$Percentage_filter == "1", ]
write.csv(alaska1,"alaska1.csv", row.names=F)
```

## 0.5 filtering for samples containing fish reads only 
## TO DO: UPDATE THIS

```{r}
#fish0.5 <- alaska0.5[alaska0.5$Sample %in% c("CD44", "KB170", "KB183", "KB196", "KB199", "KB203", "KB251", "KB257", "KB264", "KB302", "KS43", "SW2.169", "SW2x170", "AN1", "CH11", "CH25", "CH26", "CH3", "CH42", "CH45", "KJ12", "KJ13", "KJ20", "SH14", "SH53", "SH7", "SH8", "SH98", "YW14", "YW15", "YW16", "YW17", "YW2", "YW22", "YW4", "YW8"), ]

#alaska1 <- alaska1[!alaska1$Sample %in% c("CD44", "KB170", "KB183", "KB196", "KB199", "KB203", "KB251", "KB257", "KB264", "KB302", "KS43", "SW2.169", "SW2x170", "AN1", "CH11", "CH25", "CH26", "CH3", "CH42", "CH45", "KJ12", "KJ13", "KJ20", "SH14", "SH53", "SH7", "SH8", "SH98", "YW14", "YW15", "YW16", "YW17", "YW2", "YW22", "YW4", "YW8"), ]

#alaska <- rbind(fish0.5, alaska1)
#rm(fish0.5)

#alaska <- alaska[ , c("Sample", "region", "park", "site", "Best_match_references", "Best_match_PID.mean", "Local_taxa", "Local_species", "a_rep", "b_rep", "c_rep", "readcount", "totalsum")]

```
